### 2011-07-14 cover_letter.tex
IJSR paper as of 2011-07-14

\documentclass{article}

\usepackage{url}

\begin{document}

We are very grateful to the reviewers for having accepted to review our 
work, for the care they took in examining it and their resulting 
comments which have been appreciated in order to improve the quality of
the paper.

In the following three sections, we try to address separately each of
the concerns of our three reviewers.

\section{Comments of the first reviewer}

\subsection{How robust is the approach with regard to perceptual recognition
errors?}

We are aware that perception is not robust and therefore one of our next steps
is to include a probabilistic layer to the approach (as already mentioned in
the discussion section). We believe though that a robot (specially interacting
with people) should be allowed to make mistakes due to perception noise, or not
to be sure about something (for instance, "I think that..."), which represents
the certainty the robot has about a fact. In the meantime, to overcome
perception errors, the robot provides visual feedback (looking at objects) to
the user when possible and/or verifies the information with the user whenever
is needed (for instance, "do you mean the box\_01"?).

\subsection{Does the cognitive model enable a meaningful interaction with a
real user in a real-world setting?}

Due to the perception problems and computational delays, fully realistic
settings are not yet possible.  The interaction based only on a static setting,
i.e. objects don't move, the interaction with a real user is accurate enough.
We have developed some experiments, where the user interacts with the system
only through a keyboard (no gestures are included) asking questions, providing
information, etc. and the outcome is quite promising. However, for dynamic
settings where the user can point at things, move the objects around, etc., the
interaction is not fast enough and the user should perform slow movements for
the robot to create a consistent state of world.

At recent video showing the current state of the interaction can be viewed
here: \url{http://www.youtube.com/watch?v=IODx50uV_k4}

\subsection{Rework of the introduction} The paper starts off with a discussion
of Austins speech act theory, including locutionary act, illocutionary force
and perlocutionary act. As the concept of speech acts is not really picked up
in the remainder of the paper, I do not see the point in starting with such a
discussion. It would be more appropriate to motivate how a robot could benefit
from a symbolic knowledge base etc.

I very strongly recommend that section I.A. be completely revised.

\section{Comments of the second reviewer}

\begin{itemize}
\item All reviewer comments regarding citations have been addressed, as suggested by the reviewer,
\item The reviewer's remark on human tendency to explain in a consistent way the world has been added
\item Language issues have been addressed as well.
\end{itemize}

\section{Comments of the third reviewer}

\subsection{``Explain the relation of the presented approach to the Open World
Assumption''}

\subsection{``Present a systematic overview of the formalisms and
representations used (OWL, RDF, DL, ...), and choose one consistent syntax for
listing the examples''}

The paragraph 2.2 has been reorganized to clarify the vocabulary and better
define the various acronyms.

\subsection{``The authors are encouraged to explain the notion of curiosity
with respect to the Dora robot by Hawes et al. (2011, ICRA) and Hanheide et al.
(2011, IJCAI)''}

\subsection{``Give examples of the alignment of ORO with OpenCyc''}

\subsection{Typography and notation consistency}

Typography has been throughly revised for better consistency. Notation varients have been
made explicit (for instance, section 2.4).

\subsection{``p.6, II.E. explain "multi-lingual support"''}

\subsection{``Points At: please describe how you manage the critical aspect of
timing''}

PointsAt relation is computed continuously during the interaction by 
selecting the objects that "might" be pointed at by the human. The outcome
of the computation (facts like "human1 pointsAt bottle1") are then
continously updated in the knowledge base.

In the current system, the decision on whether the human is pointing at an 
object in a given time is left to the reasoning on the dialog or on the human gestures:
when the dialogue module analyses a sentence like "give me this", it tries to
resolve the atom "this" by querying the ontology.

Both sitatuation assessement and dialogue processing chains induce 
small delays, but these are roughly the same (below 200ms), and 
compensate for each other.

\subsection{``p.8, III.B.Location according to an object: the authors are
kindly referred to the work done by, e.g., K. Sjöö or J. Kelleher''}

\subsection{``p.8, III.B. "creation of imaginary objects...hinted by verbal
assertions..." -- this phenomenon is called presupposition accommodation''}

\subsection{``p.9, Alg. IV.1: the crucial bits are left out; please explain
"GenerateDescription", "Ontology.Find", "Ontology.CheckEquivalent", and
"Discrimination"''}

\subsection{``how does the system deal with restrictive information vs.
attributive information expressed by adjectives (e.g., in "the yellow banana is
big")?''}

\subsection{``the authors are kindly referred to the work by Zender et al.
(2009, IJCAI) on reference resolution in OWL-DL ontology knowledge bases using
SPARQL queries''}

\subsection{``VI. the authors are advised to present an initial overview of the
robot hardware and the common software architecture used in the systems''}

\subsection{``VI. which university in Munich? LMU or TUM?''}

\subsection{``VI. please present the anecdote of unexpected behavior elsewhere
or give some more context''}

\subsection{``Fig. 15. please make the spelling and the presentation more
consistent''}

\subsection{``the authors are kindly referred to work on the Generation and
Resolution of Referring Expressions (GRE, RRE) from the Computational
Linguistics and Natural Language Processing community, e.g. Dale\&Reiter,
Krahmer\&Theune, Kelleher ''}

\subsection{``please improve the quality of the references (sometimes, the
publisher is missing, sometimes capitalization is omitted, sometimes
abbreviations are used and sometimes full names''}

\end{document}

### 2011-07-15 paper_knowledge.tex
Forgot the main paper. My bad

\documentclass[twocolumn]{svjour3}

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{fancyvrb}
\usepackage[usenames]{color}
\usepackage{url}
\usepackage{hyperref}

\usepackage[numbers]{natbib}
%\usepackage[switch]{lineno}

\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed

\usepackage{eso-pic} %% Background
%\usepackage{floatflt}

\usepackage{enumerate}
\usepackage{paralist} %for inline lists

\usepackage{pseudocode}
\usepackage{alltt}

% fixme notes
\usepackage[draft,footnote,marginclue]{fixme}

\usepackage{subfigure}
%diagrams
\usepackage{tikz}
%\usetikzlibrary{shapes,decorations,trees}
\usetikzlibrary{shapes,trees}

% rotate text in tabular mode
\usepackage{rotating}

%Name of the speaker in a chat
\newcommand{\chatN}[1]{{\footnotesize \textsf{#1}}}
\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}

\newcommand{\stmt}[1]{{\footnotesize $\langle$\stmttt#1\relax$\rangle$}}
\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}

\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}

\newcommand{\ie}{{\textit{i.e.~}}}
\newcommand{\cf}{{\textit{cf~}}}
\newcommand{\eg}{{\textit{e.g.~}}}
% 
% \AddToShipoutPicture{%
%   \AtTextCenter{%
%     \makebox(0,0)[c]{\resizebox{\textwidth}{!}{%
%       \rotatebox{25}{\textsf{\textbf{\textcolor[gray]{0.90}{DRAFT}}}}}}%
%   }%
%  }

%-------------------------------------------------------------------------

\journalname{International Journal of Social Robotics}

\begin{document}
%\linenumbers

\title{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}

\author{
Séverin Lemaignan \and
Raquel Ros \and
E. Akin Sisbot \and
Rachid Alami \and
Michael Beetz
}

\institute{ 
	Séverin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami
	\at
	CNRS - LAAS, 7 avenue du Colonel Roche, F-31077 Toulouse, France\\
	Université de Toulouse, UPS, INSA, INP, ISAE, LAAS, F-31077 Toulouse, France\\
	\email{\{slemaign, rrosespi, sisbot, rachid\}@laas.fr}
	\and
	Séverin Lemaignan \and Michael Beetz
	\at
	Intelligent Autonomous Systems, Technische Universität München, Munich, Germany\\
	\email{\{lemaigna, beetz\}@in.tum.de}
}

\maketitle

\begin{abstract}
This paper presents how extraction, representation and use of symbolic
knowledge from real-world perception and human-robot verbal and non-verbal
interaction can actually enable a grounded and shared model of the world that
is suitable for later high-level tasks like dialogue understanding. We show how
the anchoring process itself relies on the situated nature of human-robot
interactions. We present an implementation, including a specialized symbolic
knowledge representation approach based on Description Logics, and experiments
on several robotic platforms that demonstrate these cognitive capabilities.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Grounding human interaction into the robot knowledge}

\subsection{Situated speech acts}
\label{intro_example}

A messy table, covered with cardboard boxes, books, video tapes... Thomas is
moving and packs everything with the help of Jido, its robot
(Fig.~\ref{fig|vpt}).

`` -- Jido, give me this'', says Thomas, looking at a box that contains a video
tape. The robot smoothly grasps the tape, and hands it to the human.

While this kind of interaction should hopefully sound quite familiar in a
foreseeable future, our robots are not yet quite up to the task. Neither
regarding natural language understanding nor plan-making and manipulation.

To be combined together, those abilities require an unambiguous and shared
representation of concepts (objects, agents, actions...) underlying the
interaction: what are the prerequisites for such a
human sentence --- ``Jido, give me this'' --- to be understood by the robot,
correctly interpreted in the spatial context of the interaction, and ultimately
transformed into an action?

Austin~\cite{Austin1962} would have at first glance analyzed such kind of
sentence as a \emph{speech act}, comprising of \emph{locutionary},
\emph{illocutionary} and possibly \emph{perlocutionary} acts. First, we want to
understand the direct meaning of the sentence (\emph{locutionary act}): we must
acquire the sentence, convert it into a useful syntactic form (quite probably
by mean of speech recognition), and understand the semantics of the sentence,
\ie, What is refered by ``\textit{Jido}''? What is ``\textit{give}''? What is
``\textit{me}''? And ``\textit{this}''?

Working in a situated context, we want furthermore to \emph{resolve} these
semantics atoms, \ie ground them in the sensory-motor space of the robot. For
instance, ``\textit{this}'' is a demonstrative pronoun that refers in this
context to the object the human is focusing on, whatever \textit{focusing}
means: here, Thomas is looking at something, which is a possible cue. But it
could as well point at something or refer to some previously mentioned concept.

\begin{figure}%[!ht] 
	\centering
	\includegraphics[width=0.9\linewidth]{images/pt.jpg} 
	\caption{Interacting with
	the robot in an everyday setup: the human asks for help in vague terms, the
	robot takes into account the human's spatial perspective to refine its
	understanding of the question.} 
	\label{fig|vpt} 
\end{figure}

Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
as thought by the agent must be extracted, and understood. In our example,
Thomas obviously wants an action to be performed by the robot. The action
parametrization is conveyed by the semantics attached to the words and the
grammatical structures of the sentence. In our example, the type of action is
given by the verb ``\textit{give}''. Assuming the robot has some procedural
knowledge attached to this symbol, the action type can be considered as
grounded for the robot. We can as well understand that the recipient of the
action is the human, the performer is the robot itself, and the object acted
upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
that can be extracted from the sentence that allow to fully ground the action.

Ultimately, an implicit speaker's expectation can sometimes be conveyed --- the
\emph{perlocutionary act} --- that must be taken into account. If Thomas would
have said instead `` -- Jido, this box is too heavy for me'', he implicitly
expects the robot to memorize the fact that the box's weight is over his
capabilities, and it may as well imply that Thomas is requiring help to carry
the box.

\subsection{Approach}

Extracting these speech acts and turning them into a content processable by the
robot is a difficult challenge in the general case. We base our approach on
three distinct, inter-related cognitive functions.
Figure~\ref{fig|overall-process} illustrates the general context, while
Figure~\ref{fig|main-components} summarizes the interactions between the three
main components.

\begin{inparaenum}[\itshape 1)]

\item \emph{Physical environment modeling} and \emph{spatial reasoning}
(grouped under the term \emph{situation assessment})  are in charge of building
and maintaining a coherent model of the physical world. This model is realistic
in the sense that it relies on accurate 3D models of both manipulated objects
and humans. It also has dedicated mechanisms to manage disappearing or occluded
objects.  The geometric model is used to compute several spatial properties of
the scene that actually convert the original sensory data into symbolic
beliefs. This includes relative locations of objects, visibility state,
gestures like pointing, etc.  Assuming that other agents are as well
represented in the model, the same computations are applied to analyze the
scene from each agents' point of view (\ie from their \emph{perspectives}).

\item \emph{Knowledge representation and management}: the robot is endowed with
an active knowledge base that provides a logically sound symbolic model of its
beliefs on the world, as well as models for each cognitive agent the robot
interacts with. Each of these models is independent and logically consistent.
This enable reasoning on different perspectives of the world that would be
considered otherwise inconsistent (for instance, an object can be visible for
the robot but not for the human. This object can have at the same time the
property \concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}}, in
two different models).  Our platform also features continuous storage, querying
and event triggering over the pool of facts known by the robot. It relies on
OWL ontologies (a decidable subset of the predicate logics).

Used in combination with the situation assessment framework, the robot is thus
able to maintain different symbolic models of the world, one per agent. This proves an
essential feature (\cite{Roy2005,Kruijff2010}) to enable perspective-aware
grounding of natural language, as we will see in next sections.

\item \emph{Dialogue input processing}, including natural language parsing
capabilities, disambiguation routines and interactive concept anchoring. We
focused our efforts on three classes of utterance, commonly found in
human-robot interaction: \emph{statements} (\ie new facts the human wants to
inform the robot), \emph{orders} (or more generically \emph{desires}) and
\emph{questions on declarative knowledge} (whose answers do not require
explicit planning). This would roughly cover the \emph{representative}
(sometimes referred as \emph{assertives}) and \emph{directives} type of
illocutionary acts, in Searle~\cite{Searle1976} classification.

\end{inparaenum}

Based on these three components, we have developed several grounding mechanisms
for situated verbal interaction, integrated into a broader control architecture
for indoors service robots.

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.8\linewidth]{images/grounding_robot.png}
  \caption{Generic model of cognitive abilities interactions for grounding}
  \label{fig|overall-process}
\end{figure}

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.9\linewidth]{images/grounding.png}
  \caption{The three main components of our architecture and their connections. Submodules are presented in subsequent sections.}
  \label{fig|main-components}
\end{figure}

%~ The system we describe here relies on two distinct activities that run concurrently and continuously: maintenance of a symbolic model from real world interpretation on one hand, and verbal interaction with agents on the other hand.
%~ 
%~ The first activity consists in two main steps:
%~ 
%~ \renewcommand{\labelenumi}{(\alph{enumi})}
%~ \begin{enumerate}
%~ \item The \emph{Object Identification} step is expected to provide consistent identification of objects in the environment. It does not involve any semantic functionality like automatic labelling, classification or annotation capabilities. It is in charge of solving what is merely a perception issue related to the ability to assign some identifier consistently. This is in the general case an open research issue that is outside of the scope of this paper. In our experiments, we will use simplified approaches (like tag-based object identification and motion capture).
%~ 
%~ Besides identification, other perceptual informations are extracted at this step, such as absolute size, color, and more importantly, its spatial position relative to the robot.
%~ 
%~ \item \emph{Physical environment modeling} and \emph{spatial reasoning} are then in charge of rebuilding a coherent physical model of the world based on information extracted in the first step. Once available, the geometric model is used to compute several spatial properties of the scene that actually convert the original sensory data into symbolic beliefs. This includes relative locations of objects, visibility state, gestures like pointing, etc. Assuming that other agents are as well represented in the model, the same computations can be applied to analyze the scene from each agents' point of view (\ie from their \emph{perspectives}).
%~ \end{enumerate}
%~ 
%~ The output of this second step needs to be stored for later reuse: we introduce to that end a general-purpose knowledge management system that relies on a sound logical system (the Description Logics) to store, manipulate and reason about symbolic knowledge. This system is described at length in the next section.
%~ 
%~ The verbal interaction activity relies on three main steps:
%~ 
%~ \begin{enumerate}
%~ \item The \emph{parsing} of natural language aimed at extracting grammatical structures from a sentence,
%~ 
%~ \item The \emph{resolution} of concepts, where words used by the user are associated with the underlying concept they represent, as stored in the robot knowledge base. This step may require an additional interactive disambiguation step if the robot can not successfully bind a word to the concept it represents.
%~ 
%~ \item Finally, the \emph{content analysis} step extracts the \emph{intention} of the sentence and handles it. Orders, questions, statements (new information) are separately processed.
%~ \end{enumerate}

\subsection{Related work}

Our work builds on top of years of work in the artificial intelligence
community around the symbol grounding issue: we can mention
Searle~\cite{Searle1980} who introduces it with the \emph{Chinese room}
metaphor, Harnard~\cite{Harnad1990} who coins the following (slightly
paraphrased in the robotic context) definition of symbol grounding:
\emph{``make the meanings of the meaningless symbol tokens intrinsic to the
robotic system"}, Coradeschi and Saffioti~\cite{Coradeschi2003} who focus on
robotics and define the meaning of \emph{anchoring} as \emph{``the process of
creating and maintaining the correspondence between symbols and sensor data
that refer to the same physical object''}, Ziemke~\cite{Ziemke1999} who
elaborates on the need of embodiment for symbol grounding, and
Sloman~\cite{Sloman2007} who criticizes the \emph{``concept emprism"} that
underlies symbol grounding.

Our contribution relates to two narrower fields: natural language in embodied
interaction contexts and knowledge building and representation in robotic
systems.

Processing natural language in situated context is already an established
research field. In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
challenges to be tackled: cross-modal representation systems, association of
words with perceptual and action categories, modeling of context, figuring out
the right granularity of models, integrating temporal modeling and planning,
the ability to match past (learned) experiences with the current interaction
and the ability to take into account the human perspective. This list offers an
interesting entry point to evaluate our contribution.

Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date survey of literature
on situated human-robot dialogue, focusing on formal representation systems,
bi-directionality of the interaction and context building. They point out as well
that, compared to the cognitive psychology community, the ``situated AI''
community started only recently to take into account agents focus, perspective and temporal
projection abilities.

Dialogue processing on real robots have been explored by several teams.
Brick and Scheutz~\cite{Brick2007} have contributions regarding natural language
processing in an incremental way, and how this enables instant back-channel
feedback (like nodding).

Hüwel et al.~\cite{Huwel2006} propose the concept of \textit{Situated Semantic
Unit}: these meaning atoms are extracted from sentences and expose semantic
links to other units. The parser tries to satisfy these links and rate
accordingly the semantic interpretation of the sentence. Used in conjunction
with ontologies, their approach offers good robustness to ungrammatical or
partial utterances. They validated the approach with an extensive user-study.

While mostly implemented on virtual agents, the GLAIR cognitive architecture
by Shapiro and Bona~\cite{Shapiro2009} is an architecture
explicitly built to tackle the grounding issue from the percept to the
decision. It is a three-layers architecture: a \emph{Knowledge Layer}, a
low-level \emph{Sensori-Actuator Layer} and an intermediate
\emph{Perceptuo-Motor Layer} that binds the previous two.  The knowledge layer
relies on a custom knowledge representation language (more expressive than
first-order logic), and natural language processing capabilities similar to
ours are available. The GLAIR project has been only demonstrated in a limited
set of environments, but exhibits interesting features like explicit management
of contexts of facts and memory models (long term/short term,
episodic/semantic).

Also worth mentioning, Mavridis and Roy~\cite{Mavridis2005} propose the idea of
a \emph{grounded situation model} which is an amodal model of the world where
different sensing modalities, including verbal ones (the robot is able to
\emph{imagine} objects), are merged. Their framework also allows management of
the interaction history (the human can ask for a past event). They propose an
implementation in an environment built on simple entities (a manipulator arm
and color balls).

In the field of symbolic knowledge processing for robots, Gunderson and
Gunderson~\cite{Gunderson2008} introduce the concept of \emph{reification}
(based on both recognition and pre-afference) as an intermediate step between
pattern recognition and symbol grounding. Their underlying storage of knowledge
relies on ontologies and bio-inspired memory models. While sharing similar
foundations to our work, their proposal is based on fairly simple perceptual
modalities and does not develop complex symbolic models that could enable
human-robot interaction.

Suh et al.~\cite{Suh2007} develop {\sc OMRKF}, an ontology-based reasoning
framework for robotics. They try to tackle the grounding issue by storing low-level facts
(like SIFT visual features) in a layered symbolic architecture that works well
in simple sensori-motor spaces. However this approach raises concerns regarding
scalability and management of more complex entities or interactions.

Daoutis et al.~\cite{Daoutis2009} introduce one of the first complete
architecture for grounded human-robot interaction. They successfully bind
low-level percepts (including view-point independent SIFT based object
recognition) to a high-level knowledge representation and reasoning system via
a dedicated anchoring module. They base their knowledge
model directly on the \textit{ResearchCyc} ontology (including the
\textit{MicroTheories} concept), used in combination with the \textsc{CycL}
language. This enables second-order logic modeling and access to a large
common-sense knowledge base.

Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture called \textsc{CRAM}
(Cognitive Robot Abstract Machine) that integrates
\textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
Prolog. Its underlying storage is based on an RDF/OWL\footnote{RDF/OWL is a
standard for Description Logics based ontology
(\url{http://www.w3.org/TR/owl2-primer/}).} ontology, derived from
\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KownRob} have been demonstrated on
several real-world scenarios, where natural language recipes extracted from
Internet had to be translated into plans and executed in a kitchen environment,
perceived and rebuilt on-line by the robots. While Prolog offers more flexible
modeling (no constraints on the arity of predicate, where Description Logics as
used in our work is limited to binary predicates), it is based on the closed
world assumption (if something cannot inferred to be true, it is inferred to be
false) whereas we rely on the open world assumption, which is more realistic in
real world scenarii. A probabilistic extension of \textsc{KnowRob}, called
\textsc{ProbCog}~\cite{Jain2009} is available. While in principle possible, the
CRAM architecture does not provide currently explicit support for the
interaction with humans.

\subsection{Contribution}

Besides proposing a new integration model for sensor data, natural language and
symbolic knowledge repositories, our work extends these previous
contributions by tackling more realistic human-robot interactions: open
speech; complex, dynamic, partially unknown human environments; fully
embodied (with arms, head, ability to move,...) autonomous robots that can
perform manipulation.

Four specific contributions are presented in this work: first, we introduce in
this paper a novel and versatile knowledge base that models in a
formal framework, based on first-order logics, not only the robot's own beliefs
but also every other cognitive agent the robot interacts with.  This explicit
modeling of other agents' belief states is used for the interaction and eases
the implementation of various advanced cognitive behaviours like learning
new concepts or interactive objects discrimination.

Second, we have implemented a framework to extract symbolic facts from complex
real scenes. It is based on a 3D model of the world that the robot builds
on-line by merging sensor modalities. It computes in real-time spatial
relations between perceived objects and it allows to virtually \emph{view} the
same scene from different point of view, enabling \emph{visual} and
\emph{spatial agent perspective taking}.

Third, the same symbolic knowledge base enables richer language capabilities
for the robot.  We propose a new approach to natural language grounding that is
robust, situated and more generic than what can be found in previous work
on situated language grounding. We present several examples that include a
generic recognition and semantic validation of thematic roles (\ie, the
semantic roles played by the different nominal groups around a verb), which
can be translated into input for a symbolic task planner, or disambiguation
based on attention foci.

Lastly, our architecture moves away from standard layered approaches. The way
our different components interact together is not easily represented as a
layered architecture because components interactions are mostly bidirectional.
This is especially visible for the dialogue input processing: this component
does not simply act as an alternative perceptual input to the symbolic
database, but also actively queries previously acquired knowledge to
disambiguate and validate the newly created symbolic knowledge. This behaviour
leads us to the idea of \emph{knowledge-oriented architecture} which is
detailed later on.

These points highlight some original aspects of a larger cognitive
architecture that has been deployed and tested on three distinct robotic
platforms (including both humanoid robot and service robots), demonstrating the
versatility and hardware-agnosticism of these developments.

\subsection{Article overview}

In the next section, we present in depth the so-called {\sc ORO}
(\emph{OpenRobots Ontology} server) knowledge base, a dynamic knowledge
representation system based on Description Logics ontologies and relying on the
\emph{Open World Assumption}. We present it first, along with its objectives,
since it is the knowledge \textit{hub} of the system, used pervasively by other
components.

Section 3 presents the {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge})
module that merges perceptual information with a coherent geometric model and
builds a symbolic interpretation of the world from the robot point of view. It
also computes an individual symbolic model for each agent currently present in
the environment.

Section 4 covers in details the dialogue processing component. The main
algorithms for natural language grounding are detailed. Interactions between
this component and the knowledge base are described.

Section 5 presents three experiments that were conducted on three different
robotic platforms: the \emph{naming} experiment, where a robot anchors new
knowledge in its model through verbal interaction; the \emph{Spy Game} where
either the user or the robot tries to guess which object the other player is
thinking of; and a more complex experiment situated in an everyday setup where
the robot builds models for several agents and interacts with the users using
this knowledge.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ORO, a knowledge management platform}
\label{cognitivekernel}

\subsection{On knowledge representation}
\label{knowledge_representation}

While \emph{knowledge} has no general definition that researchers agree on, for
our own purposes we define knowledge as \emph{an information interpreted in the
cultural and social context of the robot}, where an information is a
\emph{statement} or an \emph{assertion} on the world\footnote{In this paper,
statements are always triples \texttt{<subject, predicate, object>}, following
the Description Logics formalism.}. In practical terms, knowledge is made
of statements that are \emph{contextualized}, if possible \emph{synthesized}
and \emph{limited} to a domain of validity.

These three features have important consequences on the way a knowledge
representation and storage system must be designed. Let us examine them:

\paragraph{Contextualizing} The information contextualizing is the ability for
a cognitive system to connect a fact with a \emph{cultural context}, an
\emph{interpretive scope} and the set of other facts previously acquired by
the agent.
%~ Since machines are limited to syntactic (in contrast to semantic)
%processing, we are mostly looking for a syntactic (\ie, based on symbols)
%matching between concepts representations (in our case, sets of alphanumeric
%characters).\fxfatal{il faut sans doute évoquer ici la relation
%sémantique/syntactique que propose Choamsky}.

We call \textit{cultural context} a broad set of common, general facts that are
considered as widely accepted amongst the interactors (like \textit{bottles may
contain water}). This knowledge is often referred as \emph{common-sense
knowledge}.

By \emph{interpretive scope}, we mean that a concept may have different
interpretation depending on the agent, the current situation or the time frame
the statement belongs to. Since a fact in one scope can be different (or even
inconsistent) with a fact in another scope (for instance, one object can be
visible for the robot and invisible for another agent), the underlying
knowledge representation system must properly handle these interpretive
frameworks.

Note that the focus of the ORO platform is on enabling such context to be
effectively represented rather than actually identifying the current context.
While several approaches to build contextualized knowledge are proposed in this
paper (symbolic environment interpretation, perspective taking, grounded
natural language resolution, self-awareness of its own activity), much remains
to be done for a robot to actually identify its current context as well as
contexts that may be referred to.

\paragraph{Synthesis} The synthesis of knowledge is the identification process
of facts (and their components, concepts and predicates) with each others. For
instance, if the robot observes a human sitting down at a table, and at the
same time, we tell it that ``Peter is sitting at the table'', we would like the
robot to infer that ``Peter'' may be the name of the human. \textit{Synthesis}
refers to the fact that several, \textit{a priori} non correlated, facts must
be associated to the same common concept. This process requires the ability to
control the logical consistency of the knowledge corpus. For instance, if we
add to the previous example the fact that the human that is sitting is a woman,
the synthesis ``Peter is the name of the human'' is not valid anymore.

\paragraph{Domain of validity} The domain of validity of an information,
lastly, covers several aspects: temporal, situational and probabilistic. The
main role of the domain of validity is to specify the scope in which an
information is (believed to be) true. While related to the previously concept
of \emph{interpretive scopes}, the domain of validity addresses the
question whether a fact must be or not considered in a given context. This
limitation of the validity is usually not carried by the fact itself. In the
previous example, for instance, the robot observes a human sitting at a table.
This fact ``a human is sitting at the table'' is true only for a limited
period, until the human stand up. This period of time is not directly
accessible (the robot does not know how long the human plans to stay), but the
knowledge representation system must be able to deal with this uncertainty and
should explicitly label this fact as being limited in time.
%~ ~ To know if a fact is permanent or transitional (as defined by Pollock
%\fxfatal{citation de Langage et Cognition, page 50}) is difficult (especially
%considering that a feature may be considered as permanent or not depending of
%the context: within the situation ``a familly meal''\fxfatal{Check
%translation!}, the fact ``the human is sitting at the table'' could be
%considered as permanent. Conversely, ``ground is static'' is generally
%considered as a static fact, expect if we are talking of planetary mechanics
%for instance. The difficuly lies in the selection of the relevant situation in
%which reasonning must be carried out at a given time) and have currently to be
%defined in the cultural background of the robot.

Those three aspects translate into a set of features that are desirable for a knowledge representation system:
\begin{itemize}
	\item represent raw information,
	\item render a general cultural background, in form of common-sense knowledge,
	\item attach interpretive scopes to new statements,
	\item add and connect new statements to knowledge already present,
	\item store restrictions on the domain of validity of the knowledge.
\end{itemize}

Besides, active processes would be desirable to:
\begin{itemize}
	\item acquire and maintain knowledge perceived from the physical world or
	retrieved from other sources (interaction with other agents, web-based contents,...)
	\item synthesize facts as much as possible,
	\item monitor contexts and manage accordingly the validity of the stored knowledge,
	\item ensure the logical consistency of the knowledge repository, and
	explicit inconsistencies when required\footnote{One may argue that the real world is 
	inherently inconsistent. In this article, we however make a
	\textit{consistent world} assumption, in order to leverage reasoning
	capabilities of the first-order logics. This is supported by the natural
	tendency of humans themselves to provide a consistent explanation of their
	world.}.
\end{itemize}

This list does not cover all the possible features that could be exposed by a
symbolic knowledge management system. Bio-inspired memory management ---the
ability to forget or reinforce knowledge--- or curiosity ---the ability to
identify where knowledge is lacking and actively trigger behaviours to acquire
it---, to give some examples, could arguably be added to the list. However,
this first analysis sets a convenient reference frame to understand and
evaluate knowledge representation systems, including the \textsc{ORO} knowledge
management system we propose.

\begin{figure}[!ht]
\centering
  \includegraphics[width=\columnwidth]{images/snapshot_oroview.png}
  \caption{Live snapshot from the \emph{oro-view} viewer depicting the
  knowledge model of the Jido robot during an experiment. Dots represent
  instances (ABox) while circles pyramids are classes (TBox)}
  \label{fig|oro-view}
\end{figure}

\subsection{ORO Architecture}

The ORO platform~\cite{Lemaignan2010} is primarily designed as a central
knowledge storage service implemented as a server: at run-time, the robot
components can add or query statements. Figure~\ref{fig|oro-overview}
illustrates the overall architecture. The \emph{front-end} accepts and manages
connections from client components. The clients' requests are processed by a
set of internal modules: basic operations on statements, but also higher
cognitive and human-robot interaction related functionalities are available
(detailed thereafter). External plugins can also be easily hooked. The modules
rely on several parallel ontology \emph{back-ends} where the knowledge is
actually stored.

Knowledge is represented in ORO in the Description Logics formalism (using the OWL-DL -- \emph{Web Ontology Language - Description Logics} -- language), as RDF (\emph{Resource Description Framework})
triples (for instance \stmt{robot isIn kitchen}). We use the
Jena\footnote{\url{http://jena.sourceforge.net/}} framework as the underlying
library to load and build in-memory OWL model. We use it in conjunction with
the Pellet\footnote{\url{http://clarkparsia.com/pellet/}} reasoner to ensure
the continuous classification of the OWL concept graph: at run-time, newly
added statements are continuously reasoned about (\emph{classified}), and at
anytime, the server expose a complete set of statements, both the asserted ones
and the inferred ones. For instance, if \stmt{socrates type Human}
and \stmt{Human subClassOf Mortal} are asserted, the server
transparently adds the inferred statement \stmt{socrates type
Mortal}, and a later query retrieving all mortal entities would return
\concept{socrates} as well. The language of the OWL family use the \emph{Open World
Assumption} (if a fact can not be inferred as true, it does not mean that it is
inferred to be false), and the Pellet reasoner honors this assumption as well.

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.8\linewidth]{images/oro_architecture.png}
  \caption{Overview of the ORO architecture.}
  \label{fig|oro-overview}
\end{figure}

\subsection{The OpenRobots Common-Sense Ontology}
\label{ontology}

The first requirement identified in section~\ref{knowledge_representation}
refers to the modeling of a cultural background, a common-sense knowledge
assumed to be shared by all agents.

The ORO server can be loaded with an initial set of statements which we call
the \emph{OpenRobots Common Sense Ontology}. It defines a small set of concepts
(and implicitly, a vocabulary) that can be used by all the modules of the robot
to unambiguously add or query facts. Moreover, the same ontology declares rules
and logical properties that are later used for inference.

The \emph{OpenRobots Common Sense Ontology} defines a small set of classes (56 are
currently defined) and predicates (60 are currently defined) focused on
concepts useful for robot interaction with humans. It includes both very broad
categories like \concept{SpatialThing}, \concept{Event} or \concept{Action}, and much more
concrete concepts as \concept{Table}, \concept{Book} or colors. Predicates allow to
describe the state of the agents and the world with relations like \concept{isOn},
\concept{sees}, \concept{currentlyPerforms}, etc.

Several significant projects are trying to provide such a machine-processable
repository of common sense facts produced by humans (the \textsc{OpenMind}
project\footnote{\url{http://www.openmind.org/}}, for instance). These
knowledge bases are valuable but remain difficult to use in a pervasive way
because of both their incompleteness and the lack of good connections with
underlying, unambiguous concepts.
%\cite{Singh2002}

Our common sense ontology is closely aligned on the open-source
\textsc{OpenCyc}\footnote{\url{http://www.opencyc.org}} upper ontology.
\textsc{OpenCyc} defines a large taxonomy of concepts and semantic
relationships between concepts that are used in several other projects
(\textsc{WordNet, DBpedia}). This potentially eases the exchange and addition
of knowledge from these other sources. Relying on the \textit{de facto}
\textsc{OpenCyc} standard enables as well knowledge exchange with other robots
(for instance, the works previously mentioned by Daoutis and Tenorth rely on
the same \textsc{Cyc} concepts).

Figure~\ref{fig|oro-view} shows a snapshot of the knowledge model during an
experiment. Both concepts from the common sense ontology and
experiment-specific instances are visible.

\subsection{Reasoning and dynamic knowledge structuring}

As previously mentioned, ontologies in ORO are written in OWL, a language
based on Description Logics (a decidable subset of first-order logic). The
Pellet reasoner supports most of the OWL constructs and allows several type of
reasoning:

\begin{itemize}
	\item inheritance
	\item property axioms
		\begin{itemize}
		\item entailments based on predicates' domain and range,
		\item cardinality constraints (including \concept{allValue}, 
		\concept{someValue}, \concept{hasValue}),
		\item property characteristics (symmetry, transitivity)
		\end{itemize}
	\item class restrictions like: \par \footnotesize \concept{Bottle} $\equiv$
		\concept{Artifact} {\bf that} (\concept{hasShape} {\bf value}
		\concept{cylinderShape})\footnote{This example uses the \emph{Manchester
		syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}} \normalsize
	\item set operations like: \par \footnotesize \concept{Color} $\equiv$ {\bf unionOf}(\concept{blue},
		\concept{green}, \concept{orange}, \concept{black}...) \normalsize
	\item generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
		\footnotesize \concept{looksAt(?agt, ?obj)} $\land$
		\concept{pointsAt(?agt,?obj)} $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
		\normalsize 
	\end{itemize}

We provide in ORO accessors to query, add or remove all these properties and
restrictions (however, not the SWRL rules) at run-time. This allows
knowledge introspection and enables the robot to alter its own knowledge
structures (the so-called \emph{T-Box} model) during its life-time by adding
new constraints and properties to classes and predicates.

The \emph{Naming} experiment (section~\ref{naming}) gives a simple example of
such knowledge restructuring.

\subsection{ORO features}
\label{features}

Besides storing and reasoning about knowledge, we have developed in ORO several
features to manage knowledge at higher level: independent cognitive models for
each agent the robot interacts with, categorization capabilities, different
profiles of memory and events registration.

\subsubsection{Base functionalities}
\label{base}
ORO offers an extended set of methods for processing of facts at the triples level. It includes:
\begin{itemize}
	\item statement ({\it i.e.}, RDF triples) insertion, removal, update,
	\item pattern-based removal of statements,
	\item consistency check, insertion of statements with consistency
	constraint (only if the new fact does not lead to inconsistencies),
	\item fast concept lookup, with multi-lingual support,
	\item queries with patterns (for instance, \stmt{ * isOn
	table}, which means ``return me all objects on table'') and filters (for
	instance, \concept{weight < 150.0}),
	\item standard SPARQL\footnote{\url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
\end{itemize}

These basic functionalities enable the incremental construction (and
exploitation) of a consistent knowledge model, and the \emph{Common Sense Ontology}
core set of concepts helps to build assertions that are related to previous one by offering
a predefined vocabulary.

\subsubsection{Representation of alternative cognitive models}
\label{alterite}

As pictured in Figure~\ref{fig|oro-overview}, ORO stores independent cognitive
models for each agent it interacts with. When ORO actually identifies a new
agent (or infers that some instance is an agent), it automatically creates a
new, separate, in-memory OWL model. Robot's components like supervision or
situation assessment may then store the agents' beliefs in a
separate model. All knowledge processing functions available on the robot
primary model are equally available in every agent model. This allows to store
and reason on different (and possibly globally inconsistent) models of the
world.

This feature actually allows us to consider the robot to be endowed with a
\emph{theory of mind}~\cite{Scassellati2002}: the robot can explicitly model
the belief state of its interactors, opening new possibilities for the control
architecture. We present in the section~\ref{dialog} on natural language
grounding how we use this feature to make sense of the human sentences from the
human own point of view.

These multiple models can be viewed as well as different interpretive
scopes, allowing to interpret the same reality from different points of
view.

\subsubsection{Categorization}
\label{categorization}

We have implemented several algorithms (common ancestors, computation of the
best discriminant, see~\cite{Ros2010b}) to help the robot cluster a set of
concepts based on their symbolic similarities (common properties, common
ancestors). One particular application of these functions is discrimination:
when interacting with a user, the robot often needs to clarify an ambiguity
produced by its human partner. For instance, a user may refer to a ``bottle''
whereas two bottles are currently visible: the discrimination routines can
identify possible (symbolic) differences (maybe the color or the size of the
bottles) that permit the robot to ask an accurate question to the user to solve
the ambiguity. This discrimination can occur from the robot perspective or from
a specific agent perspective. Usage of these categorization abilities are
illustrated in the \emph{Spy Game} experiment (section~\ref{spygame}).

\subsubsection{Memory profiles}
\label{memory}

We have designed a simplified bio-inspired memory model that allows us to store
statements in different \emph{memory profiles}. These include \emph{short term
memory} and \emph{long term memory}. Each profile is characterized with a
lifetime, which is assigned to the stored facts. When the lifetime of a fact
expires, ORO automatically removes it.

\subsubsection{The events framework} \label{events} Lastly, ORO allows external
modules to be triggered when specific events occur. For instance, when a
logical sentence becomes true or false, or if a new instance of a certain class
is added. One immediate application is reactive supervision: a component could
for instance subscribe to events of kind \setstmt{?agent isVisible true, ?agent type Human}. As soon as the perception layer detects a human in the
robot's field of view and accordingly updates the knowledge base, the
supervision component would be triggered back.

The event framework also takes advantage of the inference capabilities of ORO.
Thus an event can be indirectly triggered if its triggering conditions can be
inferred to be true.

Next sections present how symbolic knowledge is actually produced and added to
ORO.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Geometric reasoner for situation assessment}

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic assessments. In this section we present SPARK (\emph{SPAtial
Reasoning \& Knowledge}), a situation assessment reasoner that generates
relevant symbolic information from the geometry of the environment with respect
to relations between objects, robots and humans. Moreover, the notion of
\emph{Perspective Taking}~\cite{Flavell1992,Tversky1999} is employed in the
heart of the reasoner to provide the robot with the ability to put itself at
the human’s place and to reason about the world from different perspectives.

As mentioned in the introduction, this paper do not focus on the sensor-level
perception. We rather assume that the perception of the humans and the objects
is provided as a list of unique identifiers with associated 3D meshes and 6DOF
poses.

\subsection{Capabilities}

There are a number of common properties for a robot and a human related to
their capabilities in a given situation: they can both reach, grasp, look,
point, etc. In our context, we group the robot and the human into a single
category. Thus, we define the ``agent'' as an entity that can act in the
environment and manipulate it. In this work we focus on the following
capabilities from each agents' perspective \footnote{Note that each of the
capabilities described are computed from each agent point of view, and
therefore, also stored in different models in ORO for further use at the
decisional level.}.

\begin{itemize}

\item \emph{Sees}: An important ability to know about an agent is to predict
``what it can see'', \ie what is within its field of view (FOV). A robot being
able to compute this information can then act accordingly. A typical example
would be a clarification scenario where the human is searching for an object
and the robot is able to infer that it is looking for the one that is not
visible (otherwise the user would not be searching for it).

In figure~\ref{fig::vis} the field of view of a person is illustrated with a
grey cone (broader one). While he is able to see the two small boxes on the
table in front of the him, the big box on his right is out of this FOV, and
therefore, he is not able to see it.

\item \emph{Looks At}: this relation corresponds to what the agent is focused
on, \ie, where its focus of attention is directed to. This model is based on a
narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
shows the field of attention of a person with a green cone (narrower one). In
this example only the grey box satisfies the \verb"looksAt" relation.

\begin{figure}[!t] \centering
\includegraphics[width=\columnwidth]{images/looks.jpg} \caption{Representation
of the field of view (FOV) and the field of attention (FOA) of the human.}
\label{fig::vis} \end{figure}

\item \emph{Points At}: verifies if an object is pointed at by an agent. This
relation is particularly useful during interaction when one of the agents is
referring to an object saying ``this" or ``that" while pointing at it.

If a big object occludes a smaller one, and an agent is pointing at them, the
outcome of the evaluation will result only in one relation, \ie, \stmt{Agent
\textbf{pointsAt} BigObject} since the small one is not visible to the agent.
On the contrary, if the small object is in front of the big one, then both
objects will be pointed at by the human, generating an ambiguity which should
be solved through higher level reasoning (\eg context analysis or
clarification through verbal interaction as shown later on).

\item \emph{Reachable}: it allows the robot to estimate the agents capability
to reach an object, which is fundamental for task planning. For example, if the
user asks the robot to give her an object, the robot must compute a transfer
point where the user is able to get the object afterward.

Figure~\ref{fig::reach} shows different reachability postures for each object
on the table. In the example, the bottle and the box are both reachable by the
human, but the teddy bear is too far. Instead, from the robot's perspective,
the teddy bear is reachable, while the bottle is not.

\begin{figure}[!t] 
	\centering
	\includegraphics[width=\columnwidth]{images/reach.jpg} 
	\caption{Different reaching postures for the human: the bottle and the 
	box are reachable, while the teddy bear is not.} 
	\label{fig::reach} 
\end{figure}

\end{itemize}

While the first three relations (\emph{sees, looksAt} and \emph{pointsAt}) are
computed through a model based approach, the latter one is based on the
Generalized Inverse Kinematics with pseudo inverse
method~\cite{Nakamura90,Baerlocher04} to find a collision free posture for the
agent where its end-effector is at the center of the object within a given
tolerance. The details of these computations are out of the scope of this
article.

\subsection{Locations} As humans, one way of referring to the object's
positions is based on their symbolic descriptors instead of using their precise
position. In fact, in many cases, this information is the most precise one that
we can give since we do not store the numeric coordinates of objects. The
following relations are computed with respect to the position of the agents and
the objects.

\begin{itemize} \item \emph{Location according to an agent} The predicate
\verb"isLocatedAt" represents spatial locations between agents and objects. For
example we say ``it is on my right, on your left,...'' We compute these spatial
locations by dividing the space around the referent (an agent) into $n$ regions
based on arbitrary angle values relative to the referent orientation. For
example, for $n = 4$ we would have the space divided into \emph{front, left,
right} and \emph{back}. Additionally, two proximity values, \emph{near} and
\emph{far}, may also be considered. The number of regions and proximity values
can be chosen depending on the context where the interaction takes place.

\item \emph{Location according to an object} We can also refer to object
locations with respect to other objects in the environment, such as, above,
next to, in, etc. These types of relations are widely studied in language
grounding (e.g.~\cite{Regier01} presented different models to define the
\emph{above} relation). In this work we use similar models based on the
bounding boxes and center of mass of the objects to define three main relations
(Figure~\ref{fig::sprelation}): \begin{itemize}

\item \textit{isOn}: computes if an object $O_1$ is on another object $O_2$ by
evaluating the center of mass of $O_1$ according to the bounding box of $O_2$.

\item \textit{isIn}: evaluates if an object $O_1$ is inside another object
$O_2$ based on their bounding boxes $BB_{O_1}$ and $BB_{O_2}$.

\item \textit{isNextTo}: indicates whether an object $O_1$ is next to another
object $O_2$. We cannot use a simple distance threshold to determine if two
objects are next to each other since the relation is highly dependent on the
dimensions of the objects. For instance, the maximum distance between large
objects (e.g. two houses) to consider them as being next to each other is much
larger than the maximum distance we would consider for two small objects (\eg
two bottles). Thus, the relation between the dimensions and the distances of
the objects are taken into account.

\begin{figure} 
	\centering
	\includegraphics[width=0.8\columnwidth]{images/spatial_relation.png}
	\caption{Spatial relations between two objects: A) \concept{isOn} relation, 
	B) \concept{isIn} relation, and C) \concept{isNextTo} relation.} 
	\label{fig::sprelation} 
\end{figure}

\end{itemize} 
\end{itemize}

To ensure the different agent's models are up-to-date, all these properties are
always computed on-line, each time the current state of the world changes.

SPARK can be compared to the \emph{Grounded Situation Model} (GSM) by
Mavridis and Roy~\cite{Mavridis2005} in the sense that they both provide an amodal physical
representation of the world used as mediator between the sensor space and
symbolic models. They have however a different set of features: while GSM
enable representation of time and imaginary objects (whose existence is hinted
by verbal assertions from a human), SPARK offers a richer 3D model that enable
computation of several spatial relationships between objects and an effective
implementation of the perspective taking.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The natural language grounding process}
\label{dialog}

Verbal interaction with human presents two categories of challenges: syntactic
ones, and semantic ones. The robot must be able to process and analyze the
structure of human utterances, \ie natural language sentences, and then make
sense of them. As stated in the introduction, we process three categories of
sentences: \emph{statements}, \emph{desires} and \emph{questions} that can be
answered from the declarative knowledge present in the robot knowledge base (a
choice similar to the \emph{Behaviour Cycle} in the GLAIR
architecture~\cite{Shapiro2009}). The grounding of the human discourse consists
for us either in extracting the \emph{informational} content of the sentence
to produce statements or its \emph{intentional} content (\ie, performative value)
to collect orders and questions.

We do not claim any contribution to the field of computational linguists (see
\cite{Kruijff2010} for a survey of formal approaches to natural language
processing in the robotics field): our main contribution here is the grounding
(we call it \emph{resolution}) of concepts involved in the human discourse
through the robot's own knowledge.

We have developed a dedicated module called {\sc Dialogs} that processes human
input in natural language, grounds the concepts in the robot's knowledge and
eventually translates the discourse in a set of queries or declarative OWL/RDF
statements.

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.9\linewidth]{images/dialog_module_simple.png}
  \caption{The {\sc Dialogs} module has three main steps: the parsing,
  the interpretation and the verbalization. The interpretation module is
  responsible for both the \emph{resolution} and the semantic content
  \emph{analysis and translation}.} 
  \label{fig|dialog}
\end{figure}

Figure~\ref{fig|dialog} shows the {\sc Dialogs} module architecture. The user's
input is first pre-processed. For instance, \emph{I'm} constructs are expanded
into \emph{I am} and then parsed. The parser is a custom-made, rule-based (\ie
grammar-free) tool that extracts the grammatical structure from the user's
sentence. Figure~\ref{dialog|parser_output} shows the raw output of the parser for a
moderately complex sentence (``remember that I want you to give me my nice blue
bottle.'').

\begin{figure}[!ht]
\begin{center}
\scriptsize
\begin{alltt}
>> IMPERATIVE
VP: \textbf{remember} (present simple)
    SUBSENTENCE (aim: that)
      NP: \textbf{I}
      VP: \textbf{want} (present simple)
        direct objects: 
          NP: \textbf{you}
        secondary VP: \textbf{give} ()
              direct objects:
                NP: \textbf{bottle}
              indirect objects:
                NP: \textbf{me}
\end{alltt}
\end{center}
\caption{Raw output of the {\sc Dialogs} parser after processing the
sentence: ``remember that I want you to give me my nice blue bottle.''. 
Nominal groups are not grounded yet.} 
\label{dialog|parser_output}
\end{figure}

The result of the parsing is then sent to the \emph{interpretation} module, the
core of the component.  Interpretation consists in three distinct operations:
the sentence \emph{resolution} (concepts grounding), the \emph{content
analysis} (what is the intent of the utterance: information, question or
desire) and the \emph{statement building} (translation into RDF statements).

The sentence resolution has three steps: {\it(1)} pronouns and anaphora are
replaced by, respectively, the correct speaker ID and the ID of the last object
spoken about (extracted from the dialogue history), {\it(2)} nominal groups are
disambiguated and grounded (noun phrase resolution, \cf below), and {\it(3)} 
verbal groups are resolved as well, and their associated
\emph{thematic roles} are retrieved (verbal phrase resolution).

\small
\begin{pseudocode}[ruled]{Resolution}{sentence, currentSpeaker}
\label{algo|Resolution}

\mathcal{G} \GETS \CALL{ParseNominalGroups}{sentence} \\

\FOREACH g \in \mathcal{G} \DO 
\BEGIN
   \mathcal{D} \GETS \CALL{GenerateDescription}{g} \STMTNUM{5.1em}{res.desc}\\
   candidates \GETS \CALL{Ontology.Find}{\mathcal{D}} \STMTNUM{4em}{res.onto}\\

   \IF \left|{candidates}\right| = 0 \THEN
    \BEGIN
      \OUTPUT{\mbox{Couldn't resolve the group!}} \\
      \EXIT \\
    \END
   \ELSEIF \left|{candidates}\right| = 1 \THEN
      id \GETS candidates[0] \STMTNUM{8em}{res.easy}\\

   \ELSE
      \BEGIN
	\IF \CALL{Ontology.CheckEquivalent}{candidates} \THEN
	  id \GETS candidates[0] \\
	\ELSE
	  id \GETS \CALL{Discrimination}{candidates} \\ %\STMTNUM{1em}{st.discrimination}\\
      \END \\
   \CALL{Replace}{g, id, sentence}
\END
\end{pseudocode}
\normalsize

As represented on Figure~\ref{fig|dialog}, interpretation tightly relies on the
communication with the knowledge base. All the concepts the robot manipulates
are stored in the \textit{ontology server} and retrieved through logical
queries, except for the verbs that are currently stored in a dedicated library
(the \emph{action library} on the diagram).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Technical analysis}
\label{examples}

In order to better understand the overall process we next describe the
different steps based on three examples.

In this example, we assume some initial facts are present in the knowledge
base, both in the robot's \emph{own} model and in the \emph{human's} model.
Since the robot tries to ground a human utterance, all queries are sent to the
\emph{human} model, \ie the sentence is interpreted from the human perspective.

\subsection{Informational content extraction}

Figure~\ref{dialog|ex1} shows a first example of human discourse grounding and
the extraction of informational content. We suppose that the robot knowledge
base only contains two initial statements in the human model. The user
asserts a new one: ``The yellow banana is big!''.

\begin{figure}
    \centering
	\begin{tabular}{p{7cm}}
	\emph{Initial knowledge model of} \texttt{human\_01}\\
	\hline
    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\

	\vspace{0.5em}
	\emph{Human input}\\
	\hline
	\hspace{0.3cm}``The yellow banana is big!'' \\

	\vspace{0.5em}
	\emph{Generated partial statements}\\
	\hline
	\hspace{0.3cm}\stmt{?obj type Banana} \\
    	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
    	\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = banana\_01}\\

	\vspace{0.5em}
	\emph{Newly created statements}\\
	\hline
	\hspace{0.3cm}\stmt{banana\_01 hasSize big} \\
	\end{tabular}
\caption{First example of natural language grounding: the nominal group ``the
yellow banana'' is matched with the individual \concept{banana\_01}.}
\label{dialog|ex1}
\end{figure}

We want first to match the nominal group \emph{The yellow banana} to an already
known concept, and second to translate the property \emph{is big} into a
predicate (\emph{hasSize}) to state its semantics. The general resolution
algorithm of nominal groups is described by the algorithm~\ref{algo|Resolution}.

We need to resolve the nominal group \emph{The yellow banana} to a known
concept.  A set of partial statements that describe the
concept is generated based on the grammatical parsing of the sentence
(algorithm~\ref{algo|Resolution}, \emph{(\ref{res.desc})}). In the example, a banana
(\stmt{?obj type Banana}) that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like \concept{hasColor} or \concept{hasSize} that
bind \concept{banana\_01} to adjectives are extracted from a predefined
database of $[Predicate \rightarrow AdjectiveCategory]$, and falls back on the
generic \concept{hasFeature} predicate if the adjective is not known.}.  Based
on these partial statements a query is sent to the ontology server to retrieve
possible instances that match the description (algorithm~\ref{algo|Resolution},
\emph{(\ref{res.onto})}).

In this first simple case, the concept \concept{banana\_01} is unambiguously
matched (since there is only one possible banana) and returned. We can then add
the new information provided by the human, \ie the new statement
\stmt{banana\_01 hasSize big}, to the human model in the ontology
server.

% The translation of \emph{yellow} to \stmt{hasColor yellow} is not obvious: in
% the general case, we associate a adjective to the noun it characterizes with
% the \concept{hasFeature} predicate (for instance, \emph{The sight is beautiful}
% would translate to \stmt{sight hasFeature beautiful}). But we can also manually
% set the predicate associated to a category of adjectives: It is what has been
% done for the main colours. Another example is the size: for known size
% adjectives (big, small, etc.), the \concept{hasSize} predicate is being used.

\subsection{Intentional content through verb resolution}
The sentence in the first example is built with the state verb \emph{be} at
indicative. Let us examine a different example with an action verb at
imperative mode (an order): ``Give me the banana". The process is
described in Figure~\ref{dialog|ex2}.

\begin{figure}
    \centering
	\begin{tabular}{p{7cm}}
	\emph{Initial knowledge model of} \texttt{human\_01}\\
	\hline
    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Human input}\\
	\hline
    	\hspace{0.3cm}``Give me the banana.'' \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Generated partial statements}\\
	\hline
    	\hspace{0.3cm}\stmt{?obj type Banana} \\
	\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = banana\_01}\\

	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Newly created statements}\\
	\hline
    	\hspace{0.3cm}\stmt{human\_01 desires situation\_a3f74} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 type Give} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 performedBy myself} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 actsOnObject banana\_01} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 receivedBy human\_01} \\
	\end{tabular}

\caption{Second example: processing an order.}
\label{dialog|ex2}
\end{figure}

\label{processing_of_actions}

In order to capture the intentional content of a sentence (for example, an
order) we need to retain the semantics of the verb and its complements.
\emph{Thematic roles} allow to semantically link a verb to its complements.  
There is no general agreement amongst linguists on a comprehensive list of 
thematic roles. The amount and the granularity of the roles varies a lot in 
the literature~\cite{Gutierrez2001}. We use a small set of them that matches
the relations the robot can actually achieve (we discuss
possible extensions in the conclusion). For instance, in the second example,
the verb \emph{give} has three thematic roles: \concept{performedBy},
\concept{actsOnObject} and \concept{receivedBy}.

The list of actions the robot can plan for (currently \emph{take},
\emph{place}, \emph{give}, \emph{show}, \emph{hide} and \emph{move}) along with
possible synonyms (for example, \emph{to pick} is set as a synonym of \emph{to
take}) and their associated thematic roles are stored in a predefined library
of actions. For each action we identify and store the role of the subject of
the sentence (always \concept{performedBy}); the role of the direct object
(for instance, \concept{actsOnObject}); and the role of each of the indirect
objects with their optional prepositions (for instance,
\concept{receivedBy})\footnote{Note that in example 2, ``give me the banana'',
the pronoun ``me'' appears before ``banana'', while it is an indirect
complement --- ``give it {\bf to me}''. The parser correctly handles these
cases.}. Moreover, we check with the help of the ontology that each holder of a
role has a consistent semantic. For instance, action \emph{Give} must have a
manipulable physical item (\emph{Artifact}) as direct object. Thus, if the
concept the robot finds for the thematic role \emph{actsOnObject} can not be
inferred to be an artifact, it goes back to the human saying it does not
understand.

This second example  also shows the pronouns reference resolution: ``me'' is
replaced by the id of the current speaker while ``you'' is replaced by
\concept{myself} (\concept{myself} always represent the robot itself). When
present, anaphora (references to previous concepts like ``give me the banana, I
like {\bf it}.'') are also resolved in the same step.

Once the sentence is completely resolved and translated into a formal
representation (a human desire in this case\footnote{Orders are here
represented as human desires: the human desires a specific new situation.}), we
store it in the ontology server. The robot's decisional/executive layers can
then decide whether to execute the order or not.

\subsection{Informational content extraction requiring clarification}
\begin{figure}
    \centering
	\begin{tabular}{p{7cm}}
	\emph{Initial knowledge model of} \texttt{human\_01}\\
	\hline
     	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
     	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow} \\
     	\hspace{0.3cm}\stmt{banana\_02 type Banana} \\
     	\hspace{0.3cm}\stmt{banana\_02 hasColor green} \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Human input}\\
	\hline
     	\hspace{0.3cm}``The banana is good.'' \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Generated partial statements}\\
	\hline
     	\hspace{0.3cm}\stmt{?obj type Banana} \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Robot output speech}\\
	\hline
     	\hspace{0.3cm}``The yellow one or the green one?'' \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Human answer}\\
	\hline
     	\hspace{0.3cm}``The green one.'' \\
	\end{tabular} \\

	\vspace{0.5em}

	\begin{tabular}{p{7cm}}
	\emph{Newly created statements}\\
	\hline
     	\hspace{0.3cm}\stmt{banana\_02 hasFeature good} \\
	\end{tabular}

\caption{Ambiguity resolution: in this example, ``banana'' can refer to the
yellow banana (\concept{banana\_01}) or the green one (\concept{banana\_02}).
Discrimination routines handle the disambiguation process.} \label{dialog|ex3}
\end{figure}

This last example (Figure~\ref{dialog|ex3}) shows the resolution of ambiguous
concepts. In this case the user refers to ``the banana'' while two instances of 
the \concept{Banana} class exist in the ontology. The robot needs to find out 
to which instance the user is actually referring to. To this end, 
disambiguation routines~\cite{Ros2010b} find differences between the instances 
(in the example, one banana is yellow while the other one is green) and build a 
sentence through the \emph{verbalization} module to ask the user a closed 
question that will help clarify the ambiguity: ``Is it yellow or green?'' The 
user's answer is parsed and merged with the previous sentence. The resulting, 
augmented, sentence (``Give me the green banana") goes again through all 
the interpretation steps. This process is repeated until no ambiguities arise. 
In the example, the \concept{banana\_02} is finally returned.

%If no differences \fxfatal{should we say 'in the human model', even if 
%getDiscriminantForAgent currently doesn't work?} can be found, an open question 
%(``give me more information'') is send to the human.

Several other strategies are used in parallel to disambiguate concepts without
having to ask for more information to the human:

\begin{itemize}
	\item Which objects are currently visible to the human? If only one of
	them, then it is probably the one the user is talking about. 
	\item Did a previous interaction involved a specific object that would
	still be the subject of the current sentence?
	\item Is the user looking or pointing to a specific object?
\end{itemize}

%Two cases can alter the way the discrimination routines work:
%\begin{enumerate}
%    \item If a sentence starts with {\it Learn that...}, failures during 
%    discrimination are interpreted as new concepts, and instead of marking the 
%    nominal as not resolved, and new identifier is created and add to the knowledge base.
%    \item For questions like {\it Which color is the bottle?}, the discrimination 
%    algorithm can not use the feature {\it color} to identify to bottle. The 
%    resolution algorithm pass this kind of constraints as a parameter of the 
%    discrimination routines.
%\end{enumerate}

While no examples involving questions have been detailled, \emph{W-} questions
and \emph{yes/no} questions can be processed in a similar way by
\textsc{Dialogs}. For instance, a question like: \emph{What is on the table?}
is grounded (to extract the relation \concept{isOn} and to find what \emph{table}
refers to) and transformed into the following kind of query: {\tt find ?var
[\stmt{?var isOn table1}]}.  Answers are converted back to a full sentence
by the \emph{verbalization} module, and uttered to the human.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}

This section presents several experiments where the different features of our
approach are represented.

\subsection{\emph{Naming} experiment}
\label{naming}

\begin{figure}
\centering

\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/kimp1.jpg}} \hspace{0.2em}
\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/rviz.png}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{(a) Rosie, looking for objects it may know, and (b) viewed in Willow Garage's RViz. The clusters of point are given a unique identifier by the perception that allows the supervision to create the link between the physical objects and their symbolic representation in ORO.}
\label{fig|kimpwatching}
\end{figure}

The \emph{Naming} task uses ORO to anchor perception into the robot's knowledge
through interaction with the user. This task has been implemented on the Rosie
robot at Munich university.

The robot picks an unknown object from the table, shows it to the user, and
asks about its name and type (Figure~\ref{fig|kimpwatching}). The user
interacts with the robot through a dedicated bridge, allowing to chat with
the robot with a standard Jabber instant messaging client.
Figure~\ref{tab|transcript_kimp} shows a chat session with Rosie performing the
naming experiment. The human describes the object until the robot recognizes a
type it already knows: by asking in loop the human for the categories of an
object until it can connect it to a concept it already knows, the robot
accurately anchors perception in its symbolic model and it is able to reason
about it.  The learning process starts over again with each unknown object. To
further exploit the knowledge reasoner in this experiment, we ask the robot
which object do not belong to a typical breakfast table, \ie we ask for the odd
objects. In this case, it corresponds to objects that are neither of type
\concept{Tableware} nor \concept{Food}.

The supervision (\textsc{cram}\footnote{\textsc{cram} (Cognitive Robotic
Abstract Machine) is a RPL-derived framework for rapid development of cognitive
robot control programs we currently develop.}~\cite{Beetz2010}) automatically
updates the ORO server whenever an object enters or leaves the field of view.
%(the perception is based on the \textsc{CoP} framework~\cite{Klank2009}). 
Therefore, the integration of ORO can be seen as an extension to
the robot's belief state that not only contains abstract identifiers
of the internal object representation used in plans, but also the
semantics and roles of objects in the scenario.

\begin{figure}[!ht]
\centering
	\begin{tabular}{r|p{7cm}}
	\chatN{rosie} & Servus! The Bavarian robot is up and fit! \\
	\chatN{rosie} & What is the name of the WHITE CLUSTER on the TABLE? \\
	\chatN{human} & "Rosie's mug" \emph{[Rosie adds \stmt{OBJ001 label "Rosie's mug"} stmt]}\\
	\chatN{rosie} & What kind of object is Rosie's mug? \\
	\chatN{human} & Mug \\
	\chatN{rosie} & What is a Mug? \emph{[Rosie adds \stmt{OBJ001 type Mug} but it does not know the \emph{Mug} concept]}\\
	\chatN{human} & Tableware \emph{[Rosie knows about \emph{Tableware} - belongs to the common-sense ontology. It adds \stmt{Mug subClassOf Tableware}]}\\
	 & (...) \\
	\chatN{rosie} &  What is the name of the BLACK CLUSTER on the TABLE? \\
	\chatN{human} & my\_laptop \\
	\chatN{rosie} & What kind of object is my\_laptop? \\
	\chatN{human} & Computer \\
	\chatN{rosie} & What is a Computer? \\
	\chatN{human} & Artifact \\
	\chatN{rosie} & Do you want me to learn more objects? \\
	\chatN{human} & no \\
	\chatN{rosie} & These objects are odd: my\_laptop \\
	\end{tabular}
	\caption{Transcript of a Jabber session with the robot Rosie.}
	\label{tab|transcript_kimp}
\end{figure}

An unexpected example of what the symbolic reasoning layer brings to more
traditional robotic architectures emerged during the \emph{Naming} experiment:
the perception routines provided segmented blobs corresponding to objects along
with their colors to update the information in ORO. At some point, ORO suddenly
refused to add an object. What seemed at first a communication bug between
modules, was actually the consequence of a consistency check by ORO: because of
bad light conditions, the color recognition was not very reliable, and the same
object was set to have two different colors at the same time. Since the
\concept{hasColor} predicate we had decided to use was a functional predicate
(\ie objects can have only one color), ORO inferred an inconsistency and
discarded the new fact. This kind of logical failure can be used to improve
low-level perception results by ``closing the loop'' with high-level, symbolic
knowledge.

\subsection{``Spy Game'' experiment}

\label{spygame}

This game is based on the traditional children game ``I Spy''. The idea is to
discover the object or concept one of the participants is thinking of by asking
questions such as: ``Is it green? Is it a machine? Is it on your left?'', etc.
When playing, children exploit their knowledge about the world while
categorizing and describing objects through useful discriminants that will
allow them to find out the answer as fast as possible while including
perspective taking abilities~\cite{Moll2006}.

\begin{figure}
\centering

\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/spy-game-real.jpg}} \hspace{0.2em}
\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/spy-game-mhp.jpg}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{Spy game scenario: (a) Real environment and (b) 3D environment model, viewed in \textsc{SPARK}.}
\label{fig|spyGameScenario}
\end{figure}

The scenario for this game (Figure~\ref{fig|spyGameScenario}) consists on a
face-to-face interaction where the human thinks of an object present in the
environment, while the robot queries the human until either discovering the
object or giving up~\cite{Ros2010a}. A categorization example is presented in
Figure~\ref{fig|objectsSpyGame}. The game starts with the human user giving a
first hint (communication is done through a keyboard and screen), allowing the
robot to start the search filtering those objects that fulfill this first
description. Based on this subset, ORO provides a descriptor (or set of
descriptors) that allows a maximum discrimination among objects in the subset.
The robot queries the user about the value of the descriptor (or the most
discriminant among the set of descriptors) and with this new information, the
current subset of objects is filtered again. The process is repeated until
either obtaining a single object that fulfills all the descriptor values, or
failing (\textit{i.e.} no object found).

\begin{figure}[!ht]
\centering
\begin{scriptsize}
\begin{tikzpicture} %[level/.style={sibling distance=30mm/#1},scale=0.8]
	%[edge from parent fork down,
	%every node/.style={fill=black!30,rounded corners},
	%[parent anchor=east,child anchor=west,grow=east,
	%edge from parent/.style={thick,draw}]
	\node {Artifact}
	child {node {Tableware}
		child {node {Bottle}}
		child {node {Cup}}
		}
	child {node {GameObject}}
	child {node {Furniture}
			child {node {Table}}
			child {node {Chair}}
			child {node {Shelf}}};
\end{tikzpicture}
\end{scriptsize}
\caption{Example of object categorization used in the scenario.}
\label{fig|objectsSpyGame}			
\end{figure}

%Thing
%Object
%Tableware
%	- bottle: red, blue, yellow orange
%	- cup: white
%Furniture
%	- table: hrp2table, lowtable
%	- chair: chair1 and chair2
%	- shelf: pink_shelf
%GameObject: spacenavbox, orangebox, accesskit

We have integrated the game in the LAAS-CNRS Jido robot~\cite{Alami1998a}.
Objects are identified through a tag-based vision approach\footnote{ARToolKit:
\url{http://www.hitl.washington.edu/artoolkit/}} and motion capture is used for
human tracking. Their descriptions regarding categories (type of object) and
features (color, shape) are manually given in advance. Spatial relationships
(\concept{front}, \concept{back}, \concept{left}, etc, and \concept{in}, \concept{on} and \concept{next
to}) and visibility (only visible objects for both agents can be considered in
the game) are automatically computed on-line by \textsc{SPARK} (the geometric
reasoner). Figure~\ref{fig|spyGameExample} shows an example of a round game.

\begin{figure}
\centering
	\begin{tabular}{r|p{6cm}}
		\chatN{human} & It is a tableware.\\
		\chatN{jido} & \emph{[retrieves possible objects: coke-bottle, yellow-bottle, orange-bottle, cup-with-handle]} \\
	 			& \emph{[keeps visible objects: coke-bottle, yellow-bottle, cup-with-handle]}\\
				& \emph{[obtains discriminants: type, color.]}\\
				& Which type of object is: bottle or cup? \\
		\chatN{human} & Bottle.\\
		\chatN{jido} & \emph{[obtains possible objects: coke-bottle, yellow-bottle.]}\\
				& \emph{[obtains discriminants: color.]}\\
				& What color the object is: red or yellow?\\
		\chatN{human} & Red.\\
		\chatN{jido} & \emph{[obtains possible objects: coke-bottle.]}\\
				& The object is the Coke bottle!	
	\end{tabular}\\
	\caption{Example of the robot playing Spy game.}
	\label{fig|spyGameExample}
\end{figure}

\subsection{``Moving to London'' scenario}

In order to illustrate some of different reasoning abilities of the robot,
based on the approach presented in this paper, we have designed the following
daily life situation. Tom and Jerry are moving to London, so they are packing
things in boxes. The scenario takes places in the living-room, where Jido (our
robot) is observing while they move things here and there. To assess the
reasoning abilities of the robot they ask Jido for information (entered through
keyboard). Ideally, the robot should also perform actions when required (e.g.
hand an object when asking ``give me...''). However, since it is out of the
scope of this work, we do not include any motion from the robot's side. Similar
to the Spy Game scenario, perception of objects is done through a tag-based
system and humans are detected through motion capture. We next describe in
detail three situations where we can follow the internal robot's reasoning and
the interaction with the users.

\subsubsection{Initial situation assessment}

\begin{figure}[!t]
\centering
  \subfigure[]
  {\includegraphics[width=0.95\linewidth]{images/situationAssessment1.png}\label{a}}
  \subfigure[]
  {\includegraphics[width=0.95\linewidth]{images/situationAssessment2.png}\label{b}}
	\caption{Situation assessment during initialization: a) before placing the
	objects on the table, and b) after placing them. In each image the snapshots
	correspond to: real environment (top-left sub-image); processed image to
	identify the tagged objects (bottom-left sub-image); and the 3D environment
	(right sub-image).}
	\label{fig|sitAssessRoman}
\end{figure}

First Tom enters the room and places on the table some of the things to pack: a
toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.

The initial information in the ontology corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoiTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}

There is a cardboard box on the table:

\begin{center}
    \begin{tabular}{l}
      \stmt{cardBoardBox isOn table}
    \end{tabular}	             
\end{center}

Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a).
He places a toolbox and a video tape on the table, and another video tape
inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b):

\begin{center}
    \begin{tabular}{l}
      \stmt{toolbox isOn table}\\
      \stmt{videoTape1 isOn table}\\
      \stmt{videoTape1 isNextTo toolbox}\\
      \stmt{toolbox isNextTo videoTape1}\\
      \stmt{videoTape2 isIn cardBoardBox}
    \end{tabular}       
\end{center}

\subsubsection{Implicit disambiguation through visual perspective taking}

Tom enters the room again while carrying a big box (Figure~\ref{fig|vpt}). He
approaches the table and asks Jido to handle him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module queries the ontology to
identify the object the human is referring to:

\begin{center} 
\stmt{?obj type VideoTape} 
\end{center}

There are two video tapes in the scene: one on
the table, and another one inside the cardboard box. Thus, the knowledge 
base returns both:

\begin{center}
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [videoTape1, videoTape2]}
\end{center}

However, only one is visible for Tom (the one on the
table). Thus, although there is an ambiguity from the robot's perspective
(since it can see both video tapes), based on the perspective of its human
partner it infers that Tom is referring to the video tape on the table, and not
the one inside the box which is not visible from his view. Therefore,
non-visible objects are removed obtaining:

\begin{center}
{\tt ?obj = [videoTape1]}
\end{center}

Since only one object is available, the robot infers
that the human refers to it and can eventually execute the command, \ie give
it to the human. Alternatively, the robot could first verify with the human if
that was the object being referred to or not before proceeding to execute the
action. Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and
its human partner involved in this situation.

\begin{table}
\begin{center}
\begin{tabular}{l}
\hline
Robot's beliefs about itself (\emph{robot's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible \textit{true}}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible \textit{true}}\\
\hline
\hline
Robot's beliefs about Tom (\emph{Tom's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible \textit{true}}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible \textit{false}}\\
 \hline
\end{tabular}
\end{center}
\caption{Robot's beliefs about itself and its human partner.}
\label{table|ptbeliefs}
\end{table}

\subsubsection{Explicit disambiguation through verbal interaction and gestures}
\begin{figure}[!ht]
  \centering
  \subfigure[]
  {\includegraphics[width=0.9\linewidth]{images/inTheBox1.jpg}\label{a}}
  \subfigure[]
  {\includegraphics[width=0.9\linewidth]{images/inTheBox2.jpg}\label{b}}
\caption{(a) Jerry asks Jido for the content of the box and (b) points at it.}
  \label{fig|pointing}
\end{figure}

Figure~\ref{fig|pointing}a depicts the third situation where Jerry enters the
living room without knowing where Tom had placed the video tapes. So he first
asks Jido: ``What's in the box?''. Before the robot can answer the question it
has to figure out which box Jerry is talking about. Similar to the previous
situation, there are two available boxes:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type Box}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox, toolbox]}
\end{tabular}
\end{center}

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring to: ``Which
box, the toolbox or the cardboard box?'' Jerry could now simply answer the
question. Instead, he decides to point at it while indicating: ``This box''
(Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
cardBoardBox} as being pointed at and looked at by the human and updates the
ontology with this new information using a rule available in the commonsense
ontology ({\tt pointsAt(?ag, ?obj) $\land$ looksAt(?ag, ?obj) $\to$
focusesOn(?ag, ?obj)}) The \textsc{Dialogs} module is then able to merge both
sources of information, verbal (``this'') and gestural to distinguish the box
Jerry refers to.

\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox]}
\end{tabular}
\end{center}

Finally, the \textsc{Dialogs} queries the ontology about the content of the box
and the question can be answered: ``Jido-E''. Note that the object's label is
used instead of its ID. This way we enhance interaction using familiar names
given by the users.

\begin{center}
\begin{tabular}{l}
\stmt{?obj isIn cardBoardBox}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = videoTape2}\\
\end{tabular}
\end{center}

At this point Jerry wants to know where the other tape is, and that is exactly
what he asks Jido: ``And where is the other tape?''. In this occasion, the
\textsc{Dialogs} module is able to interpret that Jerry is not referring to the
video which they were just talking about, but to the other one:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type VideoTape}\\
\stmt{?obj differentFrom videoTape2}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [videoTape1]}
\end{tabular}
\end{center}

Since there is only one possible ``other'' video (there are only two videos in
the scene), it can directly answer Jerry: ``The other tape is on the table and
next to the toolbox.''

\begin{center}
\begin{tabular}{l}
\stmt{videoTape1 isOn table}\\
\stmt{videoTape1 isNextTo toolbox}
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\subsection{Towards an event-driven, knowledge-oriented architecture for personal robotics}

In this paper, we studied knowledge flows between three components: {\it(1)}
{\sc ORO}, an ontology-based knowledge server that stores and maintains
classified RDF statements produced by the other modules in agent-specific
models and allows information to be easily retrieved, either through queries or
via an event system; {\it(2)} {\sc SPARK}, the grounded, human-aware 3D model
of the environment that performs all the spatial reasoning within our
architecture, including reasoning involving motion planing (to compute
reachability of object) and perspective taking, and {\it(3)} {\sc Dialogs}, a
natural language processor that performs simple grammatical parsing of English
language, grounds the semantic content of the utterance (if necessary, also
interacts with the user to disambiguate), and eventually generates a RDF
representation of the sentence.

These components, combined with modules dedicated to symbolic supervision and
task planning (these modules are outside of the scope of this article), compose
an architecture that we call \emph{knowledge-oriented}:

\begin{itemize}
\item{Knowledge is explicitly stored in one central and consistent repository
of facts, accessible by all modules,}
\item{Knowledge is represented in a strict formalism (RDF statements) and
with a clearly defined vocabulary (stated in the {\tt commonsense.oro.owl}
ontology),}
\item{The first two points enable both a loosely-coupled
architecture where modules can very easily be removed or replaced by other ones
as long as they share the same semantics (modules are defined by the knowledge
they produce),}
\item{and a \emph{symbolic} reactive, event-driven approach
to supervision. By managing events at the same level as
the reasoner, we take full advantage of the inference abilities of ORO to
trigger events whose \textit{True} conditions can be inferred.}
\item{Finally, this architecture allows for merging very different knowledge
modalities in a single homogeneous environment, bringing mutual benefits to
components: for instance, the dialogue processing module can perfectly run
without any geometric perception, but its disambiguation routines can
transparently benefit from it when available (since richer symbolic
descriptions of objects are then available).}
\end{itemize}

Regarding the anchoring question, this architecture is
bidirectional: the components we described provide a \textit{bottom-up}
grounding process: SPARK and \textsc{Dialogs} constantly build and push new
symbolic contents about the world to ORO where it becomes accessible to
decisional layers. In parallel, ORO relies on reasoning in a \textit{top-bottom}
way to produce new facts that may trigger in return physical behaviours.

We believe that this \emph{knowledge-oriented} approach has a strong potential
not only to enable rich human-robot interaction, but also as a broader approach
to information alignment and fusion in complex robotic systems.

The versatility of this paradigm could be illustrated by a simple imaginary
scenario with a blind robot and a deaf robot. The blind robot does not see (no
cameras or alike), but someone can verbally describe a scene to it, while the
deaf robot has a good vision system but can not hear anything.  Without any
changes to the software architecture above the language processing module or
the vision module, supervisors of both robot would be able to perform equally
well.

%(to actually implement this imaginary
%situation, the blind robot would of course need \textit{a priori} 3D models
%of objects talked about to enable planning or pick and place actions, and
%the deaf robot would require at least some gesture interpretation to
%understand orders).

We hope as well that our contribution may contribute to fill the gap between
robotics and psychology. Our architecture provides easier-to-reach entry points
to implement some classical psychology tests to robots. We presented
experiments focused on issues related to perspective taking. By explicitly
enabling independent modeling of the beliefs of each agent, our architecture is
especially well suited to set up \emph{False-Belief}
experiments~\cite{Leslie2000}, and we plan to further explore those kinds of
applications.

\subsection{Discussion}

Our system has however shortcomings and opens several questions on different
topics.  In this section, we discuss some of these limitations, possible
extensions, and how this article contributes to the larger debate on symbol
grounding for embodied agent.

\subsubsection{Modeling the real world}

The main challenge we address in this work can be formulated as \emph{How to
model real-world interaction in a symbolic way, processable by the robot to
make decisions}. In the paper we used several times the term \emph{grounding} to
describe the process of binding percepts to symbols (later organized in a
first-order logic framework).  We would like to relate it to
Sloman~\cite{Sloman2007} stance against the \emph{``Symbol Grounding meme''},
where he argues that symbolic grounding is bound to the representation of
somatic concepts (\ie roughly, the sensori-motor relationships that the robot
learns from its interaction with the world) which in turn severely constraints
the domain of concepts accessible to the robot. We could call this type of
grounding \emph{bottom-up} grounding, and Steels~\cite{Steels2007} claims it is
a solved issue.

For us, \emph{Grounding} is on the contrary a \emph{top-bottom} activity: the
robot needs to automatically bind a representation (for instance, a word
uttered by a human, an image taken from a camera, a sentence extracted from
Wikipedia) to an unambiguous, context-dependent, internal concept. This very
concept can (or not) be \textit{a priori} available to the robot as a
pre-loaded ontology (what we previously called the cultural background of the
robot).

While the examples we develop are all based on symbols that have a physical
meaning, the system deals equally well with abstract, \emph{exo-somatic},
concepts like \emph{Time}, \emph{Event} or \emph{Place}. Demonstrating this in
real experiments would be an interesting development.

Amongst the other shortcomings of our architecture, neither the \emph{domain of
validity} nor the context of a fact are represented in a satisfying way (we do
store some kind of context -- the agent mental model for instance). Those
informations are meta-informations on the knowledge, and the ORO framework,
while it allows it through \emph{statement reification}, does not offer yet
a convenient way to store them. One obvious limitation that derives from
the lack of efficient meta-knowledge is the absence of knowledge history.
With ORO, the robot always lives in the present.

Along the same lines, our current framework lacks a proper management of
uncertainty which is essential for real world environments. A probabilistic
layer should be added by attaching truth probabilities to statements, similar
to~\cite{Jain2009}.

% 
% \subsubsection{The challenge of natural language parsing}
% 
% \fxfatal{in Dialogs, no correct/systematic processing of illocutionary acts
% like "learn", "stop" -> that would be the role of the content analysis}
% \fxfatal{mention type of speech act we don't handle}
% 
% \subsubsection{Action recognition and assessment of non-verbal communication}
% 
% \fxfatal{situation assessment: we don't do any posture, gesture recognition, let be action recognition}
%

\subsubsection{On thematic roles and action models}

The current implementation relies on a small, predefined set of action verbs that can
be recognized from natural language (section~\ref{processing_of_actions}).
This constraint does not come from the resolution algorithm itself, but
rather from the difficulty to automatically extract the thematic roles associated
to a verb. 
This could be improved by integrating a symbolic task
planner to the \textsc{Dialogs} module to dynamically provide the list of
actions that the robot can process, \ie actions for which the robot can produce
a plan. Additionally, we could exploit on-line resources like
\textsc{VerbNet}~\cite{Kipper2008}, which provides a large machine-processable
lexicon of English verbs along with their thematic roles.

\subsubsection{Knowledge and embodiment}

The three experiments that were presented in the paper all illustrate how the
robot make use of its embodied nature to establish a meaningful communication
with a human: because the robot and the human share the same physical
environment and they perceive each other, we are able to create a mutual
context.

Sloman, in~\cite{Sloman2009}, argues however that the strong focus on
embodiment in the robotics community has hindered progress towards natural
human-robot interaction. Our approach has hopefully made clear that, similar to
Beetz et al.~\cite{Beetz2010}, we do not consider embodiment \emph{per se}, outside of
a broader symbolic system: our architecture is not bound to the morphology or
the low-level sensori-motor capabilities of a specific agent.

However, we can build a model of the ``human point of view'' because the robot
sees the human, and is able to deduce, at least partially, what the human
perceives or not. This relies on the embodied nature of the interaction. We
infer that a human focuses on some object because it points at it, looks at it,
and besides, the object is visible to him. This relies as well on the embodied
nature of the interaction. And this allow us, in turn, to understand the
meaning of sentences like ``Give me this''.

We hope that this contribution shows that considering embodiment as the most
challenging and fruitful characteristic of robotics in regards to the whole AI
community does not contradict with a formal, highly symbolic approach of the
representation and decision problems that arise in robotics.

Let conclude this article by a short glance back at Roy's list of challenges
for human-robot dialogue: 
\begin{itemize}

	\item While more modalities (especially, deictic gestures and social gazes)
	can be added, we have proposed in this article a \emph{cross-modal
	representation system},

	\item One of the main feature of the \textsc{Dialogs} module is its ability
	to interactively ground concepts throught disambiguation, bringing the
	ability for the robot to \emph{associate words with perceptual and action
	categories},

	\item The ORO knowledge base offers some support for the \emph{modeling of
	context}, but a lot remains to be done in this respect,

	\item \emph{Figuring out the right granularity of models} is partially
	solved by supporting both a geometric reasoning level and a purely symbolic
	level. Generally speaking, it appears that complex robotic systems need
	to operate with a dynamic granularity, depending on the task to achieve,

	\item \emph{Temporal modeling} is currently missing in our architecture,
	and symbolic and geometric \emph{planning} is accomplished outside of the
	knowledge representation loop we presented here. We see planning as an
	essential tool to build predictive knowledge, and we are looking into this
	direction,

	\item Since we provide no time management, our system is currently not able
	to \emph{match past (learned) experiences with the current interaction}.
	This ability is obviously a key step for general action recognition, and
	seems of particular importance for the robot to assess the state of the
	interaction with the human,

	\item Finally, Roy mentions \emph{the ability to take into account the
	human perspective}: this is probably our main contribution which we are now
	trying to develop even futher towards psychology-inspired experiments.

\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Downloads}

The various software components presented in this paper are all open-source.

\begin{itemize}
\item \textsc{ORO} is the stand-alone ontology server. Download and 
documentation are on \url{http://oro.openrobots.org}.
\item The \emph{OpenRobots Common Sense Ontology} can be accessed from
\url{http://kb.openrobots.org}.
\item \textsc{SPARK} is the 3D-based situation assessment module. Its homepage
is \url{http://spark.openrobots.org}.  
\item \textsc{Dialogs} is the natural-language interpretation module. Its
homepage is \url{http://dialogs.openrobots.org}. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{acknowledgements}

Part of this work has been conducted within the EU CHRIS project ({\tt
\url{http://www.chrisfp7.eu/}}) funded by the E.C. Division FP7-IST under
Contract 215805. We also acknowledge support of Agence Nationale de la
Recherche ANR (AMORCES/PSIROB project) and part of this work has been supported
by a Marie Curie Intra European Fellowship. We would like to thank as well
Patrick Tsemengue and Madhi Chouayakh for their great work on \textsc{Dialogs}.

\end{acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{spbasic}
%\bibliography{/home/slemaign/work/biblio/biblio_phd_severin}
\bibliography{/media/Documents/These/work/biblio/biblio_phd_severin}

\end{document}

### 2011-07-15 paper_knowledge.tex
Added bibliography
Integrated Raquel changes on abstract/introduction

@@
 interaction can actually enable a grounded and shared model of the world that
-is suitable for later high-level tasks like dialogue understanding. We show how
+is suitable for later high-level tasks such as dialogue understanding. We show how
 the anchoring process itself relies on the situated nature of human-robot
 interactions. We present an implementation, including a specialized symbolic
-knowledge representation approach based on Description Logics, and experiments
+knowledge representation approach based on Description Logics, and case studies
 on several robotic platforms that demonstrate these cognitive capabilities.
@@
 A messy table, covered with cardboard boxes, books, video tapes... Thomas is
-moving and packs everything with the help of Jido, its robot
+moving out and is packing everything with the help of Jido, his robot
 (Fig.~\ref{fig|vpt}).
 
-`` -- Jido, give me this'', says Thomas, looking at a box that contains a video
-tape. The robot smoothly grasps the tape, and hands it to the human.
+``Jido, give me this'', says Thomas while looking at a box that contains a
+video tape. The robot smoothly grasps the tape, and hands it to the him.
 
@@
 representation of concepts (objects, agents, actions...) underlying the
-interaction: what are the prerequisites for such a
-human sentence --- ``Jido, give me this'' --- to be understood by the robot,
-correctly interpreted in the spatial context of the interaction, and ultimately
-transformed into an action?
+interaction: what are the prerequisites for such a human sentence --``Jido,
+give me this''-- to be understood by the robot, correctly interpreted in the
+spatial context of the interaction, and eventually transformed into an action?
 
@@
 understand the direct meaning of the sentence (\emph{locutionary act}): we must
-acquire the sentence, convert it into a useful syntactic form (quite probably
-by mean of speech recognition), and understand the semantics of the sentence,
-\ie, What is refered by ``\textit{Jido}''? What is ``\textit{give}''? What is
+acquire the sentence, convert it into a useful syntactic form (probably through
+speech recognition), and understand the semantics of the sentence, \ie What is
+referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
 ``\textit{me}''? And ``\textit{this}''?
@@
 instance, ``\textit{this}'' is a demonstrative pronoun that refers in this
-context to the object the human is focusing on, whatever \textit{focusing}
-means: here, Thomas is looking at something, which is a possible cue. But it
-could as well point at something or refer to some previously mentioned concept.
+context to the object the human is focusing on (whatever \textit{focusing}
+means). In this example Thomas is looking at something, which is a possible
+cue. But it could as well be pointing at something or making reference to some
+previously mentioned concept.
 
@@
 Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
-as thought by the agent must be extracted, and understood. In our example,
+as thought by the agent, must be extracted and understood. In our example,
 Thomas obviously wants an action to be performed by the robot. The action
@@
 
-Ultimately, an implicit speaker's expectation can sometimes be conveyed --- the
-\emph{perlocutionary act} --- that must be taken into account. If Thomas would
-have said instead `` -- Jido, this box is too heavy for me'', he implicitly
-expects the robot to memorize the fact that the box's weight is over his
+Finally, the implicit speaker's expectation can sometimes be conveyed --the
+\emph{perlocutionary act}-- and should be taken into account. If Thomas would
+have said instead ``Jido, this box is too heavy for me'', he may implicitly
+expect the robot to memorize the fact that the box's weight is over his
 capabilities, and it may as well imply that Thomas is requiring help to carry
@@
 Figure~\ref{fig|main-components} summarizes the interactions between the three
-main components.
+main components:
 
@@
 \item \emph{Physical environment modeling} and \emph{spatial reasoning}
-(grouped under the term \emph{situation assessment})  are in charge of building
-and maintaining a coherent model of the physical world. This model is realistic
-in the sense that it relies on accurate 3D models of both manipulated objects
-and humans. It also has dedicated mechanisms to manage disappearing or occluded
-objects.  The geometric model is used to compute several spatial properties of
-the scene that actually convert the original sensory data into symbolic
-beliefs. This includes relative locations of objects, visibility state,
-gestures like pointing, etc.  Assuming that other agents are as well
-represented in the model, the same computations are applied to analyze the
-scene from each agents' point of view (\ie from their \emph{perspectives}).
+(grouped under the term \emph{situation assessment}): these components are in
+charge of building and maintaining a coherent model of the physical world. This
+model is realistic in the sense that it relies on accurate 3D models of both
+manipulated objects and humans. It also has dedicated mechanisms to manage
+disappearing or occluded objects.  The geometric model is used to compute
+several spatial properties of the scene that actually convert the original
+sensory data into symbolic beliefs, including relative locations of objects,
+visibility state, gestures (such as pointing), etc.  Assuming that other agents
+are also represented in the model, the same computations are applied to analyze
+the scene from each agent's point of view (\ie from their \emph{perspectives}).
 
@@
 an active knowledge base that provides a logically sound symbolic model of its
-beliefs on the world, as well as models for each cognitive agent the robot
-interacts with. Each of these models is independent and logically consistent.
-This enable reasoning on different perspectives of the world that would be
-considered otherwise inconsistent (for instance, an object can be visible for
-the robot but not for the human. This object can have at the same time the
-property \concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}}, in
-two different models).  Our platform also features continuous storage, querying
-and event triggering over the pool of facts known by the robot. It relies on
-OWL ontologies (a decidable subset of the predicate logics).
+beliefs of the world, as well as models for each cognitive agent the robot
+interacts with. Each of these models is independent and logically consistent,
+enabling reasoning on different perspectives of the world that would otherwise
+be considered inconsistent (for instance, an object can be visible for the
+robot but not for the human. This object can have at the same time the property
+\concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}} using
+two different models).  Our platform relies on OWL ontologies (a decidable
+subset of the predicate logics) featuring continuous storage, querying and
+event triggering over the pool of facts known by the robot.
 
 Used in combination with the situation assessment framework, the robot is thus
-able to maintain different symbolic models of the world, one per agent. This proves an
-essential feature (\cite{Roy2005,Kruijff2010}) to enable perspective-aware
-grounding of natural language, as we will see in next sections.
-
-\item \emph{Dialogue input processing}, including natural language parsing
-capabilities, disambiguation routines and interactive concept anchoring. We
-focused our efforts on three classes of utterance, commonly found in
-human-robot interaction: \emph{statements} (\ie new facts the human wants to
+able to maintain different symbolic models of the world, one per agent. This
+proves an essential feature (\cite{Roy2005,Kruijff2010}) to enable
+perspective-aware grounding of natural language as we will see in next
+sections.
+
+\item \emph{Dialogue input processing}: a subsystem which includes natural
+language parsing capabilities, disambiguation routines and interactive concept
+anchoring. We focused our efforts on three classes of utterances commonly found
+in human-robot interaction: \emph{statements} (\ie new facts the human wants to
 inform the robot), \emph{orders} (or more generically \emph{desires}) and
@@
 explicit planning). This would roughly cover the \emph{representative}
-(sometimes referred as \emph{assertives}) and \emph{directives} type of
-illocutionary acts, in Searle~\cite{Searle1976} classification.
+(sometimes referred as \emph{assertive}) and \emph{directive} illocutionary
+acts in Searle~\cite{Searle1976} classification.
 
@@
 
-Based on these three components, we have developed several grounding mechanisms
-for situated verbal interaction, integrated into a broader control architecture
-for indoors service robots.
+Based on these three components we have developed several grounding mechanisms
+for situated verbal interaction and integrated them into a broader control
+architecture for indoors service robots.
 
@@
 Processing natural language in situated context is already an established
-research field. In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
-challenges to be tackled: cross-modal representation systems, association of
-words with perceptual and action categories, modeling of context, figuring out
-the right granularity of models, integrating temporal modeling and planning,
-the ability to match past (learned) experiences with the current interaction
-and the ability to take into account the human perspective. This list offers an
-interesting entry point to evaluate our contribution.
-
-Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date survey of literature
-on situated human-robot dialogue, focusing on formal representation systems,
-bi-directionality of the interaction and context building. They point out as well
-that, compared to the cognitive psychology community, the ``situated AI''
-community started only recently to take into account agents focus, perspective and temporal
-projection abilities.
-
-Dialogue processing on real robots have been explored by several teams.
-Brick and Scheutz~\cite{Brick2007} have contributions regarding natural language
+research field. In~\cite{Roy2005}, Roy and Reiter summarize what they see as
+the main challenges to be tackled: cross-modal representation systems,
+association of words with perceptual and action categories, modeling of
+context, figuring out the right granularity of models, integrating temporal
+modeling and planning, ability to match past (learned) experiences with the
+current interaction and ability to take into account the human perspective.
+This list offers an interesting entry point to evaluate our contribution.
+
+Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date survey of
+literature on situated human-robot dialogue, focusing on formal representation
+systems, bi-directionality of the interaction and context building. They point
+out as well that compared to the cognitive psychology community, the ``situated
+AI'' community started only recently to take into account agents' focus of
+attention, perspective and temporal projection abilities.
+
+Dialogue processing in real robots have been explored by several teams.  Brick
+and Scheutz~\cite{Brick2007} have contributions regarding natural language
 processing in an incremental way, and how this enables instant back-channel
-feedback (like nodding).
-
-Hüwel et al.~\cite{Huwel2006} propose the concept of \textit{Situated Semantic
-Unit}: these meaning atoms are extracted from sentences and expose semantic
-links to other units. The parser tries to satisfy these links and rate
-accordingly the semantic interpretation of the sentence. Used in conjunction
-with ontologies, their approach offers good robustness to ungrammatical or
-partial utterances. They validated the approach with an extensive user-study.
-
-While mostly implemented on virtual agents, the GLAIR cognitive architecture
-by Shapiro and Bona~\cite{Shapiro2009} is an architecture
-explicitly built to tackle the grounding issue from the percept to the
-decision. It is a three-layers architecture: a \emph{Knowledge Layer}, a
-low-level \emph{Sensori-Actuator Layer} and an intermediate
-\emph{Perceptuo-Motor Layer} that binds the previous two.  The knowledge layer
-relies on a custom knowledge representation language (more expressive than
-first-order logic), and natural language processing capabilities similar to
-ours are available. The GLAIR project has been only demonstrated in a limited
-set of environments, but exhibits interesting features like explicit management
-of contexts of facts and memory models (long term/short term,
-episodic/semantic).
+feedback (like nodding). Hüwel et al.~\cite{Huwel2006} propose the concept of
+\textit{Situated Semantic Unit}: atoms are extracted from sentences exposing
+semantic links to other units. The parser tries to satisfy these links and
+rates the semantic interpretation of the sentence. Used in conjunction with
+ontologies, their approach offers robustness to ungrammatical or partial
+utterances. They validated the approach with an extensive user-study.
+
+While mostly implemented on virtual agents, the GLAIR cognitive architecture by
+Shapiro and Bona~\cite{Shapiro2009} is an architecture explicitly built to
+tackle the grounding issue from the percept to the decision. It is a
+three-layers architecture: a \emph{Knowledge Layer}, a low-level
+\emph{Sensori-Actuator Layer} and an intermediate \emph{Perceptuo-Motor Layer}
+that binds the previous two.  The knowledge layer relies on a custom knowledge
+representation language (more expressive than first-order logic), and natural
+language processing capabilities similar to ours are available. The GLAIR
+project has been only demonstrated in a limited set of environments, but
+exhibits interesting features such as explicit management of contexts of facts
+and memory models (long term/short term, episodic/semantic).
 
 Also worth mentioning, Mavridis and Roy~\cite{Mavridis2005} propose the idea of
-a \emph{grounded situation model} which is an amodal model of the world where
+a \emph{grounded situation model} which is an a-modal model of the world where
 different sensing modalities, including verbal ones (the robot is able to
@@
 Suh et al.~\cite{Suh2007} develop {\sc OMRKF}, an ontology-based reasoning
-framework for robotics. They try to tackle the grounding issue by storing low-level facts
-(like SIFT visual features) in a layered symbolic architecture that works well
-in simple sensori-motor spaces. However this approach raises concerns regarding
-scalability and management of more complex entities or interactions.
+framework for robotics. They tackle the grounding problem by storing low-level
+facts (like SIFT visual features) in a layered symbolic architecture that works
+well in simple sensori-motor spaces. However this approach raises concerns
+regarding scalability and management of more complex entities or interactions.
 
 Daoutis et al.~\cite{Daoutis2009} introduce one of the first complete
-architecture for grounded human-robot interaction. They successfully bind
+architectures for grounded human-robot interaction. They successfully bind
 low-level percepts (including view-point independent SIFT based object
 recognition) to a high-level knowledge representation and reasoning system via
-a dedicated anchoring module. They base their knowledge
-model directly on the \textit{ResearchCyc} ontology (including the
-\textit{MicroTheories} concept), used in combination with the \textsc{CycL}
-language. This enables second-order logic modeling and access to a large
-common-sense knowledge base.
-
-Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture called \textsc{CRAM}
-(Cognitive Robot Abstract Machine) that integrates
+a dedicated anchoring module. They base their knowledge model directly on the
+\textit{ResearchCyc} ontology (including the \textit{MicroTheories} concept),
+used in combination with the \textsc{CycL} language. This enables second-order
+logic modeling and access to a large common-sense knowledge base.
+
+Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture called
+\textsc{CRAM} (Cognitive Robot Abstract Machine) that integrates
 \textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
@@
 modeling (no constraints on the arity of predicate, where Description Logics as
-used in our work is limited to binary predicates), it is based on the closed
-world assumption (if something cannot inferred to be true, it is inferred to be
-false) whereas we rely on the open world assumption, which is more realistic in
-real world scenarii. A probabilistic extension of \textsc{KnowRob}, called
-\textsc{ProbCog}~\cite{Jain2009} is available. While in principle possible, the
-CRAM architecture does not provide currently explicit support for the
-interaction with humans.
-
-\subsection{Contribution}
+used in our work are limited to binary predicates), it is based on the closed
+world assumption (if something cannot be inferred to be true, it is inferred to
+be false) whereas we rely on the open world assumption, which is more realistic
+in real world scenarios. A probabilistic extension of \textsc{KnowRob}, called
+\textsc{ProbCog}~\cite{Jain2009} is also available. While in principle
+possible, currently the CRAM architecture does not provide explicit support for
+interacting with humans.
+
+\subsection{Contributions}
 
 Besides proposing a new integration model for sensor data, natural language and
-symbolic knowledge repositories, our work extends these previous
-contributions by tackling more realistic human-robot interactions: open
-speech; complex, dynamic, partially unknown human environments; fully
-embodied (with arms, head, ability to move,...) autonomous robots that can
-perform manipulation.
-
-Four specific contributions are presented in this work: first, we introduce in
-this paper a novel and versatile knowledge base that models in a
-formal framework, based on first-order logics, not only the robot's own beliefs
-but also every other cognitive agent the robot interacts with.  This explicit
-modeling of other agents' belief states is used for the interaction and eases
-the implementation of various advanced cognitive behaviours like learning
-new concepts or interactive objects discrimination.
+symbolic knowledge repositories, our work extends these previous contributions
+by tackling more realistic human-robot interactions: open speech; complex,
+dynamic, partially unknown human environments; and fully embodied (with arms,
+head, ability to move,...) autonomous robots that can perform manipulation.
+
+Four specific contributions are presented in this work: first, we introduce a
+novel and versatile knowledge base that models in a formal framework, based on
+first-order logics, not only the robot's own beliefs but also every other
+cognitive agent the robot interacts with.  This explicit modeling of other
+agents' belief states is used for the interaction and eases the implementation
+of various advanced cognitive behaviours like learning new concepts or
+interactive objects discrimination.
 
@@
 real scenes. It is based on a 3D model of the world that the robot builds
-on-line by merging sensor modalities. It computes in real-time spatial
-relations between perceived objects and it allows to virtually \emph{view} the
-same scene from different point of view, enabling \emph{visual} and
+on-line by merging different sensor modalities. It computes spatial relations
+between perceived objects in real-time and it allows to virtually \emph{view}
+the same scene from different points of view, enabling \emph{visual} and
 \emph{spatial agent perspective taking}.
@@
 our different components interact together is not easily represented as a
-layered architecture because components interactions are mostly bidirectional.
-This is especially visible for the dialogue input processing: this component
-does not simply act as an alternative perceptual input to the symbolic
-database, but also actively queries previously acquired knowledge to
+layered architecture because interactions among components are mostly
+bidirectional.  This is especially visible for the dialogue input processing:
+this component does not simply act as an alternative perceptual input to the
+symbolic database, but also actively queries previously acquired knowledge to
 disambiguate and validate the newly created symbolic knowledge. This behaviour
-leads us to the idea of \emph{knowledge-oriented architecture} which is
+leads us to the idea of \emph{knowledge-oriented architecture}, which is
 detailed later on.
 
-These points highlight some original aspects of a larger cognitive
-architecture that has been deployed and tested on three distinct robotic
-platforms (including both humanoid robot and service robots), demonstrating the
+These points highlight some original aspects of a larger cognitive architecture
+that has been deployed and tested on three distinct robotic platforms
+(including both humanoid robot and service robots), demonstrating the
 versatility and hardware-agnosticism of these developments.
@@
 
-In the next section, we present in depth the so-called {\sc ORO}
+In the next section we present in depth the so-called {\sc ORO}
 (\emph{OpenRobots Ontology} server) knowledge base, a dynamic knowledge
@@
 
-Section 3 presents the {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge})
+Section 3 presents {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge}), a
 module that merges perceptual information with a coherent geometric model and
-builds a symbolic interpretation of the world from the robot point of view. It
-also computes an individual symbolic model for each agent currently present in
-the environment.
-
-Section 4 covers in details the dialogue processing component. The main
-algorithms for natural language grounding are detailed. Interactions between
-this component and the knowledge base are described.
-
-Section 5 presents three experiments that were conducted on three different
+builds a symbolic interpretation of the world from the robot's point of view,
+as well as an individual symbolic model for each agent currently present in the
+environment.
+
+Section 4 covers in detail the dialogue processing component. The main
+algorithms for natural language grounding are described, as well as the
+interactions between these component and the knowledge base.
+
+Section 5 presents three use-cases that were conducted on three different
 robotic platforms: the \emph{naming} experiment, where a robot anchors new
@@
 either the user or the robot tries to guess which object the other player is
-thinking of; and a more complex experiment situated in an everyday setup where
+thinking of; and a more complex experiment situated in an everyday setup, where
 the robot builds models for several agents and interacts with the users using
@@
 \bibliographystyle{spbasic}
-%\bibliography{/home/slemaign/work/biblio/biblio_phd_severin}
-\bibliography{/media/Documents/These/work/biblio/biblio_phd_severin}
+\bibliography{biblio}
 

### 2011-07-15 paper_knowledge.tex
Integrated Raquel changes on ORO

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ORO, a knowledge management platform}
-\label{cognitivekernel}
-
-\subsection{On knowledge representation}
-\label{knowledge_representation}
+\section{ORO, a knowledge management platform} \label{cognitivekernel}
+
+\subsection{On knowledge representation} \label{knowledge_representation}
 
@@
 statements are always triples \texttt{<subject, predicate, object>}, following
-the Description Logics formalism.}. In practical terms, knowledge is made
-of statements that are \emph{contextualized}, if possible \emph{synthesized}
-and \emph{limited} to a domain of validity.
+the Description Logics formalism.}. In practical terms, knowledge is made of
+statements that are \emph{contextualized}, if possible \emph{synthesized} and
+\emph{limited} to a domain of validity.
 
@@
 
-\paragraph{Contextualizing} The information contextualizing is the ability for
-a cognitive system to connect a fact with a \emph{cultural context}, an
-\emph{interpretive scope} and the set of other facts previously acquired by
-the agent.
+\paragraph{Contextualizing} is the ability for a cognitive system to connect a
+fact with a \emph{cultural context}, an \emph{interpretive scope} and the set
+of other facts previously acquired by the agent.
 %~ Since machines are limited to syntactic (in contrast to semantic)
@@
 We call \textit{cultural context} a broad set of common, general facts that are
-considered as widely accepted amongst the interactors (like \textit{bottles may
-contain water}). This knowledge is often referred as \emph{common-sense
-knowledge}.
-
-By \emph{interpretive scope}, we mean that a concept may have different
-interpretation depending on the agent, the current situation or the time frame
+considered widely accepted among the inter-actors (\eg ``bottles may contain
+water''). This knowledge is often referred as \emph{common-sense knowledge}.
+
+By \emph{interpretive scope} we mean that a concept may have different
+interpretations depending on the agent, the current situation or the time frame
 the statement belongs to. Since a fact in one scope can be different (or even
@@
 
-\paragraph{Synthesis} The synthesis of knowledge is the identification process
-of facts (and their components, concepts and predicates) with each others. For
-instance, if the robot observes a human sitting down at a table, and at the
-same time, we tell it that ``Peter is sitting at the table'', we would like the
-robot to infer that ``Peter'' may be the name of the human. \textit{Synthesis}
-refers to the fact that several, \textit{a priori} non correlated, facts must
-be associated to the same common concept. This process requires the ability to
-control the logical consistency of the knowledge corpus. For instance, if we
-add to the previous example the fact that the human that is sitting is a woman,
-the synthesis ``Peter is the name of the human'' is not valid anymore.
-
-\paragraph{Domain of validity} The domain of validity of an information,
-lastly, covers several aspects: temporal, situational and probabilistic. The
-main role of the domain of validity is to specify the scope in which an
-information is (believed to be) true. While related to the previously concept
-of \emph{interpretive scopes}, the domain of validity addresses the
-question whether a fact must be or not considered in a given context. This
-limitation of the validity is usually not carried by the fact itself. In the
-previous example, for instance, the robot observes a human sitting at a table.
-This fact ``a human is sitting at the table'' is true only for a limited
-period, until the human stand up. This period of time is not directly
-accessible (the robot does not know how long the human plans to stay), but the
-knowledge representation system must be able to deal with this uncertainty and
+\paragraph{Synthesis} corresponds to the identification of facts and their
+components (concepts and predicates) with respect to other facts. For instance,
+if the robot observes a human sitting down at a table, and at the same time, we
+tell it that ``Peter is sitting at the table'', we would like the robot to
+infer that ``Peter'' may be the name of the human. \textit{Synthesis}
+refers to the fact that several, \textit{a priori} non correlated, facts
+must be associated to the same common concept. This process requires the
+ability to control the logical consistency of the knowledge corpus. To
+continue with the previous example,  if we add the fact that the human that
+is sitting is a woman, the synthesis ``Peter is the name of the human'' is
+not valid anymore.
+
+\paragraph{Domain of validity} its main role is to specify the scope in which
+an information is (believed to be) true. It covers several aspects: temporal,
+situational and probabilistic. While related to the previous concept of
+\emph{interpretive scopes}, the domain of validity addresses the question
+whether a fact must be or not considered in a given context. This validity
+limitation is not usually carried by the fact itself. In the previous example,
+for instance, the robot observes a human sitting at a table.  The fact ``a
+human is sitting at the table'' is true only for a limited period of time,
+until the human stands up. This period of time is not directly accessible
+(the robot does not know how long the human plans to stay), but the
+knowledge representation must be able to deal with this uncertainty and
 should explicitly label this fact as being limited in time.
@@
 %considered as a static fact, expect if we are talking of planetary mechanics
-%for instance. The difficuly lies in the selection of the relevant situation in
-%which reasonning must be carried out at a given time) and have currently to be
-%defined in the cultural background of the robot.
-
-Those three aspects translate into a set of features that are desirable for a knowledge representation system:
+%for instance. The difficulty lies in the selection of the relevant situation
+%in which reasoning must be carried out at a given time) and have currently to
+%be defined in the cultural background of the robot.
+
+These three aspects lead us to envisage a knowledge representation system
+characterized by the following features: 
 \begin{itemize}
 	\item represent raw information,
-	\item render a general cultural background, in form of common-sense knowledge,
+	\item render a general cultural background, in the form of common-sense knowledge,
 	\item attach interpretive scopes to new statements,
@@
 
-Besides, active processes would be desirable to:
+Besides, the following active processes would be desirable to:
 \begin{itemize}
@@
 	\item synthesize facts as much as possible,
-	\item monitor contexts and manage accordingly the validity of the stored knowledge,
+	\item monitor contexts and accordingly manage the validity of the stored knowledge,
 	\item ensure the logical consistency of the knowledge repository, and
@@
 ability to forget or reinforce knowledge--- or curiosity ---the ability to
-identify where knowledge is lacking and actively trigger behaviours to acquire
-it---, to give some examples, could arguably be added to the list. However,
-this first analysis sets a convenient reference frame to understand and
-evaluate knowledge representation systems, including the \textsc{ORO} knowledge
-management system we propose.
+identify lacking knowledge and actively trigger behaviors to acquire
+it\fixme{cite Howa}---, to give some examples, could arguably be added to the
+list. However, this first analysis sets a convenient reference frame to
+understand and evaluate knowledge representation systems, including the
+\textsc{ORO} knowledge management system we propose.
 
@@
   \caption{Live snapshot from the \emph{oro-view} viewer depicting the
-  knowledge model of the Jido robot during an experiment. Dots represent
-  instances (ABox) while circles pyramids are classes (TBox)}
+  knowledge model of the robot during an experiment. Dots represent
+  instances (ABox), while circles pyramids are classes (TBox)}
   \label{fig|oro-view}
@@
 The ORO platform~\cite{Lemaignan2010} is primarily designed as a central
-knowledge storage service implemented as a server: at run-time, the robot
-components can add or query statements. Figure~\ref{fig|oro-overview}
+knowledge storage service implemented as a server where the robot
+components can add or query statements at run-time. Figure~\ref{fig|oro-overview}
 illustrates the overall architecture. The \emph{front-end} accepts and manages
@@
 
-Knowledge is represented in ORO in the Description Logics formalism (using the OWL-DL -- \emph{Web Ontology Language - Description Logics} -- language), as RDF (\emph{Resource Description Framework})
-triples (for instance \stmt{robot isIn kitchen}). We use the
-Jena\footnote{\url{http://jena.sourceforge.net/}} framework as the underlying
-library to load and build in-memory OWL model. We use it in conjunction with
-the Pellet\footnote{\url{http://clarkparsia.com/pellet/}} reasoner to ensure
-the continuous classification of the OWL concept graph: at run-time, newly
-added statements are continuously reasoned about (\emph{classified}), and at
-anytime, the server expose a complete set of statements, both the asserted ones
-and the inferred ones. For instance, if \stmt{socrates type Human}
-and \stmt{Human subClassOf Mortal} are asserted, the server
-transparently adds the inferred statement \stmt{socrates type
-Mortal}, and a later query retrieving all mortal entities would return
-\concept{socrates} as well. The language of the OWL family use the \emph{Open World
-Assumption} (if a fact can not be inferred as true, it does not mean that it is
-inferred to be false), and the Pellet reasoner honors this assumption as well.
+Knowledge is represented in ORO in the Description Logics formalism (using the
+OWL-DL -- \emph{Web Ontology Language - Description Logics} -- language), as
+RDF (\emph{Resource Description Framework}) triples (for instance \stmt{robot
+isIn kitchen}). We use the Jena\footnote{\url{http://jena.sourceforge.net/}}
+framework as the underlying library to load and build in-memory OWL model. We
+use it in conjunction with the
+Pellet\footnote{\url{http://clarkparsia.com/pellet/}} reasoner to ensure the
+continuous classification of the OWL concept graph: at run-time, newly added
+statements are continuously reasoned about (\emph{classified}), and at anytime,
+the server expose a complete set of statements, both the asserted ones and the
+inferred ones. For instance, if \stmt{socrates type Human} and \stmt{Human
+subClassOf Mortal} are asserted, the server transparently adds the inferred
+statement \stmt{socrates type Mortal}, and a later query retrieving all mortal
+entities would return \concept{socrates} as well. The language of the OWL
+family use the \emph{Open World Assumption} (if a fact can not be inferred as
+true, it does not mean that it is inferred to be false), and the Pellet
+reasoner honors this assumption as well.
 
@@
 to unambiguously add or query facts. Moreover, the same ontology declares rules
-and logical properties that are later used for inference.
-
-The \emph{OpenRobots Common Sense Ontology} defines a small set of classes (56 are
-currently defined) and predicates (60 are currently defined) focused on
-concepts useful for robot interaction with humans. It includes both very broad
-categories like \concept{SpatialThing}, \concept{Event} or \concept{Action}, and much more
-concrete concepts as \concept{Table}, \concept{Book} or colors. Predicates allow to
-describe the state of the agents and the world with relations like \concept{isOn},
-\concept{sees}, \concept{currentlyPerforms}, etc.
+and logical properties that are later on used for inference.
+
+The \emph{OpenRobots Common Sense Ontology} defines a small set of classes (56
+are currently defined) and predicates (60 are currently defined) focused on
+concepts useful for human-robot interaction. It includes both very broad
+categories like \concept{SpatialThing}, \concept{Event} or \concept{Action},
+and much more concrete concepts as \concept{Table}, \concept{Book} or colors.
+Available predicates allow to describe the state of the agents and the world
+with relations like \concept{isOn}, \concept{sees},
+\concept{currentlyPerforms}, etc.
 
@@
 
-Our common sense ontology is closely aligned on the open-source
+Our common sense ontology is closely aligned with the open-source
 \textsc{OpenCyc}\footnote{\url{http://www.opencyc.org}} upper ontology.
@@
 (\textsc{WordNet, DBpedia}). This potentially eases the exchange and addition
-of knowledge from these other sources. Relying on the \textit{de facto}
-\textsc{OpenCyc} standard enables as well knowledge exchange with other robots
-(for instance, the works previously mentioned by Daoutis and Tenorth rely on
-the same \textsc{Cyc} concepts).
+of knowledge from these other sources. Moreover, it also enables knowledge
+exchange with other robots (for instance, the works previously mentioned by
+Daoutis and Tenorth rely on the same \textsc{Cyc} concepts\fixme{Give
+examples}).
 
@@
 Besides storing and reasoning about knowledge, we have developed in ORO several
-features to manage knowledge at higher level: independent cognitive models for
-each agent the robot interacts with, categorization capabilities, different
-profiles of memory and events registration.
+features to manage knowledge at higher level: 
+%independent cognitive models for
+%each agent the robot interacts with, categorization capabilities, different
+%profiles of memory and events registration.
 
@@
 \label{base}
-ORO offers an extended set of methods for processing of facts at the triples level. It includes:
+
+ORO offers an extended set of methods to process facts at the triples level,
+including:
 \begin{itemize}
-	\item statement ({\it i.e.}, RDF triples) insertion, removal, update,
-	\item pattern-based removal of statements,
+	\item statement (\ie RDF triples) insertion, removal, update,
+	\item pattern-based statements removal,
+	\item pattern-based queries (for instance, \stmt{ * isOn
+	table}, which means ``return me all objects on table'') and filters (for
+	instance, \concept{weight < 150.0}),
 	\item consistency check, insertion of statements with consistency
@@
 	\item fast concept lookup, with multi-lingual support,
-	\item queries with patterns (for instance, \stmt{ * isOn
-	table}, which means ``return me all objects on table'') and filters (for
-	instance, \concept{weight < 150.0}),
-	\item standard SPARQL\footnote{\url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
+	\item standard SPARQL\footnote{{\em SPARQL Protocol and RDF Query Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
 \end{itemize}
 
-These basic functionalities enable the incremental construction (and
-exploitation) of a consistent knowledge model, and the \emph{Common Sense Ontology}
-core set of concepts helps to build assertions that are related to previous one by offering
+While these basic functionalities enable the incremental construction (and
+exploitation) of a consistent knowledge model, the \emph{Common Sense Ontology}
+helps building assertions that are related to previous one by offering
 a predefined vocabulary.
@@
 agent (or infers that some instance is an agent), it automatically creates a
-new, separate, in-memory OWL model. Robot's components like supervision or
-situation assessment may then store the agents' beliefs in a
-separate model. All knowledge processing functions available on the robot
-primary model are equally available in every agent model. This allows to store
-and reason on different (and possibly globally inconsistent) models of the
-world.
+new, separate, in-memory OWL model for that agent. Thus, different robot's
+components, like supervision or situation assessment, may then store the
+agents' beliefs in separate models. All knowledge processing functions in the
+robot's primary model are equally available in every agent's model, which
+allows to store and reason on different (and possibly globally inconsistent)
+models of the world.
 
@@
 \emph{theory of mind}~\cite{Scassellati2002}: the robot can explicitly model
-the belief state of its interactors, opening new possibilities for the control
-architecture. We present in the section~\ref{dialog} on natural language
-grounding how we use this feature to make sense of the human sentences from the
-human own point of view.
+the belief state of its inter-actors, opening new possibilities for the control
+architecture. In the section~\ref{dialog} we present how we use this feature to
+make sense of the user sentences from the his/her point of view. Moreover,
+these multiple models can be viewed as well as different interpretive scopes,
+allowing to interpret the same reality from different points of view.
 
@@
 concepts based on their symbolic similarities (common properties, common
-ancestors). One particular application of these functions is discrimination:
-when interacting with a user, the robot often needs to clarify an ambiguity
-produced by its human partner. For instance, a user may refer to a ``bottle''
-whereas two bottles are currently visible: the discrimination routines can
-identify possible (symbolic) differences (maybe the color or the size of the
-bottles) that permit the robot to ask an accurate question to the user to solve
-the ambiguity. This discrimination can occur from the robot perspective or from
-a specific agent perspective. Usage of these categorization abilities are
-illustrated in the \emph{Spy Game} experiment (section~\ref{spygame}).
+ancestors). One particular application of these functions is discrimination.
+While interacting with a user, the robot quite often needs to clarify an
+ambiguity produced by its human partner. For instance, a user may refer to a
+``bottle'' whereas two bottles are currently visible. Discrimination routines
+can identify possible (symbolic) differences (\eg the color or the size of the
+bottles) that permit the robot to ask an accurate question to the user in order
+to solve the ambiguity. This discrimination can occur from the robot's
+perspective or from a specific agent's perspective. Usage of these
+categorization abilities are illustrated in the \emph{Spy Game}
+(section~\ref{spygame}).
 
@@
 
-\subsubsection{The events framework} \label{events} Lastly, ORO allows external
-modules to be triggered when specific events occur. For instance, when a
-logical sentence becomes true or false, or if a new instance of a certain class
-is added. One immediate application is reactive supervision: a component could
-for instance subscribe to events of kind \setstmt{?agent isVisible true, ?agent type Human}. As soon as the perception layer detects a human in the
-robot's field of view and accordingly updates the knowledge base, the
-supervision component would be triggered back.
-
-The event framework also takes advantage of the inference capabilities of ORO.
-Thus an event can be indirectly triggered if its triggering conditions can be
-inferred to be true.
-
-Next sections present how symbolic knowledge is actually produced and added to
+\subsubsection{The events framework} 
+\label{events}
+
+Lastly, ORO allows external modules to be triggered when specific events occur.
+For instance, when a logical sentence becomes true or false, or if a new
+instance of a certain class is added. One immediate application is reactive
+supervision: a component could for instance subscribe to events of kind
+\setstmt{?agent isVisible true, ?agent type Human}. As soon as the perception
+layer detects a human in the robot's field of view and accordingly updates the
+knowledge base, the supervision component would be triggered back. The event
+framework also takes advantage of the inference capabilities of ORO. Thus an
+event can be indirectly triggered if its triggering conditions can be inferred
+to be true.
+
+Next sections describe how symbolic knowledge is actually produced and added to
 ORO.

### 2011-07-15 paper_knowledge.tex
Integrated Raquel changes on SPARK

@@
 
-Anchoring perceptions in a symbolic model requires perception abilities and
-their symbolic assessments. In this section we present SPARK (\emph{SPAtial
-Reasoning \& Knowledge}), a situation assessment reasoner that generates
-relevant symbolic information from the geometry of the environment with respect
-to relations between objects, robots and humans. Moreover, the notion of
-\emph{Perspective Taking}~\cite{Flavell1992,Tversky1999} is employed in the
-heart of the reasoner to provide the robot with the ability to put itself at
-the human’s place and to reason about the world from different perspectives.
+Anchoring perceptions in a symbolic model requires first perception abilities
+and second, a symbolic assessment of the percepts. In this section we present
+SPARK (\emph{SPAtial Reasoning \& Knowledge}), a situation assessment reasoner
+that generates relevant symbolic information from the geometry of the
+environment with respect to relations between objects and human capabilities.
+Moreover, the notion of \emph{Perspective
+Taking}~\cite{Flavell1992,Tversky1999} is employed in the heart of the
+reasoning to provide the robot with the ability to put itself at the human's
+place and to reason about the world from different perspectives.
 
@@
 their capabilities in a given situation: they can both reach, grasp, look,
-point, etc. In our context, we group the robot and the human into a single
-category. Thus, we define the ``agent'' as an entity that can act in the
-environment and manipulate it. In this work we focus on the following
-capabilities from each agents' perspective \footnote{Note that each of the
-capabilities described are computed from each agent point of view, and
-therefore, also stored in different models in ORO for further use at the
-decisional level.}.
+point, etc. In our context, we group robots and humans into a single category.
+Thus, we define agents as entities that can act in the environment and
+manipulate it. In this work we focus on the following capabilities from each
+agent's perspective\footnote{Note that each of the capabilities described are
+computed from each agent point of view, and therefore, also stored in different
+models in ORO for further use at the decisional level.}.
 
@@
 \item \emph{Looks At}: this relation corresponds to what the agent is focused
-on, \ie, where its focus of attention is directed to. This model is based on a
+on, \ie where its focus of attention is directed to. This model is based on a
 narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
 shows the field of attention of a person with a green cone (narrower one). In
-this example only the grey box satisfies the \verb"looksAt" relation.
+this example only the grey box satisfies the \concept{looksAt} relation.
 
@@
 If a big object occludes a smaller one, and an agent is pointing at them, the
-outcome of the evaluation will result only in one relation, \ie, \stmt{Agent
-\textbf{pointsAt} BigObject} since the small one is not visible to the agent.
-On the contrary, if the small object is in front of the big one, then both
-objects will be pointed at by the human, generating an ambiguity which should
-be solved through higher level reasoning (\eg context analysis or
-clarification through verbal interaction as shown later on).
+outcome of the evaluation will result only in one relation, \ie \stmt{agent\_01
+pointsAt object\_01} since the small one is not visible to the agent.  On the
+contrary, if the small object is in front of the big one, then both objects
+will satisfy the relation, which may generate an ambiguity (which object the
+agent refers to?) that should be solved through higher level reasoning (\eg
+context analysis or clarification through verbal interaction as shown later
+on).
 
@@
 
-While the first three relations (\emph{sees, looksAt} and \emph{pointsAt}) are
-computed through a model based approach, the latter one is based on the
-Generalized Inverse Kinematics with pseudo inverse
+While the first three relations (\concept{sees}, \concept{looksAt} and
+\concept{pointsAt}) are computed through a model based approach, the latter one
+is based on the Generalized Inverse Kinematics with pseudo inverse
 method~\cite{Nakamura90,Baerlocher04} to find a collision free posture for the
@@
 
-\subsection{Locations} As humans, one way of referring to the object's
+\subsection{Locations} As humans, one way of referring to object's
 positions is based on their symbolic descriptors instead of using their precise
@@
 
-\begin{itemize} \item \emph{Location according to an agent} The predicate
-\verb"isLocatedAt" represents spatial locations between agents and objects. For
-example we say ``it is on my right, on your left,...'' We compute these spatial
-locations by dividing the space around the referent (an agent) into $n$ regions
-based on arbitrary angle values relative to the referent orientation. For
-example, for $n = 4$ we would have the space divided into \emph{front, left,
-right} and \emph{back}. Additionally, two proximity values, \emph{near} and
-\emph{far}, may also be considered. The number of regions and proximity values
-can be chosen depending on the context where the interaction takes place.
-
-\item \emph{Location according to an object} We can also refer to object
-locations with respect to other objects in the environment, such as, above,
-next to, in, etc. These types of relations are widely studied in language
-grounding (e.g.~\cite{Regier01} presented different models to define the
+\begin{itemize}
+
+\item \emph{Location according to an agent}: The predicate
+\concept{isLocatedAt} represents spatial locations between agents and objects.
+For example we say ``it is on my right, on your left,...'' We compute these
+spatial locations by dividing the space around the referent (an agent) into $n$
+regions based on arbitrary angle values relative to the referent orientation.
+For example, for $n = 4$ we would have the space divided into \emph{front,
+left, right} and \emph{back}. Additionally, two proximity values, \emph{near}
+and \emph{far}, may also be considered. The number of regions and proximity
+values can be chosen depending on the context where the interaction takes
+place.
+
+\item \emph{Location according to an object}: We can also refer to object
+locations with respect to other objects in the environment, such as \emph{above,
+next to, in}, etc. These types of relations are widely studied in language
+grounding (\eg \cite{Regier01} presented different models to define the
 \emph{above} relation). In this work we use similar models based on the
 bounding boxes and center of mass of the objects to define three main relations
-(Figure~\ref{fig::sprelation}): \begin{itemize}
-
-\item \textit{isOn}: computes if an object $O_1$ is on another object $O_2$ by
+(Figure~\ref{fig::sprelation}):
+
+\begin{itemize}
+	\item \concept{isOn}: computes if an object $O_1$ is on another object $O_2$ by
 	evaluating the center of mass of $O_1$ according to the bounding box of $O_2$.
 
-\item \textit{isIn}: evaluates if an object $O_1$ is inside another object
+	\item \concept{isIn}: evaluates if an object $O_1$ is inside another object
 	$O_2$ based on their bounding boxes $BB_{O_1}$ and $BB_{O_2}$.
 
-\item \textit{isNextTo}: indicates whether an object $O_1$ is next to another
+	\item \concept{isNextTo}: indicates whether an object $O_1$ is next to another
 	object $O_2$. We cannot use a simple distance threshold to determine if two
@@
 	dimensions of the objects. For instance, the maximum distance between large
-objects (e.g. two houses) to consider them as being next to each other is much
+	objects (\eg two houses) to consider them as being next to each other is much
 	larger than the maximum distance we would consider for two small objects (\eg
@@
 
-SPARK can be compared to the \emph{Grounded Situation Model} (GSM) by
-Mavridis and Roy~\cite{Mavridis2005} in the sense that they both provide an amodal physical
-representation of the world used as mediator between the sensor space and
-symbolic models. They have however a different set of features: while GSM
-enable representation of time and imaginary objects (whose existence is hinted
-by verbal assertions from a human), SPARK offers a richer 3D model that enable
-computation of several spatial relationships between objects and an effective
-implementation of the perspective taking.
+SPARK can be compared to the \emph{Grounded Situation Model} (GSM) introduced
+by Mavridis and Roy~\cite{Mavridis2005} in the sense that they both provide an
+amodal physical representation of the world used as a mediator between the
+sensor space and symbolic models. They have however a different set of
+features: while GSM enables representation of time and imaginary objects (whose
+existence is hinted by verbal assertions from a human), SPARK offers a richer
+3D model that enables the computation of several spatial relationships between
+objects and an effective implementation of perspective taking capabilities.
 

### 2011-07-15 paper_knowledge.tex
Integrated Raquel changes on Dialogs + conclusion

@@
 choice similar to the \emph{Behaviour Cycle} in the GLAIR
-architecture~\cite{Shapiro2009}). The grounding of the human discourse consists
-for us either in extracting the \emph{informational} content of the sentence
-to produce statements or its \emph{intentional} content (\ie, performative value)
-to collect orders and questions.
-
-We do not claim any contribution to the field of computational linguists (see
-\cite{Kruijff2010} for a survey of formal approaches to natural language
-processing in the robotics field): our main contribution here is the grounding
-(we call it \emph{resolution}) of concepts involved in the human discourse
-through the robot's own knowledge.
+architecture~\cite{Shapiro2009}). In our work, the grounding process of the
+human discourse consists in extracting either the \emph{informational} content
+of the sentence to produce statements or its \emph{intentional} content (\ie
+performative value) to collect orders and questions. We do not claim any
+contribution to the field of computational linguists (see \cite{Kruijff2010}
+for a survey of formal approaches to natural language processing in the
+robotics field). Our main contribution here is the grounding (we call it
+\emph{resolution}) of concepts involved in the human discourse through the
+robot's own knowledge.
 
@@
 sentence. Figure~\ref{dialog|parser_output} shows the raw output of the parser for a
-moderately complex sentence (``remember that I want you to give me my nice blue
-bottle.'').
+moderately complex sentence.
 
@@
 \caption{Raw output of the {\sc Dialogs} parser after processing the
-sentence: ``remember that I want you to give me my nice blue bottle.''. 
+sentence: ``remember that I want you to give me my nice blue bottle.'' 
 Nominal groups are not grounded yet.} 
@@
 
-The result of the parsing is then sent to the \emph{interpretation} module, the
+The output of the parser is then sent to the \emph{interpretation} module, the
 core of the component.  Interpretation consists in three distinct operations:
@@
 The sentence resolution has three steps: {\it(1)} pronouns and anaphora are
-replaced by, respectively, the correct speaker ID and the ID of the last object
-spoken about (extracted from the dialogue history), {\it(2)} nominal groups are
+replaced by the correct speaker ID and the ID of the last object spoken about
+(extracted from the dialogue history) respectively, {\it(2)} nominal groups are
 disambiguated and grounded (noun phrase resolution, \cf below), and {\it(3)}
-verbal groups are resolved as well, and their associated
-\emph{thematic roles} are retrieved (verbal phrase resolution).
+verbal groups are resolved and their associated \emph{thematic roles} are
+retrieved (verbal phrase resolution).
 
@@
 	\ELSE
-	  id \GETS \CALL{Discrimination}{candidates} \\ %\STMTNUM{1em}{st.discrimination}\\
+	  id \GETS \CALL{Discrimination}{candidates} \\ %\STMTNUM{0em}{st.discrimination}\\
       \END \\
@@
 
-As represented on Figure~\ref{fig|dialog}, interpretation tightly relies on the
+As represented in Figure~\ref{fig|dialog}, interpretation tightly relies on the
 communication with the knowledge base. All the concepts the robot manipulates
@@
 queries, except for the verbs that are currently stored in a dedicated library
-(the \emph{action library} on the diagram).
+(the \emph{action library} in the diagram).
 
@@
 In order to better understand the overall process we next describe the
-different steps based on three examples.
-
-In this example, we assume some initial facts are present in the knowledge
-base, both in the robot's \emph{own} model and in the \emph{human's} model.
-Since the robot tries to ground a human utterance, all queries are sent to the
-\emph{human} model, \ie the sentence is interpreted from the human perspective.
+different steps of the approach based on three examples. In these examples, we
+assume that some initial facts are present in the knowledge base, both in the
+robot's own model and in the human's model.  Since the robot tries to ground a
+human utterance, all queries are sent to the human model in order to interpret
+it from the human perspective.
 
@@
 Figure~\ref{dialog|ex1} shows a first example of human discourse grounding and
-the extraction of informational content. We suppose that the robot knowledge
+the extraction of informational content. We assume that the robot knowledge
 base only contains two initial statements in the human model. The user
@@
 \caption{First example of natural language grounding: the nominal group ``the
-yellow banana'' is matched with the individual \concept{banana\_01}.}
+yellow banana'' is matched with the individual \concept{banana\_01}.
+``$\Rightarrow$'' represents the output of the ontology server.}
 \label{dialog|ex1}
@@
 
-We want first to match the nominal group \emph{The yellow banana} to an already
+We first want to match the nominal group \emph{The yellow banana} to an already
 known concept, and second to translate the property \emph{is big} into a
 predicate (\emph{hasSize}) to state its semantics. The general resolution
-algorithm of nominal groups is described by the algorithm~\ref{algo|Resolution}.
-
-We need to resolve the nominal group \emph{The yellow banana} to a known
-concept.  A set of partial statements that describe the
-concept is generated based on the grammatical parsing of the sentence
-(algorithm~\ref{algo|Resolution}, \emph{(\ref{res.desc})}). In the example, a banana
-(\stmt{?obj type Banana}) that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like \concept{hasColor} or \concept{hasSize} that
-bind \concept{banana\_01} to adjectives are extracted from a predefined
-database of $[Predicate \rightarrow AdjectiveCategory]$, and falls back on the
-generic \concept{hasFeature} predicate if the adjective is not known.}.  Based
-on these partial statements a query is sent to the ontology server to retrieve
-possible instances that match the description (algorithm~\ref{algo|Resolution},
-\emph{(\ref{res.onto})}).
+algorithm of nominal groups is described in algorithm~\ref{algo|Resolution}.
+
+To resolve the nominal group \emph{The yellow banana} a set of partial
+statements that describe the concept is generated based on the grammatical
+parsing of the sentence (algorithm~\ref{algo|Resolution},
+\emph{(\ref{res.desc})}). In the example, a banana (\stmt{?obj type Banana})
+that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
+\concept{hasColor} or \concept{hasSize} that bind \concept{banana\_01} to
+adjectives are extracted from a predefined database of $[Predicate \rightarrow
+AdjectiveCategory]$, and falls back on the generic \concept{hasFeature}
+predicate if the adjective is not known.}.  Based on these partial statements a
+query is sent to the ontology server to retrieve possible instances that match
+the description (algorithm~\ref{algo|Resolution}, \emph{(\ref{res.onto})}).
 
 In this first simple case, the concept \concept{banana\_01} is unambiguously
-matched (since there is only one possible banana) and returned. We can then add
+matched (since there is only one possible banana) and returned. Finally, we can now add
 the new information provided by the human, \ie the new statement
@@
 There is no general agreement amongst linguists on a comprehensive list of 
-thematic roles. The amount and the granularity of the roles varies a lot in 
-the literature~\cite{Gutierrez2001}. We use a small set of them that matches
+thematic roles. The amount and the granularity of roles varies a lot in 
+the literature~\cite{Gutierrez2001}. We thus use a small set of them, which matches
 the relations the robot can actually achieve (we discuss
@@
 take}) and their associated thematic roles are stored in a predefined library
-of actions. For each action we identify and store the role of the subject of
+of actions. For each action we identify and store: the role of the subject in
 the sentence (always \concept{performedBy}); the role of the direct object
@@
 complement --- ``give it {\bf to me}''. The parser correctly handles these
-cases.}. Moreover, we check with the help of the ontology that each holder of a
-role has a consistent semantic. For instance, action \emph{Give} must have a
-manipulable physical item (\emph{Artifact}) as direct object. Thus, if the
-concept the robot finds for the thematic role \emph{actsOnObject} can not be
-inferred to be an artifact, it goes back to the human saying it does not
+cases.}. Moreover, through the ontology we check that each holder of a
+role is semantically consistent. For instance, the action \emph{Give} must have a
+manipulable physical item (\concept{Artifact}) as direct object. Thus, if the
+concept the robot finds for the thematic role \concept{actsOnObject} can not be
+inferred to be an artifact, the robot goes back to the human saying it does not
 understand.
@@
 replaced by the id of the current speaker while ``you'' is replaced by
-\concept{myself} (\concept{myself} always represent the robot itself). When
-present, anaphora (references to previous concepts like ``give me the banana, I
+\concept{myself} (\concept{myself} always represents the robot itself). When
+present, anaphoras (references to previous concepts like ``give me the banana, I
 like {\bf it}.'') are also resolved in the same step.
@@
 Once the sentence is completely resolved and translated into a formal
-representation (a human desire in this case\footnote{Orders are here
+representation (a human desire in this example\footnote{Orders are here
 represented as human desires: the human desires a specific new situation.}), we
@@
      	\hspace{0.3cm}\stmt{?obj type Banana} \\
+	\hspace{0.7cm} $\Rightarrow$ \concept{?obj = [banana\_01, banana\_02]}
+	\end{tabular} \\
+
+	\vspace{0.5em}
+
+	\begin{tabular}{p{7cm}}
+	\emph{Discrimination process}\\
+	\hline
+     	\hspace{0.3cm}\concept{discriminate([banana\_01, banana\_02])} \\
+	\hspace{0.7cm} $\Rightarrow$ \concept{?hasColor = [yellow, green]}
 	\end{tabular} \\
@@
 
+	\begin{tabular}{p{7cm}}
+	\emph{Extended human input}\\
+	\hline
+     	\hspace{0.3cm}``The green banana is good.'' \\
+	\end{tabular} \\
+
+	\vspace{0.5em}
+
+	\begin{tabular}{p{7cm}}
+	\emph{Generated partial statements}\\
+	\hline
+     	\hspace{0.3cm}\stmt{?obj type Banana} \\
+     	\hspace{0.3cm}\stmt{?obj hasColor green} \\
+	\hspace{0.7cm} $\Rightarrow$ \concept{?obj = [banana\_02]}
+	\end{tabular} \\
+
+	\vspace{0.5em}
 	\begin{tabular}{p{7cm}}
@@
 user's answer is parsed and merged with the previous sentence. The resulting, 
-augmented, sentence (``Give me the green banana") goes again through all 
+augmented, sentence (``The green banana is good") goes again through all 
 the interpretation steps. This process is repeated until no ambiguities arise. 
@@
 	still be the subject of the current sentence?
-	\item Is the user looking or pointing to a specific object?
+	\item Is the user looking or pointing at a specific object?
 \end{itemize}
@@
 
-While no examples involving questions have been detailled, \emph{W-} questions
+While no examples involving questions have been detailed, \emph{W-} questions
 and \emph{yes/no} questions can be processed in a similar way by
-\textsc{Dialogs}. For instance, a question like: \emph{What is on the table?}
+\textsc{Dialogs}. For instance, a question like ``What is on the table?''
 is grounded (to extract the relation \concept{isOn} and to find what \emph{table}
@@
 \hspace{1.8cm} (a) \hspace{3.9cm} (b)
-\caption{(a) Rosie, looking for objects it may know, and (b) viewed in Willow Garage's RViz.
-The clusters of point are given a unique identifier by the perception that allows the supervision to create the link between the physical objects and their symbolic representation in ORO.}
+\caption{(a) Rosie, looking for objects it may know, and (b) viewed in Willow
+Garage's RViz. The clusters of point are given a unique identifier by the
+perception that allows the supervision to create the link between the physical
+objects and their symbolic representation in ORO.} 
 \label{fig|kimpwatching}
@@
 	\chatN{rosie} & What is the name of the WHITE CLUSTER on the TABLE? \\
-	\chatN{human} & "Rosie's mug" \emph{[Rosie adds \stmt{OBJ001 label "Rosie's mug"} stmt]}\\
-	\chatN{rosie} & What kind of object is Rosie's mug? \\
+	\chatN{human} & ``Rosie's mug'' \\
+	\chatN{rosie} &\emph{[add \stmt{obj\_01 label "Rosie's mug"}]}\\
+				& What kind of object is ``Rosie's mug''? \\
 	\chatN{human} & Mug \\
-	\chatN{rosie} & What is a Mug? \emph{[Rosie adds \stmt{OBJ001 type Mug} but it does not know the \emph{Mug} concept]}\\
-	\chatN{human} & Tableware \emph{[Rosie knows about \emph{Tableware} - belongs to the common-sense ontology. It adds \stmt{Mug subClassOf Tableware}]}\\
+	\chatN{rosie} & \emph{[add \stmt{obj\_01 type Mug}]}\\
+			& What is a Mug?\\ 
+	\chatN{human} & Tableware\\
+	\chatN{rosie} & \emph{[adds \stmt{Mug subClassOf Tableware}. Rosie knows about \emph{Tableware} --it belongs to the common-sense ontology--.}\\
 	 & (...) \\
 	\chatN{rosie} &  What is the name of the BLACK CLUSTER on the TABLE? \\
-	\chatN{human} & my\_laptop \\
-	\chatN{rosie} & What kind of object is my\_laptop? \\
+	\chatN{human} & ``my\_laptop'' \\
+	\chatN{rosie} & What kind of object is ``my\_laptop''? \\
 	\chatN{human} & Computer \\
@@
 	\chatN{human} & no \\
-	\chatN{rosie} & These objects are odd: my\_laptop \\
+	\chatN{rosie} & These objects are odd: ``my\_laptop'' \\
 	\end{tabular}
@@
 either obtaining a single object that fulfills all the descriptor values, or
-failing (\textit{i.e.} no object found).
+failing (\ie no object found).
 
@@
 features (color, shape) are manually given in advance. Spatial relationships
-(\concept{front}, \concept{back}, \concept{left}, etc, and \concept{in}, \concept{on} and \concept{next
-to}) and visibility (only visible objects for both agents can be considered in
-the game) are automatically computed on-line by \textsc{SPARK} (the geometric
-reasoner). Figure~\ref{fig|spyGameExample} shows an example of a round game.
+(\concept{front}, \concept{back}, \concept{left}, etc, and \concept{in},
+\concept{on} and \concept{next to}) and visibility (only visible objects for
+both agents can be considered in the game) are automatically computed on-line
+by \textsc{SPARK} (the geometric reasoner). Figure~\ref{fig|spyGameExample}
+shows an example of a round game.
 
@@
 reasoning abilities of the robot they ask Jido for information (entered through
-keyboard). Ideally, the robot should also perform actions when required (e.g.
+keyboard). Ideally, the robot should also perform actions when required (\eg
 hand an object when asking ``give me...''). However, since it is out of the
@@
 	\item One of the main feature of the \textsc{Dialogs} module is its ability
-	to interactively ground concepts throught disambiguation, bringing the
+	to interactively ground concepts through disambiguation, bringing the
 	ability for the robot to \emph{associate words with perceptual and action
@@
 	human perspective}: this is probably our main contribution which we are now
-	trying to develop even futher towards psychology-inspired experiments.
+	trying to develop even further towards psychology-inspired experiments.
 

### 2011-07-15 cover_letter.tex
Moved paragraph on perlocutionnary acts to a footnote

@@
 
-\subsection{Rework of the introduction} The paper starts off with a discussion
+\subsection{Rework of the introduction}
+
+Reviewer comment: ``The paper starts off with a discussion
 of Austins speech act theory, including locutionary act, illocutionary force
@@
 discussion. It would be more appropriate to motivate how a robot could benefit
-from a symbolic knowledge base etc.
+from a symbolic knowledge base etc.''
+
+Austin's \emph{speech acts} were actually one of the starting point of our
+analysis of dialogue understanding, and we think it is worthwhile to briefly mention
+them in the introduction, as one of the theoritical ground of our work.
+
+The reviewer however rightfully notes that \emph{perlocutionnary acts}, while mentioned, are not
+actually used further. We moved this paragraph to a simple footnote.
 

### 2011-07-15 paper_knowledge.tex
Moved paragraph on perlocutionnary acts to a footnote

@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{images/pt.jpg} 
+	\includegraphics[width=1.0\linewidth]{images/pt.jpg} 
 	\caption{Interacting with
@@
 upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
-that can be extracted from the sentence that allow to fully ground the action.
-
-Finally, the implicit speaker's expectation can sometimes be conveyed --the
-\emph{perlocutionary act}-- and should be taken into account. If Thomas would
-have said instead ``Jido, this box is too heavy for me'', he may implicitly
-expect the robot to memorize the fact that the box's weight is over his
-capabilities, and it may as well imply that Thomas is requiring help to carry
-the box.
+that can be extracted from the sentence that allow to fully ground the action
+\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deals
+with implicit speaker's expectations. We will not consider it further in this
+article}.
+
+%Finally, the implicit speaker's expectation can sometimes be conveyed --the
+%\emph{perlocutionary act}-- and should be taken into account. If Thomas would
+%have said instead ``Jido, this box is too heavy for me'', he may implicitly
+%expect the robot to memorize the fact that the box's weight is over his
+%capabilities, and it may as well imply that Thomas is requiring help to carry
+%the box.
 

### 2011-07-15 cover_letter.tex
Addressed comment about multi-lingual support

@@
 
-Typography has been throughly revised for better consistency. Notation varients have been
+Typography has been throughly revised for better consistency. Notation variants have been
 made explicit (for instance, section 2.4).
 
-\subsection{``p.6, II.E. explain "multi-lingual support"''}
+\subsection{``p.6, II.E. Explain "multi-lingual support"''}
+
+The "multi-lingual support" has been reworded:
+
+``fast concept lookup, with possible multi-lingual support (through the
+{\tt @lang} XML annotation, labels of concept can be translated, and
+specific translations can be queried for)''
 

### 2011-07-15 cover_letter.tex
Addressed comment on consistency of Fig 15 typography

@@
 
-In the following three sections, we try to address separately each of
+Several reviewers noted linguistic issues and terminology/typography 
+inconsistencies. We have done our best to carefully proof-read the article,
+with the help of a native English speaker.
+
+In the following three sections, we address separately each of
 the concerns of our three reviewers.
@@
 \begin{itemize}
-\item All reviewer comments regarding citations have been addressed, as suggested by the reviewer,
+\item All reviewer comments regarding citations have been addressed, following reviewer's suggestions,
 \item The reviewer's remark on human tendency to explain in a consistent way the world has been added
@@
 
+We thank the reviewer for this input that has been added to the article.
+
 \subsection{``p.9, Alg. IV.1: the crucial bits are left out; please explain
@@
 
-\subsection{``Fig. 15. please make the spelling and the presentation more
-consistent''}
-
 \subsection{``the authors are kindly referred to work on the Generation and

### 2011-07-15 paper_knowledge.tex
Addressed comment about multi-lingual support
Addressed comment on consistency of Fig 15 typography

@@
 	\item statement (\ie RDF triples) insertion, removal, update,
+
 	\item pattern-based statements removal,
+
 	\item pattern-based queries (for instance, \stmt{ * isOn
@@
 	instance, \concept{weight < 150.0}),
+
 	\item consistency check, insertion of statements with consistency
 	constraint (only if the new fact does not lead to inconsistencies),
-	\item fast concept lookup, with multi-lingual support,
-	\item standard SPARQL\footnote{{\em SPARQL Protocol and RDF Query Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
+
+	\item fast concept lookup, with possible multi-lingual support (through the
+	\concept{@lang} XML annotation, labels of concept can be translated, and
+	specific translations can be queried for),
+
+	\item standard SPARQL\footnote{{\em SPARQL Protocol and RDF Query
+	Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
+
 \end{itemize}
@@
 features: while GSM enables representation of time and imaginary objects (whose
-existence is hinted by verbal assertions from a human), SPARK offers a richer
-3D model that enables the computation of several spatial relationships between
+existence is hinted by verbal assertions from a human, also called
+\emph{presupposition accomodation}), SPARK offers a richer 3D model that
+enables the computation of several spatial relationships between
 objects and an effective implementation of perspective taking capabilities.
@@
 	\chatN{rosie} & Servus! The Bavarian robot is up and fit! \\
-	\chatN{rosie} & What is the name of the WHITE CLUSTER on the TABLE? \\
+	\chatN{rosie} & What is the name of the white cluster on the table? \\
 	\chatN{human} & ``Rosie's mug'' \\
-	\chatN{rosie} &\emph{[add \stmt{obj\_01 label "Rosie's mug"}]}\\
+	\chatN{rosie} &\emph{[\stmt{obj\_01 label "Rosie's mug"} added]}\\
 				& What kind of object is ``Rosie's mug''? \\
 	\chatN{human} & Mug \\
-	\chatN{rosie} & \emph{[add \stmt{obj\_01 type Mug}]}\\
+	\chatN{rosie} & \emph{[\stmt{obj\_01 type Mug} added]}\\
 			& What is a Mug?\\ 
 	\chatN{human} & Tableware\\
-	\chatN{rosie} & \emph{[adds \stmt{Mug subClassOf Tableware}. Rosie knows about \emph{Tableware} --it belongs to the common-sense ontology--.}\\
+	\chatN{rosie} & \emph{[\stmt{Mug subClassOf Tableware} added. Rosie already
+	knows about \emph{Tableware} (it belongs to the common-sense ontology), so
+	the interaction stops.]}\\
 	& (...) \\
-	\chatN{rosie} &  What is the name of the BLACK CLUSTER on the TABLE? \\
+	\chatN{rosie} &  What is the name of the black cluster on the table? \\
 	\chatN{human} & ``my\_laptop'' \\

### 2011-07-15 cover_letter.tex
Addressed question on TUM or LMU?
Addressed question on robots architectures + some typography fix

@@
 
-\subsection{``VI. the authors are advised to present an initial overview of the
+\subsection{``Present an initial overview of the
 robot hardware and the common software architecture used in the systems''}
 
+Paragraph added in the section 6 introduction.
+
 \subsection{``VI. which university in Munich? LMU or TUM?''}
+
+The study was conducted with TUM in Munich. This has been corrected in 
+the paper as well.
 

### 2011-07-15 paper_knowledge.tex
Addressed question on TUM or LMU?
Addressed question on robots architectures + some typography fix

@@
     	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
-    	\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = banana\_01}\\
+    	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
 
@@
     	\hspace{0.3cm}\stmt{?obj type Banana} \\
-	\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = banana\_01}\\
+	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
 
@@
 approach are represented.
+
+Experiments have been conducted on two different platforms: the
+\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
+holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
+System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot from
+LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
+Genom/pocolibs} stack.
 
@@
 \begin{center}
-\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [videoTape1, videoTape2]}
+\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1, videoTape2]}
 \end{center}
@@
 \begin{center}
-{\tt ?obj = [videoTape1]}
+\concept{?obj = [videoTape1]}
 \end{center}
@@
 \stmt{?obj type Box}\\
-\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox, toolbox]}
+\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox, toolbox]}
 \end{tabular}
@@
 question. Instead, he decides to point at it while indicating: ``This box''
-(Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
-cardBoardBox} as being pointed at and looked at by the human and updates the
-ontology with this new information using a rule available in the commonsense
-ontology ({\tt pointsAt(?ag, ?obj) $\land$ looksAt(?ag, ?obj) $\to$
-focusesOn(?ag, ?obj)}) The \textsc{Dialogs} module is then able to merge both
-sources of information, verbal (``this'') and gestural to distinguish the box
-Jerry refers to.
+(Figure~\ref{fig|pointing}b). The robot's perception identifies the
+\concept{cardBoardBox} as being pointed at and looked at by the human and
+updates the ontology with this new information using a rule available in the
+commonsense ontology ({\tt pointsAt(?agt, ?obj) $\land$ looksAt(?agt, ?obj)
+$\to$ focusesOn(?agt, ?obj)}) The \textsc{Dialogs} module is then able to merge
+both sources of information, verbal (``this'') and gestural to distinguish the
+box Jerry refers to.
 
@@
 $\to$ \stmt{Jerry focusesAt carboardBox}\\
-\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox]}
+\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
 \end{tabular}
@@
 \stmt{?obj isIn cardBoardBox}\\
-\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = videoTape2}\\
+\hspace{0.7cm}$\Rightarrow$ \concept{?obj = videoTape2}\\
 \end{tabular}
@@
 \stmt{?obj differentFrom videoTape2}\\
-\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [videoTape1]}
+\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1]}
 \end{tabular}

### 2011-07-15 paper_knowledge.tex
Fixed some issues with biblio + added Sisbot2011 on SPARK

@@
 and second, a symbolic assessment of the percepts. In this section we present
-SPARK (\emph{SPAtial Reasoning \& Knowledge}), a situation assessment reasoner
+SPARK (\emph{SPAtial Reasoning \& Knowledge}, \cite{Sisbot2011}), a situation assessment reasoner
 that generates relevant symbolic information from the geometry of the

### 2011-07-15 cover_letter.tex
Fixed some issues with biblio + added Sisbot2011 on SPARK
Addressed the question on the Open World Assumption

@@
 Assumption''}
+
+The Open World assumption mainly allows us to model that we may not know
+if a fact is true or not.
+
+One very common example is the {\tt knowsAbout} predicate. This predicate
+defines if some agent knows about some object.
+
+It is important (especially for dialog) to be able to equally model that an agent knows or 
+does not know about the object, or that the robot simply does not know if 
+the human knows about it.
+
+The Open World Assumption has important consequences on the reasoning
+abilities of the system, since it prevents the classical "negation by 
+failure" and leads to tighter modelling requirement to locally "close the
+world" (for instance, the {\tt Color} class is defined as the equivalent of
+the union of the 10 or so color instances we consider).
 
@@
 
-\subsection{``please improve the quality of the references (sometimes, the
-publisher is missing, sometimes capitalization is omitted, sometimes
-abbreviations are used and sometimes full names''}
-
 \end{document}

### 2011-07-15 paper_knowledge.tex
Raquel changes to the experiments and conclusion parts

--- 2011-07-15 paper_knowledge.tex
@@
 on situated language grounding. We present several examples that include a
-generic recognition and semantic validation of thematic roles (\ie, the
+generic recognition and semantic validation of thematic roles (\ie the
 semantic roles played by the different nominal groups around a verb), which
@@
 
-Section 5 presents three use-cases that were conducted on three different
-robotic platforms: the \emph{naming} experiment, where a robot anchors new
+Section 5 presents three use-cases that were conducted on two different
+robotic platforms: the \emph{Naming} experiment, where a robot anchors new
 knowledge in its model through verbal interaction; the \emph{Spy Game} where
@@
 
-Anchoring perceptions in a symbolic model requires first perception abilities
-and second, a symbolic assessment of the percepts. In this section we present
-SPARK (\emph{SPAtial Reasoning \& Knowledge}), a situation assessment reasoner
+Anchoring perceptions in a symbolic model requires perception abilities
+and their symbolic assessments. In this section we present
+SPARK (\emph{SPAtial Reasoning \& Knowledge}, \cite{Sisbot2011}), a situation assessment reasoner
 that generates relevant symbolic information from the geometry of the
-environment with respect to relations between objects and human capabilities.
+environment with respect to relations between objects, robots and humans.
 Moreover, the notion of \emph{Perspective
-Taking}~\cite{Flavell1992,Tversky1999} is employed in the heart of the
-reasoning to provide the robot with the ability to put itself at the human's
+Taking}~\cite{Flavell1992,Tversky1999} is employed at the heart of the
+reasoner to provide the robot with the ability to put itself at the human's
 place and to reason about the world from different perspectives.
 
-As mentioned in the introduction, this paper do not focus on the sensor-level
+As mentioned in the introduction, this paper does not focus on the sensor-level
 perception. We rather assume that the perception of the humans and the objects
@@
 There are a number of common properties for a robot and a human related to
-their capabilities in a given situation: they can both reach, grasp, look,
-point, etc. In our context, we group robots and humans into a single category.
+their capabilities in a given situation: they can both reach, grasp, look at,
+point at, etc. In our context, we group robots and humans into a single category.
 Thus, we define agents as entities that can act in the environment and
@@
 computed from each agent point of view, and therefore, also stored in different
-models in ORO for further use at the decisional level.}.
+models in ORO for further use at the decisional level.}:
 
@@
 holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
-System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot from
+System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from
 LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
@@
 \centering
-
-\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/kimp1.jpg}} \hspace{0.2em}
-\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/rviz.png}}
+\includegraphics[width=0.45\columnwidth]{images/kimp1.jpg}
+\includegraphics[width=0.45\columnwidth]{images/rviz.png}
+%\parbox[c]{4.3cm}{\includegraphics[width=0.5\columnwidth]{images/kimp1.jpg}}
+%\parbox[c]{4.1cm}{\includegraphics[width=4cm]{images/rviz.png}}
 \flushleft
 \hspace{1.8cm} (a) \hspace{3.9cm} (b)
-\caption{(a) Rosie, looking for objects it may know, and (b) viewed in Willow
+\caption{(a) Rosie, looking for objects it may know, and (b) view from Willow
 Garage's RViz. The clusters of point are given a unique identifier by the
@@
 
-The robot picks an unknown object from the table, shows it to the user, and
+The robot selects an unknown object from the table, shows it to the user, and
 asks about its name and type (Figure~\ref{fig|kimpwatching}). The user
@@
 naming experiment. The human describes the object until the robot recognizes a
-type it already knows: by asking in loop the human for the categories of an
-object until it can connect it to a concept it already knows, the robot
+type it already knows.
+More precisely, the robot recursively asks the human for the categories the object belongs to until reaching a known one.
+At this point, the robot is able to connect the new object information to already known concepts.
+Therefore, the robot
 accurately anchors perception in its symbolic model and it is able to reason
@@
 
-The supervision (\textsc{cram}\footnote{\textsc{cram} (Cognitive Robotic
+The supervision module (\textsc{cram}\footnote{\textsc{cram} (Cognitive Robotic
 Abstract Machine) is a RPL-derived framework for rapid development of cognitive
@@
 %(the perception is based on the \textsc{CoP} framework~\cite{Klank2009}). 
-Therefore, the integration of ORO can be seen as an extension to
+Therefore, the integration of ORO can be seen as an extension of
 the robot's belief state that not only contains abstract identifiers
@@
 discarded the new fact. This kind of logical failure can be used to improve
-low-level perception results by ``closing the loop'' with high-level, symbolic
+low-level perception outcomes by ``closing the loop'' with high-level, symbolic
 knowledge.
 
-\subsection{``Spy Game'' experiment}
+\subsection{\emph{Spy Game} experiment}
 
@@
 \centering
-
-\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/spy-game-real.jpg}} \hspace{0.2em}
-\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/spy-game-mhp.jpg}}
+\includegraphics[width=0.45\columnwidth]{images/spy-game-real.jpg}
+\includegraphics[width=0.45\columnwidth]{images/spy-game-mhp.jpg}
+%\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/spy-game-real.jpg}} \hspace{0.2em}
+%\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/spy-game-mhp.jpg}}
 \flushleft
@@
 
-We have integrated the game in the LAAS-CNRS Jido robot~\cite{Alami1998a}.
+We have integrated the game on the LAAS-CNRS Jido robot.
 Objects are identified through a tag-based vision approach\footnote{ARToolKit:
@@
 		\chatN{human} & It is a tableware.\\
-		\chatN{jido} & \emph{[retrieves possible objects: coke-bottle, yellow-bottle, orange-bottle, cup-with-handle]} \\
-	 			& \emph{[keeps visible objects: coke-bottle, yellow-bottle, cup-with-handle]}\\
-				& \emph{[obtains discriminants: type, color.]}\\
+		\chatN{jido} & \emph{[retrieve possible objects: coke-bottle, yellow-bottle, orange-bottle, cup-with-handle]} \\
+	 			& \emph{[keep visible objects: coke-bottle, yellow-bottle, cup-with-handle]}\\
+				& \emph{[obtain discriminants: type, color.]}\\
 				& Which type of object is: bottle or cup? \\
 		\chatN{human} & Bottle.\\
-		\chatN{jido} & \emph{[obtains possible objects: coke-bottle, yellow-bottle.]}\\
-				& \emph{[obtains discriminants: color.]}\\
-				& What color the object is: red or yellow?\\
+		\chatN{jido} & \emph{[obtain possible objects: coke-bottle, yellow-bottle.]}\\
+				& \emph{[obtain discriminants: color.]}\\
+				& What color is the object: red or yellow?\\
 		\chatN{human} & Red.\\
@@
 
-In order to illustrate some of different reasoning abilities of the robot,
-based on the approach presented in this paper, we have designed the following
+In order to illustrate some of different reasoning abilities of the robot presented in this paper, we have designed the following
 daily life situation. Tom and Jerry are moving to London, so they are packing
@@
 scope of this work, we do not include any motion from the robot's side. Similar
-to the Spy Game scenario, perception of objects is done through a tag-based
+to the \emph{Spy Game} scenario, perception of objects is done through a tag-based
 system and humans are detected through motion capture. We next describe in
@@
 
-First Tom enters the room and places on the table some of the things to pack: a
+First Tom enters the room with some of the things they need to pack: a
 toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
@@
 
-The initial information in the ontology corresponds to:
+The initial information in ORO corresponds to:
 
@@
       \stmt{toolbox type Box}\\
-      \stmt{videoTape1 type VideoiTape}\\
+      \stmt{videoTape1 type VideoTape}\\
       \stmt{videoTape1 label "Lord of the Robots"}\\
@@
 
-There is a cardboard box on the table:
+SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:
 
@@
 He places a toolbox and a video tape on the table, and another video tape
-inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b):
+inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:
 
@@
 
-Since only one object is available, the robot infers
+Since only one object is available now, the robot infers
 that the human refers to it and can eventually execute the command, \ie give
@@
 question. Instead, he decides to point at it while indicating: ``This box''
-(Figure~\ref{fig|pointing}b). The robot's perception identifies the
-\concept{cardBoardBox} as being pointed at and looked at by the human and
-updates the ontology with this new information using a rule available in the
-commonsense ontology ({\tt pointsAt(?agt, ?obj) $\land$ looksAt(?agt, ?obj)
-$\to$ focusesOn(?agt, ?obj)}) The \textsc{Dialogs} module is then able to merge
-both sources of information, verbal (``this'') and gestural to distinguish the
-box Jerry refers to.
-
+(Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
+cardBoardBox} as being pointed at and looked at by the human and updates the
+ontology with this new information using a rule available in the commonsense
+ontology:
 \begin{center}
@@
 $\to$ \stmt{Jerry focusesAt carboardBox}\\
+\end{tabular}
+\end{center}
+In the meantime, the \textsc{Dialogs} module is processing the human verbal input.
+When trying to resolve the reference ``this'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
+\begin{center}
+\begin{tabular}{l}
+\stmt{Jerry focusesAt ?obj}\\
 \hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
@@
 modalities in a single homogeneous environment, bringing mutual benefits to
-components: for instance, the dialogue processing module can perfectly run
+components. For instance, the dialogue processing module can perfectly run
 without any geometric perception, but its disambiguation routines can
@@
 Regarding the anchoring question, this architecture is
-bidirectional: the components we described provide a \textit{bottom-up}
+bidirectional. The components we described provide a \textit{bottom-up}
 grounding process: SPARK and \textsc{Dialogs} constantly build and push new
@@
 cameras or alike), but someone can verbally describe a scene to it, while the
-deaf robot has a good vision system but can not hear anything.  Without any
-changes to the software architecture above the language processing module or
-the vision module, supervisors of both robot would be able to perform equally
-well.
+deaf robot has a good vision system but can not hear.  Without any
+changes to the software architecture described, supervisors of both robots would be able to perform equally well.
 
@@
 enabling independent modeling of the beliefs of each agent, our architecture is
-especially well suited to set up \emph{False-Belief}
-experiments~\cite{Leslie2000}, and we plan to further explore those kinds of
-applications.
+especially well suited to set up cognitive and psychological experiments (such as the \emph{False-Belief} experiment~\cite{Leslie2000}), which we plan to further explore.
 
@@
 
-For us, \emph{Grounding} is on the contrary a \emph{top-bottom} activity: the
+For us, \emph{grounding} is on the contrary a \emph{top-bottom} activity: the
 robot needs to automatically bind a representation (for instance, a word
 uttered by a human, an image taken from a camera, a sentence extracted from
-Wikipedia) to an unambiguous, context-dependent, internal concept. This very
-concept can (or not) be \textit{a priori} available to the robot as a
+Wikipedia) to an unambiguous, context-dependent, internal concept. This 
+concept may (or may not) be \textit{a priori} available to the robot as a
 pre-loaded ontology (what we previously called the cultural background of the
@@
 robot make use of its embodied nature to establish a meaningful communication
-with a human: because the robot and the human share the same physical
+with a human. Mainly, because the robot and the human share the same physical
 environment and they perceive each other, we are able to create a mutual
@@
 human-robot interaction. Our approach has hopefully made clear that, similar to
-Beetz et al.~\cite{Beetz2010}, we do not consider embodiment \emph{per se}, outside of
-a broader symbolic system: our architecture is not bound to the morphology or
+Beetz et al.~\cite{Beetz2010}, we do not consider embodiment \emph{per se} outside of
+a broader symbolic system, \ie our architecture is not bound to the morphology or
 the low-level sensori-motor capabilities of a specific agent.
@@
 However, we can build a model of the ``human point of view'' because the robot
-sees the human, and is able to deduce, at least partially, what the human
-perceives or not. This relies on the embodied nature of the interaction. We
-infer that a human focuses on some object because it points at it, looks at it,
-and besides, the object is visible to him. This relies as well on the embodied
-nature of the interaction. And this allow us, in turn, to understand the
+perceives the human, and is able to deduce, at least partially, what the human
+perceives or not. We infer that a human focuses on some object because it points at it, looks at it, and besides, the object is visible to him. This relies on the embodied
+nature of the interaction. And this allows us, in turn, to understand the
 meaning of sentences like ``Give me this''.
@@
 
-Let conclude this article by a short glance back at Roy's list of challenges
+Let us conclude this article by a short glance back at Roy's list of challenges
 for human-robot dialogue: 
@@
 	\item While more modalities (especially, deictic gestures and social gazes)
-	can be added, we have proposed in this article a \emph{cross-modal
+	can be added, we have actually proposed a \emph{cross-modal
 	representation system},

### 2011-07-16 paper_knowledge.tex
Proof read of the first 10 pages

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Grounding human interaction into the robot knowledge}
+\section{Grounding human interaction into robot knowledge}
 
@@
 
-``Jido, give me this'', says Thomas while looking at a box that contains a
+``Jido, give me that'', says Thomas while looking at a box that contains a
 video tape. The robot smoothly grasps the tape, and hands it to the him.
@@
 While this kind of interaction should hopefully sound quite familiar in a
-foreseeable future, our robots are not yet quite up to the task. Neither
+foreseeable future, our robots are not yet quite up to the task, neither
 regarding natural language understanding nor plan-making and manipulation.
@@
 representation of concepts (objects, agents, actions...) underlying the
-interaction: what are the prerequisites for such a human sentence --``Jido,
-give me this''-- to be understood by the robot, correctly interpreted in the
+interaction: what are the prerequisites for such a sentence --``Jido,
+give me that''-- to be understood by the robot, correctly interpreted in the
 spatial context of the interaction, and eventually transformed into an action?
 
-Austin~\cite{Austin1962} would have at first glance analyzed such kind of
+Austin~\cite{Austin1962} would have at first glance analyzed this kind of
 sentence as a \emph{speech act}, comprising of \emph{locutionary},
@@
 referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
-``\textit{me}''? And ``\textit{this}''?
+``\textit{me}''? And ``\textit{that}''?
 
@@
 parametrization is conveyed by the semantics attached to the words and the
-grammatical structures of the sentence. In our example, the type of action is
+grammatical structure of the sentence. In our example, the type of action is
 given by the verb ``\textit{give}''. Assuming the robot has some procedural
@@
 upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
-that can be extracted from the sentence that allow to fully ground the action
+that can be extracted from the sentence that allow for fully grounding the action
 \footnote{The third class of speech acts, the \emph{perlocutionary acts}, deals
@@
 
-Extracting these speech acts and turning them into a content processable by the
+Extracting these speech acts and turning them into content processable by the
 robot is a difficult challenge in the general case. We base our approach on
@@
 (sometimes referred as \emph{assertive}) and \emph{directive} illocutionary
-acts in Searle~\cite{Searle1976} classification.
+acts in Searle's~\cite{Searle1976} classification.
 
@@
 for situated verbal interaction and integrated them into a broader control
-architecture for indoors service robots.
+architecture for indoor service robots.
 
@@
 elaborates on the need of embodiment for symbol grounding, and
-Sloman~\cite{Sloman2007} who criticizes the \emph{``concept emprism"} that
+Sloman~\cite{Sloman2007} who criticizes the \emph{``concept empricism"} that
 underlies symbol grounding.
@@
 
-Processing natural language in situated context is already an established
+Processing natural language in situated contexts is already an established
 research field. In~\cite{Roy2005}, Roy and Reiter summarize what they see as
@@
 
-Dialogue processing in real robots have been explored by several teams.  Brick
+Dialogue processing in real robots has been explored by several teams.  Brick
 and Scheutz~\cite{Brick2007} have contributions regarding natural language
 processing in an incremental way, and how this enables instant back-channel
-feedback (like nodding). Hüwel et al.~\cite{Huwel2006} propose the concept of
-\textit{Situated Semantic Unit}: atoms are extracted from sentences exposing
+feedback (like nodding). H\"uwel et al.~\cite{Huwel2006} propose the concept of the
+\textit{Situated Semantic Unit}: atoms are extracted from sentences, exposing
 semantic links to other units. The parser tries to satisfy these links and
@@
 Also worth mentioning, Mavridis and Roy~\cite{Mavridis2005} propose the idea of
-a \emph{grounded situation model} which is an a-modal model of the world where
+a \emph{grounded situation model} which is an amodal model of the world where
 different sensing modalities, including verbal ones (the robot is able to
-\emph{imagine} objects), are merged. Their framework also allows management of
+\emph{imagine} objects), are merged. Their framework also allows the management of
 the interaction history (the human can ask for a past event). They propose an
 implementation in an environment built on simple entities (a manipulator arm
-and color balls).
+and colored balls).
 
@@
 \textit{ResearchCyc} ontology (including the \textit{MicroTheories} concept),
-used in combination with the \textsc{CycL} language. This enables second-order
+used in combination with the CycL language. This enables second-order
 logic modeling and access to a large common-sense knowledge base.
@@
 (\url{http://www.w3.org/TR/owl2-primer/}).} ontology, derived from
-\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KownRob} have been demonstrated on
-several real-world scenarios, where natural language recipes extracted from
+OpenCyc. \textsc{CRAM} and \textsc{KownRob} have been demonstrated on
+several real-world scenarios, where natural language recipes extracted from the
 Internet had to be translated into plans and executed in a kitchen environment,
@@
 agents' belief states is used for the interaction and eases the implementation
-of various advanced cognitive behaviours like learning new concepts or
-interactive objects discrimination.
+of various advanced cognitive behaviors like learning new concepts or
+interactive object discrimination.
 
@@
 on-line by merging different sensor modalities. It computes spatial relations
-between perceived objects in real-time and it allows to virtually \emph{view}
-the same scene from different points of view, enabling \emph{visual} and
+between perceived objects in real-time and it allows for virtually \emph{viewing}
+of the same scene from different points of view, enabling \emph{visual} and
 \emph{spatial agent perspective taking}.
@@
 robust, situated and more generic than what can be found in previous work
-on situated language grounding. We present several examples that include a
+on situated language grounding. We present several examples that include
 generic recognition and semantic validation of thematic roles (\ie the
@@
 that has been deployed and tested on three distinct robotic platforms
-(including both humanoid robot and service robots), demonstrating the
+(including both humanoid robots and service robots), demonstrating the
 versatility and hardware-agnosticism of these developments.
@@
 robotic platforms: the \emph{Naming} experiment, where a robot anchors new
-knowledge in its model through verbal interaction; the \emph{Spy Game} where
+knowledge in its model through verbal interaction; the \emph{Spy Game}, where
 either the user or the robot tries to guess which object the other player is
@@
 While \emph{knowledge} has no general definition that researchers agree on, for
-our own purposes we define knowledge as \emph{an information interpreted in the
-cultural and social context of the robot}, where an information is a
-\emph{statement} or an \emph{assertion} on the world\footnote{In this paper,
+our own purposes we define knowledge as \emph{information interpreted in the
+cultural and social context of the robot}, where information is a
+\emph{statement} or an \emph{assertion} about the world\footnote{In this paper,
 statements are always triples \texttt{<subject, predicate, object>}, following
 the Description Logics formalism.}. In practical terms, knowledge is made of
-statements that are \emph{contextualized}, if possible \emph{synthesized} and
+statements that are \emph{contextualized}, if possible \emph{synthesized}, and
 \emph{limited} to a domain of validity.
 
-These three features have important consequences on the way a knowledge
+These three features have important consequences for the way a knowledge
 representation and storage system must be designed. Let us examine them:
@@
 We call \textit{cultural context} a broad set of common, general facts that are
-considered widely accepted among the inter-actors (\eg ``bottles may contain
+considered widely accepted among the interactors (\eg ``bottles may contain
 water''). This knowledge is often referred as \emph{common-sense knowledge}.
@@
 effectively represented rather than actually identifying the current context.
-While several approaches to build contextualized knowledge are proposed in this
+While several approaches for building contextualized knowledge are proposed in this
 paper (symbolic environment interpretation, perspective taking, grounded
@@
 infer that ``Peter'' may be the name of the human. \textit{Synthesis}
-refers to the fact that several, \textit{a priori} non correlated, facts
-must be associated to the same common concept. This process requires the
+refers to the fact that several, \textit{a priori} uncorrelated, facts
+must be associated with the same common concept. This process requires the
 ability to control the logical consistency of the knowledge corpus. To
@@
 
-\paragraph{Domain of validity} its main role is to specify the scope in which
-an information is (believed to be) true. It covers several aspects: temporal,
+\paragraph{Domain of validity}: its main role is to specify the scope in which
+ information is (believed to be) true. It covers several aspects: temporal,
 situational and probabilistic. While related to the previous concept of
@@
 These three aspects lead us to envisage a knowledge representation system
-characterized by the following features: 
+characterized by the following abilities: 
 \begin{itemize}
@@
 
-Besides, the following active processes would be desirable to:
+Besides, the following active processes would be desirable:
 \begin{itemize}
@@
 	explicit inconsistencies when required\footnote{One may argue that the real world is 
-	inherently inconsistent. In this article, we however make a
+	inherently inconsistent. In this article, we make a
 	\textit{consistent world} assumption, in order to leverage reasoning
@@
 This list does not cover all the possible features that could be exposed by a
-symbolic knowledge management system. Bio-inspired memory management ---the
-ability to forget or reinforce knowledge--- or curiosity ---the ability to
+symbolic knowledge management system. Bio-inspired memory management (the
+ability to forget or reinforce knowledge) or curiosity (the ability to
 identify lacking knowledge and actively trigger behaviors to acquire
-it\fixme{cite Howa}---, to give some examples, could arguably be added to the
+it\fixme{cite Howa}), to give some examples, could arguably be added to the
 list. However, this first analysis sets a convenient reference frame to
@@
 cognitive and human-robot interaction related functionalities are available
-(detailed thereafter). External plugins can also be easily hooked. The modules
+(detailed thereafter). External plugins can also be easily added. The modules
 rely on several parallel ontology \emph{back-ends} where the knowledge is
@@
 continuous classification of the OWL concept graph: at run-time, newly added
-statements are continuously reasoned about (\emph{classified}), and at anytime,
+statements are continuously reasoned about (\emph{classified}), and at any time,
 the server expose a complete set of statements, both the asserted ones and the
@@
 and much more concrete concepts as \concept{Table}, \concept{Book} or colors.
-Available predicates allow to describe the state of the agents and the world
+Available predicates allow us to describe the state of the agents and the world
 with relations like \concept{isOn}, \concept{sees},
@@
 Our common sense ontology is closely aligned with the open-source
-\textsc{OpenCyc}\footnote{\url{http://www.opencyc.org}} upper ontology.
-\textsc{OpenCyc} defines a large taxonomy of concepts and semantic
+OpenCyc\footnote{\url{http://www.opencyc.org}} upper ontology.
+OpenCyc defines a large taxonomy of concepts and semantic
 relationships between concepts that are used in several other projects
@@
 exchange with other robots (for instance, the works previously mentioned by
-Daoutis and Tenorth rely on the same \textsc{Cyc} concepts\fixme{Give
+Daoutis and Tenorth rely on the same Cyc concepts\fixme{Give
 examples}).
@@
 We provide in ORO accessors to query, add or remove all these properties and
-restrictions (however, not the SWRL rules) at run-time. This allows
+restrictions (except the SWRL rules) at run-time. This allows
 knowledge introspection and enables the robot to alter its own knowledge
@@
 exploitation) of a consistent knowledge model, the \emph{Common Sense Ontology}
-helps building assertions that are related to previous one by offering
+helps build assertions that are related to previous one by offering
 a predefined vocabulary.
@@
 agent (or infers that some instance is an agent), it automatically creates a
-new, separate, in-memory OWL model for that agent. Thus, different robot's
+new, separate, in-memory OWL model for that agent. Thus, different robot
 components, like supervision or situation assessment, may then store the
@@
 robot's primary model are equally available in every agent's model, which
-allows to store and reason on different (and possibly globally inconsistent)
+allows us to store and reason on different (and possibly globally inconsistent)
 models of the world.
@@
 \emph{theory of mind}~\cite{Scassellati2002}: the robot can explicitly model
-the belief state of its inter-actors, opening new possibilities for the control
-architecture. In the section~\ref{dialog} we present how we use this feature to
-make sense of the user sentences from the his/her point of view. Moreover,
-these multiple models can be viewed as well as different interpretive scopes,
-allowing to interpret the same reality from different points of view.
-
-These multiple models can be viewed as well as different interpretive
-scopes, allowing to interpret the same reality from different points of
-view.
+the belief state of its interactors, opening new possibilities for the control
+architecture. In section~\ref{dialog} we present how we use this feature to
+make sense of user sentences from his/her point of view. Moreover,
+these multiple models can be viewed as different interpretive scopes,
+allowing the robot to interpret the same reality from different points of view.
 
@@
 ambiguity produced by its human partner. For instance, a user may refer to a
-``bottle'' whereas two bottles are currently visible. Discrimination routines
+``bottle'' where two bottles are currently visible. Discrimination routines
 can identify possible (symbolic) differences (\eg the color or the size of the
@@
 
-Lastly, ORO allows external modules to be triggered when specific events occur.
-For instance, when a logical sentence becomes true or false, or if a new
+Lastly, ORO allows external modules to be triggered when specific events occur,
+ for instance, when a logical sentence becomes true or false, or if a new
 instance of a certain class is added. One immediate application is reactive
@@
 \item \emph{Looks At}: this relation corresponds to what the agent is focused
-on, \ie where its focus of attention is directed to. This model is based on a
+on, \ie where its focus of attention is directed. This model is based on a
 narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
@@
 
-\item \emph{Points At}: verifies if an object is pointed at by an agent. This
+\item \emph{Points At}: verifies whether an object is pointed at by an agent. This
 relation is particularly useful during interaction when one of the agents is
@@
 
-\item \emph{Reachable}: it allows the robot to estimate the agents capability
+\item \emph{Reachable}: it allows the robot to estimate the agent's capability
 to reach an object, which is fundamental for task planning. For example, if the
@@
 
-\subsection{Locations} As humans, one way of referring to object's
-positions is based on their symbolic descriptors instead of using their precise
-position. In fact, in many cases, this information is the most precise one that
-we can give since we do not store the numeric coordinates of objects. The
+\subsection{Locations}
+
+One way of referring to object's
+positions is based on a human's symbolic descriptors, instead of using their precise
+position. In fact, in many cases, this information is the most precise information available
+ since we do not store the numeric coordinates of objects. The
 following relations are computed with respect to the position of the agents and
@@
 \concept{isLocatedAt} represents spatial locations between agents and objects.
-For example we say ``it is on my right, on your left,...'' We compute these
+For example we say ``it is on my right, on your left, ...'' We compute these
 spatial locations by dividing the space around the referent (an agent) into $n$
@@
 order) we need to retain the semantics of the verb and its complements.
-\emph{Thematic roles} allow to semantically link a verb to its complements.  
+\emph{Thematic roles} allow for semantically linking a verb to its complements.  
 There is no general agreement amongst linguists on a comprehensive list of 
@@
 asks about its name and type (Figure~\ref{fig|kimpwatching}). The user
-interacts with the robot through a dedicated bridge, allowing to chat with
+interacts with the robot through a dedicated bridge, allowing him/her to chat with
 the robot with a standard Jabber instant messaging client.

### 2011-07-16 paper_knowledge.tex
Add references to spark

@@
 that can be extracted from the sentence that allow for fully grounding the action
-\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deals
+\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deal
 with implicit speaker's expectations. We will not consider it further in this
@@
 \concept{isLocatedAt} represents spatial locations between agents and objects.
-For example we say ``it is on my right, on your left, ...'' We compute these
+For example we say ``it is on my right, on your left, ...'' These type of direction commands are studied
+in the literature in the context of language grounding (\eg \cite{OKeefe1999,Matuszek10}). We compute these
 spatial locations by dividing the space around the referent (an agent) into $n$
@@
 grounding (\eg \cite{Regier01} presented different models to define the
-\emph{above} relation). In this work we use similar models based on the
+\emph{above} relation; \cite{Kelleher09} presented two computational models of topological and projective spatial prepositions; \cite{Blisard05} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
+In this work we use similar models based on the
 bounding boxes and center of mass of the objects to define three main relations

### 2011-07-16 cover_letter.tex

@@
-\documentclass{article}
+\documentclass[11pt]{article}
+\usepackage[inner=2cm,outer=2cm, top=2cm, bottom=2cm]{geometry}               % See geometry.pdf to learn the layout options. There are lots.
 
@@
 
+\title{Responses to reviewers' comments and list of modifications done\\ on paper !!!!!SOROXX paper ID!!!!!\\
+\LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
+\author{S\'everin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami \and Michael Beetz}
+
 \begin{document}
 
-We are very grateful to the reviewers for having accepted to review our 
-work, for the care they took in examining it and their resulting 
-comments which have been appreciated in order to improve the quality of
-the paper.
+\maketitle
 
-Several reviewers noted linguistic issues and terminology/typography 
-inconsistencies. We have done our best to carefully proof-read the article,
-with the help of a native English speaker.
+The authors would like to thank the reviewers for their constructive feedback that helped us to strengthen the paper.
 
-In the following three sections, we address separately each of
-the concerns of our three reviewers.
+Below follow our answers to the reviewers’ comments and suggestions.
+Each answer also indicates the changes made in the paper to address the reviewer’s concern.
+An important number of language corrections are made, and the paper is proof-read by a native English speaker.
 
-\section{Comments of the first reviewer}
+%We are very grateful to the reviewers for having accepted to review our 
+%work, for the care they took in examining it and their resulting 
+%comments which have been appreciated in order to improve the quality of
+%the paper.
+
+%Several reviewers noted linguistic issues and terminology/typography 
+%inconsistencies. We have done our best to carefully proof-read the article,
+%with the help of a native English speaker.
+
+%In the following three sections, we address separately each of
+%the concerns of our three reviewers.
+
+\section{Reviewer 1}
+%\section{Comments of the first reviewer}
 
@@
 to be sure about something (for instance, "I think that..."), which represents
-the certainty the robot has about a fact. In the meantime, to overcome
+the uncertainty the robot has about a fact. In the meantime, to overcome
 perception errors, the robot provides visual feedback (looking at objects) to
@@
 
-\section{Comments of the second reviewer}
+\section{Reviewer 2}
+% \section{Comments of the second reviewer}
 
@@
 
-\section{Comments of the third reviewer}
+\section{Reviewer 3}
+%\section{Comments of the third reviewer}
 
@@
 \subsection{``p.8, III.B.Location according to an object: the authors are
-kindly referred to the work done by, e.g., K. Sjöö or J. Kelleher''}
+kindly referred to the work done by, e.g., K. Sj\"o\"o or J. Kelleher''}
+
+Kelleher's work, as well as Blisard's, are now correctly cited in the section.
 

### 2011-07-16 cover_letter.tex
Added SORO num + typo

@@
 
-\title{Responses to reviewers' comments and list of modifications done\\ on paper !!!!!SOROXX paper ID!!!!!\\
+\title{Responses to reviewers' comments and list of modifications done\\ on paper SORO226\\
 \LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
@@
 
-At recent video showing the current state of the interaction can be viewed
+A recent video showing the current state of the interaction can be viewed
 here: \url{http://www.youtube.com/watch?v=IODx50uV_k4}
@@
 actually used further. We moved this paragraph to a simple footnote.
-
-I very strongly recommend that section I.A. be completely revised.
 

### 2011-07-16 paper_knowledge.tex
Added a reference to Zender's work

@@
 \author{
-Séverin Lemaignan \and
+S\'everin Lemaignan \and
 Raquel Ros \and
@@
 \institute{ 
-	Séverin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami
+	S\'everin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami
 	\at
@@
 	\and
-	Séverin Lemaignan \and Michael Beetz
+	S\'everin Lemaignan \and Michael Beetz
 	\at
@@
 \concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}} using
-two different models).  Our platform relies on OWL ontologies (a decidable
-subset of the predicate logics) featuring continuous storage, querying and
-event triggering over the pool of facts known by the robot.
+two different models).  Our platform relies on OWL-DL\footnote{\emph{Web
+Ontology Language - Description Logics}, a decidable subset of the first-order
+logics, \url{http://www.w3.org/TR/owl2-primer/}} ontologies and features
+continuous storage, querying and event triggering over the pool of facts known
+by the robot.
 
@@
 
-Dialogue processing in real robots has been explored by several teams.  Brick
+Dialogue processing in real robots have been explored by several teams.  Brick
 and Scheutz~\cite{Brick2007} have contributions regarding natural language
 processing in an incremental way, and how this enables instant back-channel
-feedback (like nodding). H\"uwel et al.~\cite{Huwel2006} propose the concept of the
-\textit{Situated Semantic Unit}: atoms are extracted from sentences, exposing
+feedback (like nodding). Hüwel et al.~\cite{Huwel2006} propose the concept of
+\textit{Situated Semantic Unit}: atoms are extracted from sentences exposing
 semantic links to other units. The parser tries to satisfy these links and
@@
 utterances. They validated the approach with an extensive user-study.
+
+Zender et al.~\cite{Zender2009} address the generation of referring expressions
+(\emph{GRE}) in situated dialogue for topological knowledge.  They consider
+both the reference resolution and reference description tasks, and rely on
+OWL-DL representation and SPARQL\footnote{{\em SPARQL Protocol and RDF Query
+Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} to extract
+\emph{topological contexts} from their knowledge base.
 
@@
 \textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
-Prolog. Its underlying storage is based on an RDF/OWL\footnote{RDF/OWL is a
-standard for Description Logics based ontology
-(\url{http://www.w3.org/TR/owl2-primer/}).} ontology, derived from
-OpenCyc. \textsc{CRAM} and \textsc{KownRob} have been demonstrated on
-several real-world scenarios, where natural language recipes extracted from the
+Prolog. Its underlying storage is based on an OWL ontology, derived from
+\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KownRob} have been demonstrated on
+several real-world scenarios, where natural language recipes extracted from
 Internet had to be translated into plans and executed in a kitchen environment,
@@
 
-	\item standard SPARQL\footnote{{\em SPARQL Protocol and RDF Query
-	Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} queries.
+	\item standard SPARQL queries.
 
@@
               direct objects:
-                NP: \textbf{bottle}
+                NP: my \emph{nice blue} \textbf{bottle}
               indirect objects:

### 2011-07-16 paper_knowledge.tex
Fixed biblio issues between Akin and I

@@
 For example we say ``it is on my right, on your left, ...'' These type of direction commands are studied
-in the literature in the context of language grounding (\eg \cite{OKeefe1999,Matuszek10}). We compute these
+in the literature in the context of language grounding (\eg \cite{O'Keefe1999,Matuszek2010}). We compute these
 spatial locations by dividing the space around the referent (an agent) into $n$
@@
 next to, in}, etc. These types of relations are widely studied in language
-grounding (\eg \cite{Regier01} presented different models to define the
-\emph{above} relation; \cite{Kelleher09} presented two computational models of topological and projective spatial prepositions; \cite{Blisard05} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
+grounding (\eg \cite{Regier2001} presented different models to define the
+\emph{above} relation; \cite{Kelleher2009} presented two computational models of topological and projective spatial prepositions; \cite{Blisard2005} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
 In this work we use similar models based on the

### 2011-07-16 cover_letter.tex
Added a reference to Zender's work
Addressed comment on 'Give examples of OpenCyc alignement

@@
 
-Below follow our answers to the reviewers’ comments and suggestions.
-Each answer also indicates the changes made in the paper to address the reviewer’s concern.
+Below follow our answers to the reviewersÕ comments and suggestions.
+Each answer also indicates the changes made in the paper to address the reviewerÕs concern.
 An important number of language corrections are made, and the paper is proof-read by a native English speaker.
@@
 
+Numerous predicates in {\tt commonsense.oro.owl} are defined in the 
+OpenCyc namespace: {\tt farFrom, near, hold-Underspecified, objectFoundInLocation}.
+
+Similarly, most of the upper classes in the ontology belong to OpenCyc 
+(like {\tt SpatialThing, PartiallyTangible, TemporalThing}...).
+
+Overall, about 50 entites (on the >180 defined in {\tt commonsense.oro.owl}) 
+come from OpenCyc.
+
+Since most of the upper classes are OpenCyc concepts, this should 
+guarantee to a certain extend the alignment of our ontology with other ontologies.
+
 \subsection{Typography and notation consistency}
@@
 
+A reference to Zender's work has been added to the \emph{Related work} section.
+
 \subsection{``Present an initial overview of the

### 2011-07-16 cover_letter.tex
Addressed remarks concerning curiosity + generation of refering expressions and added relevant papers

@@
 
+A mention to the Dora robot has been added in section 2.1.
+
 \subsection{``Give examples of the alignment of ORO with OpenCyc''}
@@
 
-\subsection{``the authors are kindly referred to the work by Zender et al.
-(2009, IJCAI) on reference resolution in OWL-DL ontology knowledge bases using
-SPARQL queries''}
+\subsection{``Mention work on Generation and Resolution of Referring Expressions''}
 
-A reference to Zender's work has been added to the \emph{Related work} section.
+Thanks for pointing out the work that has been done in this research community.
+Even if our contribution focuses more on the grounding of concepts from 
+multi-modal knowledge source, these contribution are of definitive 
+interest for us.
+
+A reference to Zender's work has been added to the \emph{Related work} 
+section, along with a reference to Dale and Reiter paper from 1995.
 
@@
 
-\subsection{``the authors are kindly referred to work on the Generation and
-Resolution of Referring Expressions (GRE, RRE) from the Computational
-Linguistics and Natural Language Processing community, e.g. Dale\&Reiter,
-Krahmer\&Theune, Kelleher ''}
-
 \end{document}

### 2011-07-16 paper_knowledge.tex
Addressed remarks concerning curiosity + generation of refering expressions and added relevant papers

@@
 Zender et al.~\cite{Zender2009} address the generation of referring expressions
-(\emph{GRE}) in situated dialogue for topological knowledge.  They consider
+(\emph{GRE},~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
 both the reference resolution and reference description tasks, and rely on
@@
 symbolic knowledge management system. Bio-inspired memory management (the
-ability to forget or reinforce knowledge) or curiosity (the ability to
-identify lacking knowledge and actively trigger behaviors to acquire
-it\fixme{cite Howa}), to give some examples, could arguably be added to the
-list. However, this first analysis sets a convenient reference frame to
-understand and evaluate knowledge representation systems, including the
-\textsc{ORO} knowledge management system we propose.
+ability to forget or reinforce knowledge) or curiosity (the ability to identify
+lacking knowledge and actively trigger behaviours to acquire it -- Hawes et
+al.~\cite{Hawes2011} have contributions in this field, with their \emph{Dora}
+robot which is endowed with motivation mechanisms to explore unknown parts of
+the environment), to give some examples, could arguably be added to the list.
+However, this first analysis sets a convenient reference frame to understand
+and evaluate knowledge representation systems, including the \textsc{ORO}
+knowledge management system we propose.
 

### 2011-07-16 cover_letter.tex
Addressed concerns regarding missing part in the algorithm

@@
 
+\usepackage{pseudocode}
 \usepackage{url}
@@
 
+{\tt GenerateDescription} translates the parsing tree of each nominal group
+into statements. A set of rules (like the one for adjectives mentioned 
+in footnote, section 5.1) help with this task.
+
+The detailled algorithm is provided in appendix to this letter.
+We would be happy to add it to the paper as well if the editor want to consider it.
+
+{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are simple, atomic calls to
+the ontology server. The first one send a SPARQL query build from the set of 
+partial statement build at the {\tt GenerateDescription} step, the second
+one check for each candidate pair if {\tt candidate1 owl:sameAs candidate2}.
+This check uses the classified knowledge model.
+
+The {\tt Discrimation} algorithm is explained in details in Ros2010. The
+reference has been made more explicit.
+
 \subsection{``how does the system deal with restrictive information vs.
@@
 
+\section{Appendix}
+
+\subsection{GenerateDescription algorithms}
+\label{GenerateDescription}
+\begin{pseudocode}[ruled]{GenerateDescription}{group}
+\label{algo|GenerateDescription}
+
+\PROCEDURE{GenerateDescription}{group} 
+   noun \GETS \CALL{GetNoun}{group} \\ 
+   \IF \CALL{Ontology.Lookup}{noun} \in (Instances) \THEN
+   		\BEGIN
+		id \GETS \CALL{Ontology.lookup}{noun}\\	
+		\RETURN {\mathcal{D} + \{ *\ {\tt sameAs}\ <id> \}}\\
+		\END
+   \ELSE
+    	\mathcal{D} = \mathcal{D} + \{ *\ {\tt type}\ <noun>\} \\
+
+   \\
+   det \GETS \CALL{GetDeterminant}{group} \\
+   \IF det \in {\mbox(possessives)} \THEN
+       \mathcal{D} = \mathcal{D} + \{ *\ {\tt isRelatedTo}\ <possessor>\} \\
+
+    \IF det \in {\mbox(demonstratives)} \THEN
+        \BEGIN
+        \IF \CALL{Ontology.Check}{\{<currentSpeaker>\ {\tt focusesOn}\ *\}} \THEN 
+            \mathcal{D} = \mathcal{D} + \{<currentSpeaker>\ {\tt focusesOn}\ *\}
+        \ELSE
+            \mathcal{D} = \mathcal{D} + \CALL{AnaphoricMatching}{} \\
+        \END \\
+   \\
+   adjs \GETS \CALL{GetAdjectives}{group} \\
+   \FOREACH adj \in adjs \DO
+   	\BEGIN
+   		\IF adj == <other> \THEN 
+   			\BEGIN
+   			id \GETS \CALL{History.GetMatchingGroup}{group} \STMTNUM{8em}{st.history}\\
+   			\mathcal{D} = \mathcal{D} + \{ *\ {\tt differentFrom}\ <id> \}\\
+   			\RETURN{D}\\
+			\END   		
+   		\ELSE
+	     	\mathcal{D} = \mathcal{D} + \{ *\ {\tt hasFeature}\ <adj>\} \\
+    \END\\
+
+   \\  
+   nounComplements \GETS \CALL{GetNounComplements}{group} \\
+   \FOREACH nouncmpl \in nounComplements \DO
+     \mathcal{D} = \mathcal{D} + {\CALL{GenerateDescription}{nouncmpl}}\\
+
+   \\  
+   relativeClauses \GETS \CALL{GetSubordinateRelativeClauses}{group} \\
+   \FOREACH relative \in relativeClauses \DO
+   	\BEGIN
+   	 \mathcal{G} \GETS \CALL{GetNominalGroups}{relative} \\
+   	 \FOREACH g \in \mathcal{G} \DO
+     	\mathcal{D} = \mathcal{D} + {\CALL{GenerateDescription}{g}}
+    \END\\
+
+   \\
+   \RETURN{\mathcal{D}} 
+\ENDPROCEDURE
+\end{pseudocode}
+
+The method called at (\ref{st.history}) consists in looking through sentences 
+that have been stored in the conversation history, then extracting their nominal
+group in order to retrieve the identifier of the most recently mentioned 
+concept that holds the same characteristics as the nominal group that is being 
+processed (algorithm~\ref{algo|History}).
+
+\begin{pseudocode}[ruled]{History.GetMatchingGroup}{group}
+\label{algo|History}
+\PROCEDURE{History.GetMatchingGroup}{group}
+\COMMENT{Extract Nominal group from sentences stored in the history}\\
+\mathcal{H} \GETS \CALL{History.GetAllNominalGoup}{}\\
+\COMMENT{Generate description of the nominal group that is being processed.} \\
+\COMMENT{The adjective  "other" is to be removed before calling this routine} \\
+	\mathcal{G} \GETS \CALL{GenerateDescription}{group} \\
+
+	candidates \GETS \mathcal{H} \cap \mathcal{G}\\
+	\IF \left|{candidates}\right| = 0 \THEN
+    \BEGIN
+      \OUTPUT{\mbox{Couldn't find another object with the same characteristics!}} \\
+      \EXIT \\
+    \END
+   \ELSEIF \left|{candidates}\right| = 1 \THEN
+      id \GETS candidates[0]
+   \ELSE
+   	  id \GETS \CALL{Discrimination}{candidates}\\
+   \RETURN{id}
+\ENDPROCEDURE
+\end{pseudocode}
+
 \end{document}

### 2011-07-16 paper_knowledge.tex
Addressed concerns regarding missing part in the algorithm

@@
    \ELSEIF \left|{candidates}\right| = 1 \THEN
-      id \GETS candidates[0] \STMTNUM{8em}{res.easy}\\
+      id \GETS candidates[0]\\
 
@@
 	\ELSE
-	  id \GETS \CALL{Discrimination}{candidates} \\ %\STMTNUM{0em}{st.discrimination}\\
+	  id \GETS \CALL{Discrimination}{candidates} \STMTNUM{0em}{st.discrimination}\\
       \END \\
@@
 to which instance the user is actually referring to. To this end,
-disambiguation routines~\cite{Ros2010b} find differences between the instances 
-(in the example, one banana is yellow while the other one is green) and build a 
-sentence through the \emph{verbalization} module to ask the user a closed 
-question that will help clarify the ambiguity: ``Is it yellow or green?'' The 
-user's answer is parsed and merged with the previous sentence. The resulting, 
-augmented, sentence (``The green banana is good") goes again through all 
-the interpretation steps. This process is repeated until no ambiguities arise. 
+disambiguation routines (algorithm~\ref{algo|Resolution},
+\emph{(\ref{st.discrimination})}, see \cite{Ros2010b} for details of the routines)
+find differences between the instances (in the example, one banana is yellow
+while the other one is green) and build a sentence through the
+\emph{verbalization} module to ask the user a closed question that will
+help clarify the ambiguity: ``Is it yellow or green?'' The user's answer is
+parsed and merged with the previous sentence. The resulting, augmented,
+sentence (``The green banana is good") goes again through all the
+interpretation steps. This process is repeated until no ambiguities arise. 
 In the example, the \concept{banana\_02} is finally returned.

### 2011-07-16 paper_knowledge.tex
Raquel's changes

@@
 
-\subsection{Situated speech acts}
-\label{intro_example}
+%\subsection{Situated speech acts}
+%\label{intro_example}
 
@@
 regarding natural language understanding nor plan-making and manipulation.
-
 To be combined together, those abilities require an unambiguous and shared
@@
 
-Austin~\cite{Austin1962} would have at first glance analyzed this kind of
-sentence as a \emph{speech act}, comprising of \emph{locutionary},
-\emph{illocutionary} and possibly \emph{perlocutionary} acts. First, we want to
-understand the direct meaning of the sentence (\emph{locutionary act}): we must
-acquire the sentence, convert it into a useful syntactic form (probably through
+The first step is to understand the meaning of the sentence. To this end, we must acquire the sentence, convert it into a useful syntactic form (probably through
 speech recognition), and understand the semantics of the sentence, \ie What is
@@
 
-Working in a situated context, we want furthermore to \emph{resolve} these
+Working in a situated context, we need to \emph{resolve} these
 semantics atoms, \ie ground them in the sensory-motor space of the robot. For
@@
 previously mentioned concept.
+
+The next step is to extract and understand the \emph{intent} of the utterance
+as thought by the agent.
+In our example, Thomas obviously wants an action to be performed by the robot.
+The action parametrization is conveyed by the semantics attached to the words and the grammatical structure of the sentence.
+In our example, the type of action is
+given by the verb ``\textit{give}''. Assuming the robot has some procedural
+knowledge attached to this symbol, the action type can be considered as
+grounded for the robot. We can as well understand that the recipient of the
+action is the human, the performer is the robot itself, and the object acted
+upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
+that can be extracted from the sentence that allow for fully grounding the action.\footnote{This analysis has been inspired on the work of Austin et al.~\cite{Austin1962}, where this type of sentences correspond to \emph{speech acts}, comprising of \emph{locutionary act} (the meaning), \emph{illocutionary} (the intent) and possibly \emph{perlocutionary acts} (implicit speaker's expectation).}
+
+%\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deal
+%with implicit speaker's expectations. We will not consider it further in this
+%article}.
+%Austin~\cite{Austin1962} would have at first glance analyzed this kind of
+%sentence as a \emph{speech act}, comprising of \emph{locutionary},
+%\emph{illocutionary} and possibly \emph{perlocutionary} acts. 
+%Firstly, we want to understand the direct meaning of the sentence (\emph{locutionary act}): 
+%Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
+%as thought by the agent, must be extracted and understood. 
+%Finally, the implicit speaker's expectation can sometimes be conveyed --the
+%\emph{perlocutionary act}-- and should be taken into account. If Thomas would
+%have said instead ``Jido, this box is too heavy for me'', he may implicitly
+%expect the robot to memorize the fact that the box's weight is over his
+%capabilities, and it may as well imply that Thomas is requiring help to carry
+%the box.
 
@@
 
-Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
-as thought by the agent, must be extracted and understood. In our example,
-Thomas obviously wants an action to be performed by the robot. The action
-parametrization is conveyed by the semantics attached to the words and the
-grammatical structure of the sentence. In our example, the type of action is
-given by the verb ``\textit{give}''. Assuming the robot has some procedural
-knowledge attached to this symbol, the action type can be considered as
-grounded for the robot. We can as well understand that the recipient of the
-action is the human, the performer is the robot itself, and the object acted
-upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
-that can be extracted from the sentence that allow for fully grounding the action
-\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deal
-with implicit speaker's expectations. We will not consider it further in this
-article}.
-
-%Finally, the implicit speaker's expectation can sometimes be conveyed --the
-%\emph{perlocutionary act}-- and should be taken into account. If Thomas would
-%have said instead ``Jido, this box is too heavy for me'', he may implicitly
-%expect the robot to memorize the fact that the box's weight is over his
-%capabilities, and it may as well imply that Thomas is requiring help to carry
-%the box.
-
 \subsection{Approach}
 
-Extracting these speech acts and turning them into content processable by the
-robot is a difficult challenge in the general case. We base our approach on
-three distinct, inter-related cognitive functions.
+In this work we propose an approach aiming at grounding the interaction with users, verbal and non-verbal, following the steps described above.
+To this end, we have developed three distinct, inter-related cognitive components, which we believe fulfill the minimum requisites for situated human interaction.
 Figure~\ref{fig|overall-process} illustrates the general context, while
-Figure~\ref{fig|main-components} summarizes the interactions between the three
-main components:
+Figure~\ref{fig|main-components} summarizes the interactions between these three
+main components.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Technical analysis}
-\label{examples}
-
-In order to better understand the overall process we next describe the
-different steps of the approach based on three examples. In these examples, we
+%\section{Technical analysis}
+%\label{examples}
+
+In order to better understand the overall process the {\sc Dialogs} module and its relation with ORO, we next describe the different steps of the approach based on three examples.
+In these examples, we
 assume that some initial facts are present in the knowledge base, both in the
@@
 statements that describe the concept is generated based on the grammatical
-parsing of the sentence (algorithm~\ref{algo|Resolution},
-\emph{(\ref{res.desc})}). In the example, a banana (\stmt{?obj type Banana})
+parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). In the example, a banana (\stmt{?obj type Banana})
 that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
@@
 query is sent to the ontology server to retrieve possible instances that match
-the description (algorithm~\ref{algo|Resolution}, \emph{(\ref{res.onto})}).
+the description (algorithm~\ref{algo|Resolution}(\ref{res.onto})).
 
@@
 to which instance the user is actually referring to. To this end,
-disambiguation routines (algorithm~\ref{algo|Resolution},
-\emph{(\ref{st.discrimination})}, see \cite{Ros2010b} for details of the routines)
+disambiguation routines (algorithm~\ref{algo|Resolution}(\ref{st.discrimination}), see \cite{Ros2010b} for details of the routines)
 find differences between the instances (in the example, one banana is yellow
@@
 
-\subsection{``Moving to London'' scenario}
+\subsection{\emph{Moving to London} scenario}
 
@@
 SPARK.
-
 The initial information in ORO corresponds to:
@@
 \vspace{0.10cm}
-
 	SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:

### 2011-07-17 paper_knowledge.tex
Raquel: Fixed typos/grammar
(Raquel, I prefer to keep with UK spelling when possible, thus 'behaviour' and not 'behavior' for instance)

@@
 ability to forget or reinforce knowledge) or curiosity (the ability to identify
-lacking knowledge and actively trigger behaviours to acquire it -- Hawes et
-al.~\cite{Hawes2011} have contributions in this field, with their \emph{Dora}
-robot which is endowed with motivation mechanisms to explore unknown parts of
-the environment), to give some examples, could arguably be added to the list.
+lacking knowledge and actively trigger behaviours to acquire it --Hawes et
+al.~\cite{Hawes2011} have contributed in this area with \emph{Dora}, a
+robot endowed with motivation mechanisms to explore unknown regions of
+the environment--), to give some examples, could arguably be added to the list.
 However, this first analysis sets a convenient reference frame to understand
@@
 Knowledge is represented in ORO in the Description Logics formalism (using the
-OWL-DL -- \emph{Web Ontology Language - Description Logics} -- language), as
+OWL-DL --\emph{Web Ontology Language - Description Logics}-- language), as
 RDF (\emph{Resource Description Framework}) triples (for instance \stmt{robot
@@
 statements are continuously reasoned about (\emph{classified}), and at any time,
-the server expose a complete set of statements, both the asserted ones and the
+the server exposes a complete set of statements, both the asserted ones and the
 inferred ones. For instance, if \stmt{socrates type Human} and \stmt{Human
@@
 relation is particularly useful during interaction when one of the agents is
-referring to an object saying ``this" or ``that" while pointing at it.
+referring to an object saying ``this" or ``that" while pointing at it. Section~\ref{disamb-gestures} describes in more detail the combination of both sources of information (verbal and non-verbal).
 
@@
 to reach an object, which is fundamental for task planning. For example, if the
-user asks the robot to give her an object, the robot must compute a transfer
+user asks the robot to give him/her an object, the robot must compute a transfer
 point where the user is able to get the object afterward.
@@
 grounding (\eg \cite{Regier2001} presented different models to define the
-\emph{above} relation; \cite{Kelleher2009} presented two computational models of topological and projective spatial prepositions; \cite{Blisard2005} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
+\emph{above} relation; \cite{Kelleher2009} introduced two computational models of topological and projective spatial prepositions; \cite{Blisard2005} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
 In this work we use similar models based on the
@@
 
-In order to better understand the overall process the {\sc Dialogs} module and its relation with ORO, we next describe the different steps of the approach based on three examples.
+In order to better understand the overall process of the {\sc Dialogs} module and its relation with ORO, we next describe the different steps of the approach based on three examples.
 In these examples, we

### 2011-07-17 cover_letter.tex
Raquel: polishing the rebuttal letter

@@
 \usepackage{url}
+
+\newcommand{\stmt}[1]{{\footnotesize $\langle$\stmttt#1\relax$\rangle$}}
 
@@
 
-The authors would like to thank the reviewers for their constructive feedback that helped us to strengthen the paper.
+The authors would like to thank the reviewers for their constructive feedback
+that helped us to strengthen the paper.
 
-Below follow our answers to the reviewersÕ comments and suggestions.
-Each answer also indicates the changes made in the paper to address the reviewerÕs concern.
-An important number of language corrections are made, and the paper is proof-read by a native English speaker.
+We next answer the reviewers comments and suggestions. Besides, each answer
+indicates the changes made in the paper to address the reviewers' concerns. An
+important number of language corrections have been made, and the paper has been
+proof-read by a native English speaker.
 
@@
 with people) should be allowed to make mistakes due to perception noise, or not
-to be sure about something (for instance, "I think that..."), which represents
+to be sure about something (for instance, ``I think that..."), which represents
 the uncertainty the robot has about a fact. In the meantime, to overcome
-perception errors, the robot provides visual feedback (looking at objects) to
-the user when possible and/or verifies the information with the user whenever
-is needed (for instance, "do you mean the box\_01"?).
+perception errors, the robot provides visual feedback (looking at objects,
+showing in screen its 3D model of the environment) to the user when possible
+and/or verifies the information with the user whenever is needed (for instance,
+``do you mean the box\_01"?).
 
@@
 
-Due to the perception problems and computational delays, fully realistic
-settings are not yet possible.  The interaction based only on a static setting,
-i.e. objects don't move, the interaction with a real user is accurate enough.
-We have developed some experiments, where the user interacts with the system
-only through a keyboard (no gestures are included) asking questions, providing
-information, etc. and the outcome is quite promising. However, for dynamic
-settings where the user can point at things, move the objects around, etc., the
-interaction is not fast enough and the user should perform slow movements for
-the robot to create a consistent state of world.
+Due to perception problems and computational delays, fully realistic settings
+are not yet possible. In static settings, where neither humans nor objects
+move, the interaction with real users is meaningful and the flow of information
+between robot and user is accurate enough.  We have developed some experiments
+where the user interacts with the system only through a keyboard (no gestures
+are included) asking questions, providing information, etc. and the outcome is
+quite promising. However, for dynamic settings, where the user can point at
+things, move the objects around, etc., the interaction is not fast enough and
+the user should perform slow movements for the robot to create a consistent
+state of world.
 
@@
 
-Reviewer comment: ``The paper starts off with a discussion
-of Austins speech act theory, including locutionary act, illocutionary force
-and perlocutionary act. As the concept of speech acts is not really picked up
-in the remainder of the paper, I do not see the point in starting with such a
-discussion. It would be more appropriate to motivate how a robot could benefit
-from a symbolic knowledge base etc.''
+%Reviewer comment: ``The paper starts off with a discussion
+%of Austins speech act theory, including locutionary act, illocutionary force
+%and perlocutionary act. As the concept of speech acts is not really picked up
+%in the remainder of the paper, I do not see the point in starting with such a
+%discussion. It would be more appropriate to motivate how a robot could benefit
+%from a symbolic knowledge base etc.''
 
-Austin's \emph{speech acts} were actually one of the starting point of our
-analysis of dialogue understanding, and we think it is worthwhile to briefly mention
-them in the introduction, as one of the theoritical ground of our work.
+The analysis of dialogue understanding we have followed has been inspired by
+Austin's \emph{speech acts}. Thus, we believe it is worthwhile to briefly mention
+them in the introduction as a theoretical ground of our work.
 
-The reviewer however rightfully notes that \emph{perlocutionnary acts}, while mentioned, are not
-actually used further. We moved this paragraph to a simple footnote.
+However, since we do not directly refer to them in the remainder of the paper,
+we refer to Austin's definitions in a footnote.
 
@@
 \begin{itemize}
-\item All reviewer comments regarding citations have been addressed, following reviewer's suggestions,
-\item The reviewer's remark on human tendency to explain in a consistent way the world has been added
+\item All reviewer's comments regarding citations have been addressed.
+\item The reviewer's remark on human tendency to explain in a consistent way the world has been added.
 \item Language issues have been addressed as well.
@@
 
-The Open World assumption mainly allows us to model that we may not know
-if a fact is true or not.
-
-One very common example is the {\tt knowsAbout} predicate. This predicate
-defines if some agent knows about some object.
-
-It is important (especially for dialog) to be able to equally model that an agent knows or 
-does not know about the object, or that the robot simply does not know if 
-the human knows about it.
+The Open World assumption mainly allows us to model that we may not know if a
+fact is true or not. A common example is the {\tt knowsAbout} predicate. This
+predicate defines if some agent knows about some object. It is important
+(especially for dialog) to model that an agent knows or  does not know about an
+object, or that the robot simply does not know if the human knows about it.
 
 The Open World Assumption has important consequences on the reasoning
-abilities of the system, since it prevents the classical "negation by 
-failure" and leads to tighter modelling requirement to locally "close the
+abilities of the system, since it prevents the classical ``negation by 
+failure" and leads to tighter modelling requirement to locally ``close the
 world" (for instance, the {\tt Color} class is defined as the equivalent of
@@
 
-The paragraph 2.2 has been reorganized to clarify the vocabulary and better
+Paragraph 2.2 has been reorganized to clarify the vocabulary and better
 define the various acronyms.
@@
 
-A mention to the Dora robot has been added in section 2.1.
+A reference to the Dora robot has been added in section 2.1. However, at the
+time when this work had been submitted for review, neither of the two proposed
+papers by the reviewer were available (and in fact, IJCAI is taking place this
+week).
 
@@
 
-Numerous predicates in {\tt commonsense.oro.owl} are defined in the 
-OpenCyc namespace: {\tt farFrom, near, hold-Underspecified, objectFoundInLocation}.
-
-Similarly, most of the upper classes in the ontology belong to OpenCyc 
-(like {\tt SpatialThing, PartiallyTangible, TemporalThing}...).
-
-Overall, about 50 entites (on the >180 defined in {\tt commonsense.oro.owl}) 
-come from OpenCyc.
+Numerous predicates in {\tt commonsense.oro.owl} are defined in the OpenCyc
+namespace: {\tt farFrom, near, hold-Underspecified, objectFoundInLocation} are
+some of examples. Similarly, most of the upper classes in our ontology belong
+to OpenCyc (e.g. {\tt SpatialThing, PartiallyTangible, TemporalThing},...).
+Overall, about 50 entities (from the more than 180 defined in {\tt
+commonsense.oro.owl}) come from OpenCyc.
 
 Since most of the upper classes are OpenCyc concepts, this should 
-guarantee to a certain extend the alignment of our ontology with other ontologies.
+guarantee, to a certain extend, the alignment of our ontology with other ontologies.
 
@@
 
-\subsection{``p.6, II.E. Explain "multi-lingual support"''}
+\subsection{``p.6, II.E. Explain ``multi-lingual support"''}
 
-The "multi-lingual support" has been reworded:
-
-``fast concept lookup, with possible multi-lingual support (through the
-{\tt @lang} XML annotation, labels of concept can be translated, and
-specific translations can be queried for)''
+The ``multi-lingual support" part has been reworded (Section 2.5.1).
 
@@
 
-PointsAt relation is computed continuously during the interaction by 
-selecting the objects that "might" be pointed at by the human. The outcome
-of the computation (facts like "human1 pointsAt bottle1") are then
-continously updated in the knowledge base.
+SPARK is continuously updating the human non-verbal queues included in this
+work (i.e., looking at, pointing at, etc.). The outcome of these computation
+(facts like {\tt $\langle$human1 pointsAt bottle\_01$\rangle$}) are then
+constantly been sent to the knowledge base.
 
-In the current system, the decision on whether the human is pointing at an 
-object in a given time is left to the reasoning on the dialog or on the human gestures:
-when the dialogue module analyses a sentence like "give me this", it tries to
-resolve the atom "this" by querying the ontology.
+In the current approach, the timing aspect is managed during the analysis of
+dialogue. When the user makes reference to an object through demonstratives
+(e.g. this, that), the DIALOGS module queries the ontology for objects being
+pointed at. The assumption for this mechanism to work is that the human is
+constantly pointing while the dialog analysis takes place.  Both situation
+assessment and dialogue processing chains induce small delays, but these are
+roughly the same (below 200ms), and compensate for each other.
 
-Both sitatuation assessement and dialogue processing chains induce 
-small delays, but these are roughly the same (below 200ms), and 
-compensate for each other.
+Section 3.1. and 5.3.3. clarify this point.
 
@@
 
-Kelleher's work, as well as Blisard's, are now correctly cited in the section.
+Kelleher's work, as well as Blisard's, are now cited in the section.
 
@@
 
-We thank the reviewer for this input that has been added to the article.
+We thank the reviewer for this input, which has been added to the article.
 
 \subsection{``p.9, Alg. IV.1: the crucial bits are left out; please explain
-"GenerateDescription", "Ontology.Find", "Ontology.CheckEquivalent", and
-"Discrimination"''}
+``GenerateDescription", ``Ontology.Find", ``Ontology.CheckEquivalent", and
+``Discrimination"''}
 
@@
 
-The detailled algorithm is provided in appendix to this letter.
-We would be happy to add it to the paper as well if the editor want to consider it.
+The detailled algorithm is provided in appendix to this letter.  We would be
+happy to add it to the paper as well if the editor wants to consider it.
 

### 2011-07-17 cover_letter.tex
Addressed question on 'restrictive information'

@@
 \subsection{``how does the system deal with restrictive information vs.
-attributive information expressed by adjectives (e.g., in "the yellow banana is
+attributive information expressed by adjectives (e.g., in ``the yellow banana is
 big")?''}
+
+When sentences like ``the yellow banana is big" are processed, they are
+recognized as statements to be added to the ontology because of the state verb ``to be''
+({\sc Dialogs} maintains a list of such verbs).
+
+The nominal group is first resolved as usual (``the yellow banana''), and a new statement
+is then generated that describe the new feature to be added to the concept.
+
+Please refer to the {\sc GenerateDescription} algorithm in appendix to see how the
+new statements are created.
 

### 2011-07-17 paper_knowledge.tex
Commented out the anecdote on light detection error (naming experiment)
This to address reviewer 3 comment and shorter the paper

@@
 
-An unexpected example of what the symbolic reasoning layer brings to more
-traditional robotic architectures emerged during the \emph{Naming} experiment:
-the perception routines provided segmented blobs corresponding to objects along
-with their colors to update the information in ORO. At some point, ORO suddenly
-refused to add an object. What seemed at first a communication bug between
-modules, was actually the consequence of a consistency check by ORO: because of
-bad light conditions, the color recognition was not very reliable, and the same
-object was set to have two different colors at the same time. Since the
-\concept{hasColor} predicate we had decided to use was a functional predicate
-(\ie objects can have only one color), ORO inferred an inconsistency and
-discarded the new fact. This kind of logical failure can be used to improve
-low-level perception outcomes by ``closing the loop'' with high-level, symbolic
-knowledge.
+% An unexpected example of what the symbolic reasoning layer brings to more
+% traditional robotic architectures emerged during the \emph{Naming} experiment:
+% the perception routines provided segmented blobs corresponding to objects along
+% with their colors to update the information in ORO. At some point, ORO suddenly
+% refused to add an object. What seemed at first a communication bug between
+% modules, was actually the consequence of a consistency check by ORO: because of
+% bad light conditions, the color recognition was not very reliable, and the same
+% object was set to have two different colors at the same time. Since the
+% \concept{hasColor} predicate we had decided to use was a functional predicate
+% (\ie objects can have only one color), ORO inferred an inconsistency and
+% discarded the new fact. This kind of logical failure can be used to improve
+% low-level perception outcomes by ``closing the loop'' with high-level, symbolic
+% knowledge.
 

### 2011-07-17 paper_knowledge.tex
Shorten 'article overview' + formatting + missing references

@@
 \usepackage{alltt}
-
-% fixme notes
-\usepackage[draft,footnote,marginclue]{fixme}
 
@@
 \textit{ResearchCyc} ontology (including the \textit{MicroTheories} concept),
-used in combination with the CycL language. This enables second-order
+used in combination with the {\sc CycL} language. This enables second-order
 logic modeling and access to a large common-sense knowledge base.
@@
 
-In the next section we present in depth the so-called {\sc ORO}
-(\emph{OpenRobots Ontology} server) knowledge base, a dynamic knowledge
-representation system based on Description Logics ontologies and relying on the
-\emph{Open World Assumption}. We present it first, along with its objectives,
+In the next section we present the {\sc ORO}
+(\emph{OpenRobots Ontology} server) knowledge base. We present it first, along with its objectives,
 since it is the knowledge \textit{hub} of the system, used pervasively by other
@@
 
-Section 4 covers in detail the dialogue processing component. The main
-algorithms for natural language grounding are described, as well as the
-interactions between these component and the knowledge base.
+Section 4 covers in detail the dialogue processing component.
 
@@
 \emph{statement} or an \emph{assertion} about the world\footnote{In this paper,
-statements are always triples \texttt{<subject, predicate, object>}, following
-the Description Logics formalism.}. In practical terms, knowledge is made of
+statements are always triples \stmt{subject predicate object}, \ie binary
+relations between entities.}. In practical terms, knowledge is made of
 statements that are \emph{contextualized}, if possible \emph{synthesized}, and
@@
 exchange with other robots (for instance, the works previously mentioned by
-Daoutis and Tenorth rely on the same Cyc concepts\fixme{Give
-examples}).
+Daoutis and Tenorth rely on the same Cyc concepts).
 
@@
 
-\item \emph{Points At}: verifies whether an object is pointed at by an agent. This
-relation is particularly useful during interaction when one of the agents is
-referring to an object saying ``this" or ``that" while pointing at it. Section~\ref{disamb-gestures} describes in more detail the combination of both sources of information (verbal and non-verbal).
+\item \emph{Points At}: verifies whether an object is pointed at by an agent.
+This relation is particularly useful during interaction when one of the agents
+is referring to an object saying ``this" or ``that" while pointing at it.
+Section~\ref{disamb-gestures} describes in more detail the combination of both
+sources of information (verbal and non-verbal).
 
@@
 is grounded (to extract the relation \concept{isOn} and to find what \emph{table}
-refers to) and transformed into the following kind of query: {\tt find ?var
+refers to) and transformed into the following kind of query: \concept{find ?var
 [\stmt{?var isOn table1}]}.  Answers are converted back to a full sentence
@@
 \subsubsection{Explicit disambiguation through verbal interaction and gestures}
+\label{disamb-gestures}
+
 \begin{figure}[!ht]

### 2011-07-17 cover_letter.tex
Commented out the anecdote on light detection error (naming experiment)
This to address reviewer 3 comment and shorter the paper
Added a note in the cover letter regarding the article length
Raquel's comments

@@
 
-\title{Responses to reviewers' comments and list of modifications done\\ on paper SORO226\\
+\title{Answers to reviewers' comments and list of modifications done\\ on paper SORO226\\
 \LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
@@
 
-The authors would like to thank the reviewers for their constructive feedback
-that helped us to strengthen the paper.
+The authors would like to thank the reviewers for having accepted to review our work and their constructive feedback, which helped us to strengthen the paper.
 
@@
 indicates the changes made in the paper to address the reviewers' concerns. An
-important number of language corrections have been made, and the paper has been
-proof-read by a native English speaker.
+important number of language corrections have been made with the collaboration of a native English speaker.
+
+Please also note that the overall length of the paper has actually 
+increased due to the use of the official Springer \LaTeX template.
 
@@
 Austin's \emph{speech acts}. Thus, we believe it is worthwhile to briefly mention
-them in the introduction as a theoretical ground of our work.
-
-However, since we do not directly refer to them in the remainder of the paper,
+them in the introduction as a theoretical ground of our work. However, since we do not directly refer to them in the remainder of the paper,
 we refer to Austin's definitions in a footnote.
@@
 A reference to the Dora robot has been added in section 2.1. However, at the
-time when this work had been submitted for review, neither of the two proposed
-papers by the reviewer were available (and in fact, IJCAI is taking place this
+time when this work had been submitted for review, neither of the two papers proposed
+by the reviewer were available (and in fact, IJCAI is taking place this
 week).
@@
 In the current approach, the timing aspect is managed during the analysis of
-dialogue. When the user makes reference to an object through demonstratives
+dialogue. When the user makes reference to an object through demonstrative pronouns
 (e.g. this, that), the DIALOGS module queries the ontology for objects being
@@
 
-\subsection{``p.8, III.B. "creation of imaginary objects...hinted by verbal
-assertions..." -- this phenomenon is called presupposition accommodation''}
+\subsection{``p.8, III.B. creation of imaginary objects...hinted by verbal
+assertions... -- this phenomenon is called presupposition accommodation''}
 
@@
 into statements. A set of rules (like the one for adjectives mentioned 
-in footnote, section 5.1) help with this task.
+in footnote, section 5.1) help with this task. The detailed algorithm is provided in appendix to this letter.  We would be happy to add it to the paper as well if the editor wants to consider it.
 
-The detailled algorithm is provided in appendix to this letter.  We would be
-happy to add it to the paper as well if the editor wants to consider it.
-
-{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are simple, atomic calls to
-the ontology server. The first one send a SPARQL query build from the set of 
-partial statement build at the {\tt GenerateDescription} step, the second
-one check for each candidate pair if {\tt candidate1 owl:sameAs candidate2}.
+{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are atomic calls to
+the ontology server. The first one sends a SPARQL query based on the set of 
+partial statements built at the {\tt GenerateDescription} step. The second
+one checks if {\tt $\langle$candidate1 owl:sameAs candidate2$\rangle$} for each candidate pair.
 This check uses the classified knowledge model.
 
-The {\tt Discrimation} algorithm is explained in details in Ros2010. The
+The {\tt Discrimation} algorithm is explained in detail in Ros2010. The
 reference has been made more explicit.
@@
 ({\sc Dialogs} maintains a list of such verbs).
-
-The nominal group is first resolved as usual (``the yellow banana''), and a new statement
-is then generated that describe the new feature to be added to the concept.
-
-Please refer to the {\sc GenerateDescription} algorithm in appendix to see how the
+The nominal group is first resolved as usual (``the yellow banana''), and a new statement is then generated describing the new feature to be added to the concept.
+Please refer to the {\sc GenerateDescription} algorithm in the appendix to see how the
 new statements are created.
@@
 Thanks for pointing out the work that has been done in this research community.
-Even if our contribution focuses more on the grounding of concepts from 
-multi-modal knowledge source, these contribution are of definitive 
-interest for us.
-
+Even if our current work focuses more on the grounding of concepts from 
+multi-modal knowledge sources, these contributions are of great interest to us for future work.
 A reference to Zender's work has been added to the \emph{Related work} 
-section, along with a reference to Dale and Reiter paper from 1995.
+section, along with a reference to the Dale and Reiter paper from 1995.
 
@@
 
-Paragraph added in the section 6 introduction.
+A brief description of both platforms has been included in the introduction of Section 5.
 
@@
 
-The study was conducted with TUM in Munich. This has been corrected in 
-the paper as well.
+The study was conducted with TUM in Munich. This information has been included in the paper.
 
@@
 or give some more context''}
+
+To avoid increasing the length of the paper while providing more context about this anecdote, we have finally removed it.
 
@@
 The method called at (\ref{st.history}) consists in looking through sentences 
-that have been stored in the conversation history, then extracting their nominal
-group in order to retrieve the identifier of the most recently mentioned 
+that have been stored in the conversation history and extracting their nominal
+groups. From these we retrieve the identifier of the most recently mentioned 
 concept that holds the same characteristics as the nominal group that is being 
@@
 \COMMENT{Generate description of the nominal group that is being processed.} \\
-\COMMENT{The adjective  "other" is to be removed before calling this routine} \\
+\COMMENT{The adjective  ``other" is to be removed before calling this routine} \\
 	\mathcal{G} \GETS \CALL{GenerateDescription}{group} \\

### 2011-07-17 paper_knowledge.tex
Raquel's comments

@@
 the anchoring process itself relies on the situated nature of human-robot
-interactions. We present an implementation, including a specialized symbolic
+interactions. We present an integrated approach, including a specialized symbolic
 knowledge representation approach based on Description Logics, and case studies
@@
 moving out and is packing everything with the help of Jido, his robot
-(Fig.~\ref{fig|vpt}).
-
-``Jido, give me that'', says Thomas while looking at a box that contains a
+(Fig.~\ref{fig|vpt}). ``Jido, give me that'', says Thomas while looking at a box that contains a
 video tape. The robot smoothly grasps the tape, and hands it to the him.
@@
 Zender et al.~\cite{Zender2009} address the generation of referring expressions
-(\emph{GRE},~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
+(\emph{GRE}~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
 both the reference resolution and reference description tasks, and rely on
@@
 	\centering
-	\includegraphics[width=\columnwidth]{images/reach.jpg} 
+	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
 	\caption{Different reaching postures for the human: the bottle and the 
@@
 statements that describe the concept is generated based on the grammatical
-parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). In the example, a banana (\stmt{?obj type Banana})
+parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). The parsed tree of each nominal group is translated into statements based on a set of rules.
+In the example, a banana (\stmt{?obj type Banana})
 that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
@@
 predicate if the adjective is not known.}.  Based on these partial statements a
-query is sent to the ontology server to retrieve possible instances that match
+SPARQL query is sent to the ontology server to retrieve possible instances that match
 the description (algorithm~\ref{algo|Resolution}(\ref{res.onto})).
@@
 This section presents several experiments where the different features of our
-approach are represented.
-
-Experiments have been conducted on two different platforms: the
+approach are represented. The experiments have been conducted on two different platforms: the
 \emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
 holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
-System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from
-LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
-Genom/pocolibs} stack.
+System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
+Genom/pocolibs} stack. Both platforms share the use of ORO, the ontology server proposed in this work.
 

### 2011-07-17 paper_knowledge.tex
Added biographies

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\section*{Biographies}
+
+\subsection*{Séverin Lemaignan}
+
+Séverin Lemaignan received the M.S. degree in mechanical engineering from the
+École des Arts et Métiers, Paris, France and the Technical University of
+Karlsruhe, Germany.  After working for a year as research engineer in mobile
+robotics at INRIA, he is now Ph.D student at LAAS-CNRS, Toulouse, France and
+Munich Technical University, Germany. He mainly works on knowledge
+representation for cognitive robotic architectures, natural human-robot
+interaction and high-level robotic simulation.
+
+\subsection*{Raquel Ros}
+
+Raquel Ros is a Marie-Curie fellow at the Robotics and Interactions Group in
+LAAS-CNRS, where she focuses on implementation of cognitive models for social
+robots, knowledge grounding and enhancement of human-robot interaction. She
+received her PhD in computer science from the Universitat Autònoma de Barcelona
+in 2008. She was then awarded with a Marie-Curie fellowship and joined
+LAAS-CNRS. Her main research interests are decision-making, cognitive models
+for social robotics and long-term interaction.
+
+\subsection*{E. Akin Sisbot}
+
+Emrah Akin Sisbot received the B.S. degree in computer engineering from
+Galatasaray University, Istanbul, Turkey, in 2003 and the M.S. degree in
+artificial intelligence from the University Paul Sabatier, Toulouse, France, in
+2004. In 2008 he received his Ph.D degree from University of Toulouse with his
+research on Human-Aware Motion Planning. Until 2011 he was a Postdoc in
+Robotics and Interactions Group in LAAS-CNRS where he focused on Spatial
+Reasoning and Knowledge Management. He recently joined to University of
+Washington as a Research Associate, where he works on Interactive RGB-D SLAM
+and robot navigation. His research interests include motion planning,
+human-robot interaction, spatial reasoning and mapping.
+
+\subsection*{Rachid Alami}
+
+Rachid Alami received a Docteur-Ingénieur degree in 1983 in computer science
+from the Institut National Polytechnique de Toulouse and the ``Habilitation à
+Diriger des Recherches'' in 1996 from Paul Sabatier University. He joined the
+Robotics and Artificial Intelligence group at LAAS-CNRS, Toulouse, France in
+1980. He is now a Senior Research Scientist and the head of Robotics and
+InteractionS group at LAAS-CNRS. His current research interests include the
+study of control architectures for autonomous robots, task and motion planning,
+multi-robot cooperation, personal robots, and human-robot interaction.
+
+\subsection*{Michael Beetz}
+
+Michael Beetz received his diploma degree in Computer Science with distinction
+from the University of Kaiserslautern. He received his MSc, MPhil, and PhD
+degrees from Yale University in 1993, 1994, and 1996 and his Venia Legendi from
+the University of Bonn in 2000. Michael Beetz was a member of the steering
+committee of the European network of excellence in AI planning (PLANET) and
+coordinating the research area ``robot planning''. He is associate editor of
+the AI Journal. His research interests include plan-based control of robotic
+agents, knowledge processing and representation for robots, integrated robot
+learning, and cognitive perception.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \end{document}

### 2011-07-17 cover_letter.tex
cover letter: improved section 3.1

@@
 abilities of the system, since it prevents the classical ``negation by 
-failure" and leads to tighter modelling requirement to locally ``close the
+failure". Lack of negation by failure leads to tighter modelling requirement to locally ``close the
 world" (for instance, the {\tt Color} class is defined as the equivalent of
-the union of the 10 or so color instances we consider).
+the union of 10 color instances).
 

### 2011-07-17 cover_letter.tex
Fixed wrong section reference in cover letter
Removed link to YouTube video

@@
 state of world.
-
-A recent video showing the current state of the interaction can be viewed
-here: \url{http://www.youtube.com/watch?v=IODx50uV_k4}
 
@@
 into statements. A set of rules (like the one for adjectives mentioned 
-in footnote, section 5.1) help with this task. The detailed algorithm is provided in appendix to this letter.  We would be happy to add it to the paper as well if the editor wants to consider it.
+in footnote, section 4.1) help with this task. The detailed algorithm is provided in appendix to this letter.  We would be happy to add it to the paper as well if the editor wants to consider it.
 

### 2011-07-18 paper_knowledge.tex
Last round of Raquel changes

@@
 model is realistic in the sense that it relies on accurate 3D models of both
-manipulated objects and humans. It also has dedicated mechanisms to manage
+manipulable objects and humans. It also has dedicated mechanisms to manage
 disappearing or occluded objects.  The geometric model is used to compute
@@
 Zender et al.~\cite{Zender2009} address the generation of referring expressions
-(\emph{GRE}~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
+(GRE~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
 both the reference resolution and reference description tasks, and rely on
@@
 These points highlight some original aspects of a larger cognitive architecture
-that has been deployed and tested on three distinct robotic platforms
+that has been deployed and tested on several robotic platforms
 (including both humanoid robots and service robots), demonstrating the
@@
 
-\subsection{Article overview}
+%\subsection{Article overview}
 
@@
 components.
-
-Section 3 presents {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge}), a
+Section~\ref{spark} presents {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge}), a
 module that merges perceptual information with a coherent geometric model and
@@
 environment. 
-
-Section 4 covers in detail the dialogue processing component.
-
-Section 5 presents three use-cases that were conducted on two different
-robotic platforms: the \emph{Naming} experiment, where a robot anchors new
+Section~\ref{dialog} covers in detail the dialogue processing component and its 
+relation with ORO.
+Section~\ref{experiments} presents three use-cases that were conducted on two 
+different robotic platforms: the \emph{Naming} experiment, where a robot anchors new
 knowledge in its model through verbal interaction; the \emph{Spy Game}, where
@@
 this knowledge. 
+Finally, Section~\ref{conclusion} concludes and discusses the work presented 
+in this paper.
 
@@
 statements that are \emph{contextualized}, if possible \emph{synthesized}, and
-\emph{limited} to a domain of validity.
-
-These three features have important consequences for the way a knowledge
-representation and storage system must be designed. Let us examine them:
+\emph{limited} to a domain of validity. These three features have important 
+consequences for the way a knowledge representation and storage system must 
+be designed. Let us examine them:
 
@@
 
-Note that the focus of the ORO platform is on enabling such context to be
+Note that the focus of ORO is on enabling such context to be
 effectively represented rather than actually identifying the current context.
@@
 
-\paragraph{Domain of validity}: its main role is to specify the scope in which
+\paragraph{Domain of validity} specifies the scope in which
  information is (believed to be) true. It covers several aspects: temporal,
@@
 assumed to be shared by all agents.
-
 The ORO server can be loaded with an initial set of statements which we call
@@
 
-As previously mentioned, ontologies in ORO are written in OWL, a language
-based on Description Logics (a decidable subset of first-order logic). The
-Pellet reasoner supports most of the OWL constructs and allows several type of
+As previously mentioned, ontologies in ORO are written in OWL. The
+Pellet reasoner supports most of the OWL constructs and allows several types of
 reasoning:
@@
 new constraints and properties to classes and predicates.
-
 The \emph{Naming} experiment (section~\ref{naming}) gives a simple example of
@@
 exploitation) of a consistent knowledge model, the \emph{Common Sense Ontology}
-helps build assertions that are related to previous one by offering
+helps build assertions that are related to previous ones by offering
 a predefined vocabulary.
@@
 the belief state of its interactors, opening new possibilities for the control
-architecture. In section~\ref{dialog} we present how we use this feature to
+architecture. In section~\ref{exp:disambVPT} we present an example of how we use this feature to
 make sense of user sentences from his/her point of view. Moreover,
@@
 We have implemented several algorithms (common ancestors, computation of the
-best discriminant, see~\cite{Ros2010b}) to help the robot cluster a set of
-concepts based on their symbolic similarities (common properties, common
-ancestors). One particular application of these functions is discrimination.
+best discriminant~\cite{Ros2010b}) to help the robot cluster a set of
+concepts based on their symbolic similarities. One particular application of these functions is discrimination.
 While interacting with a user, the robot quite often needs to clarify an
@@
 perspective or from a specific agent's perspective. Usage of these
-categorization abilities are illustrated in the \emph{Spy Game}
-(section~\ref{spygame}).
+categorization abilities are illustrated in Sections~\ref{dialogs:disamb} and~\ref{spygame}.
 
@@
 
-We have designed a simplified bio-inspired memory model that allows us to store
+We have designed a simplified bio-inspired memory model to store
 statements in different \emph{memory profiles}. These include \emph{short term
@@
 
-Lastly, ORO allows external modules to be triggered when specific events occur,
- for instance, when a logical sentence becomes true or false, or if a new
+Lastly, ORO allows external modules to be triggered when specific events occur. For instance, when a logical sentence becomes true or false, or if a new
 instance of a certain class is added. One immediate application is reactive
@@
 \section{Geometric reasoner for situation assessment}
+\label{spark}
 
 Anchoring perceptions in a symbolic model requires perception abilities
-and their symbolic assessments. In this section we present
-SPARK (\emph{SPAtial Reasoning \& Knowledge}, \cite{Sisbot2011}), a situation assessment reasoner
+and their symbolic interpretation. In this section we present
+SPARK (\emph{SPAtial Reasoning \& Knowledge}~\cite{Sisbot2011}), a situation assessment reasoner
 that generates relevant symbolic information from the geometry of the
@@
 ``what it can see'', \ie what is within its field of view (FOV). A robot being
-able to compute this information can then act accordingly. A typical example
+able to compute this information can then act accordingly. An example
 would be a clarification scenario where the human is searching for an object
-and the robot is able to infer that it is looking for the one that is not
+and the robot is able to infer that he/she is looking for the one that is not
 visible (otherwise the user would not be searching for it).
-
 In figure~\ref{fig::vis} the field of view of a person is illustrated with a
 grey cone (broader one). While he is able to see the two small boxes on the
-table in front of the him, the big box on his right is out of this FOV, and
+table in front of him, the big box on his right is out of his FOV, and
 therefore, he is not able to see it.
@@
 agent refers to?) that should be solved through higher level reasoning (\eg
-context analysis or clarification through verbal interaction as shown later
-on).
+context analysis or clarification through verbal interaction).
 
@@
 point where the user is able to get the object afterward. 
-
 Figure~\ref{fig::reach} shows different reachability postures for each object
-on the table. In the example, the bottle and the box are both reachable by the
+on the table. In the example, the bottle and the box are both reachable for the
 human, but the teddy bear is too far. Instead, from the robot's perspective,
-the teddy bear is reachable, while the bottle is not.
+the teddy bear is reachable for it, while the bottle is not.
 
@@
 
-One way of referring to object's
-positions is based on a human's symbolic descriptors, instead of using their precise
-position. In fact, in many cases, this information is the most precise information available
- since we do not store the numeric coordinates of objects. The
-following relations are computed with respect to the position of the agents and
-the objects.
+One way of referring to object's positions is based on human's symbolic 
+descriptors, instead of using their precise position. In fact, in many cases, 
+this information is the most precise information available since humans do not 
+store the numeric coordinates of objects. These type of descriptors have been 
+studied in the context of language grounding 
+(\cite{O'Keefe1999,Matuszek2010,Regier2001,Kelleher2009,Blisard2005}). In this 
+work we focus on the following relations which are computed with respect to the 
+position of the agents and the objects:
 
@@
 \concept{isLocatedAt} represents spatial locations between agents and objects.
-For example we say ``it is on my right, on your left, ...'' These type of direction commands are studied
-in the literature in the context of language grounding (\eg \cite{O'Keefe1999,Matuszek2010}). We compute these
+For example we say ``it is on my right, on your left, ...'' We compute these
 spatial locations by dividing the space around the referent (an agent) into $n$
@@
 locations with respect to other objects in the environment, such as \emph{above,
-next to, in}, etc. These types of relations are widely studied in language
-grounding (\eg \cite{Regier2001} presented different models to define the
-\emph{above} relation; \cite{Kelleher2009} introduced two computational models of topological and projective spatial prepositions; \cite{Blisard2005} employed contour points of segmented objects to determine \emph{front, behind, left,right} relations).
-In this work we use similar models based on the
-bounding boxes and center of mass of the objects to define three main relations
-(Figure~\ref{fig::sprelation}):
+next to, in}, etc. In this work we compute three main relations based on the
+bounding box and center of mass of the objects (Figure~\ref{fig::sprelation}):
 
@@
 
-To ensure the different agent's models are up-to-date, all these properties are
+To ensure the different agent models are up-to-date, all these properties are
 always computed on-line, each time the current state of the world changes.
@@
 amodal physical representation of the world used as a mediator between the
-sensor space and symbolic models. They have however a different set of
-features: while GSM enables representation of time and imaginary objects (whose
+sensor space and symbolic models. They have however different features: while GSM enables representation of time and imaginary objects (whose
 existence is hinted by verbal assertions from a human, also called
@@
 
-We have developed a dedicated module called {\sc Dialogs} that processes human
+To this end we have developed a dedicated module called {\sc Dialogs} that processes human
 input in natural language, grounds the concepts in the robot's knowledge and
@@
 grammar-free) tool that extracts the grammatical structure from the user's
-sentence. Figure~\ref{dialog|parser_output} shows the raw output of the parser for a
-moderately complex sentence.
+sentence. Figure~\ref{dialog|parser_output} shows an example of the raw output of the parser for a moderately complex sentence.
 
@@
 
-The sentence resolution has three steps: {\it(1)} pronouns and anaphora are
-replaced by the correct speaker ID and the ID of the last object spoken about
-(extracted from the dialogue history) respectively, {\it(2)} nominal groups are
-disambiguated and grounded (noun phrase resolution, \cf below), and {\it(3)}
+The sentence resolution has three steps: {\it(i)} pronouns and anaphora are
+replaced by the correct speaker ID and the ID of the last object referred to
+(extracted from the dialogue history) respectively, {\it(ii)} nominal groups are
+disambiguated and grounded (noun phrase resolution), and {\it(iii)}
 verbal groups are resolved and their associated \emph{thematic roles} are
@@
 communication with the knowledge base. All the concepts the robot manipulates
-are stored in the \textit{ontology server} and retrieved through logical
+are stored in the ontology server and retrieved through logical
 queries, except for the verbs that are currently stored in a dedicated library
@@
 
-In order to better understand the overall process of the {\sc Dialogs} module and its relation with ORO, we next describe the different steps of the approach based on three examples.
-In these examples, we
-assume that some initial facts are present in the knowledge base, both in the
-robot's own model and in the human's model.  Since the robot tries to ground a
+In order to better understand the overall process of the {\sc Dialogs} module 
+and its relation with ORO, we next describe the different steps of the approach 
+based on three examples. In these examples we assume that some initial facts 
+are present in the knowledge base, both in the robot's own model and in the 
+human's model.  Since the robot tries to ground a
 human utterance, all queries are sent to the human model in order to interpret
@@
 \subsection{Informational content extraction}
-
-Figure~\ref{dialog|ex1} shows a first example of human discourse grounding and
-the extraction of informational content. We assume that the robot knowledge
-base only contains two initial statements in the human model. The user
-asserts a new one: ``The yellow banana is big!''.
 
@@
 
+Figure~\ref{dialog|ex1} shows a first example of human discourse grounding and
+the extraction of informational content. We assume that the robot knowledge
+base only contains two initial statements in the human model. The user
+asserts a new one: ``The yellow banana is big!''. 
 We first want to match the nominal group \emph{The yellow banana} to an already
-known concept, and second to translate the property \emph{is big} into a
-predicate (\emph{hasSize}) to state its semantics. The general resolution
-algorithm of nominal groups is described in algorithm~\ref{algo|Resolution}.
+known concept (algorithm~\ref{algo|Resolution}), and second to translate the property \emph{is big} into a predicate (\emph{hasSize}) to state its semantics.
 
@@
 manipulable physical item (\concept{Artifact}) as direct object. Thus, if the
-concept the robot finds for the thematic role \concept{actsOnObject} can not be
+concept the robot finds for the thematic role \concept{actsOnObject} cannot be
 inferred to be an artifact, the robot goes back to the human saying it does not
@@
 \subsection{Informational content extraction requiring clarification}
+\label{dialogs:disamb}
 \begin{figure}
@@
 \section{Experiments}
+\label{experiments}
 
@@
 holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
-System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
+System}, \url{http://www.ros.org}} middleware; and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
 Genom/pocolibs} stack. Both platforms share the use of ORO, the ontology server proposed in this work.
@@
 Garage's RViz. The clusters of point are given a unique identifier by the
-perception that allows the supervision to create the link between the physical
+perception module. This allows the supervision module to create the link between the physical
 objects and their symbolic representation in ORO.} 
@@
 	knows about \emph{Tableware} (it belongs to the common-sense ontology), so
-	the interaction stops.]}\\
+	the description of this object ends.]}\\
 	& (...) \\
@@
 \subsection{\emph{Spy Game} experiment}
-
 \label{spygame}
@@
 description. Based on this subset, ORO provides a descriptor (or set of
-descriptors) that allows a maximum discrimination among objects in the subset.
+descriptors) that allows maximum discrimination among objects in the subset.
 The robot queries the user about the value of the descriptor (or the most
@@
 both agents can be considered in the game) are automatically computed on-line
-by \textsc{SPARK} (the geometric reasoner). Figure~\ref{fig|spyGameExample}
+by \textsc{SPARK}. Figure~\ref{fig|spyGameExample}
 shows an example of a round game.
@@
 				& \emph{[obtain discriminants: type, color.]}\\
-				& Which type of object is: bottle or cup? \\
+				& Which type of object it is: bottle or cup? \\
 		\chatN{human} & Bottle.\\
@@
 daily life situation. Tom and Jerry are moving to London, so they are packing
-things in boxes. The scenario takes places in the living-room, where Jido (our
-robot) is observing while they move things here and there. To assess the
+things in boxes. The scenario takes places in the living-room, where Jido is observing while they move things here and there. To assess the
 reasoning abilities of the robot they ask Jido for information (entered through
-keyboard). Ideally, the robot should also perform actions when required (\eg
+keyboard). Ideally, the robot should perform actions when required (\eg
 hand an object when asking ``give me...''). However, since it is out of the
@@
 First Tom enters the room with some of the things they need to pack: a
-toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
+toolbox and two videos. He leaves one of the video tapes (``Jido-E'') inside one of
 the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
-table. We next describe how the situation assessment takes place, \ie how the
+table. We next describe how the situation assessment takes place at each step, \ie how the
 ontology is updated with the information obtained from the geometric reasoner
@@
 
-Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a).
-He places a toolbox and a video tape on the table, and another video tape
-inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:
+Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a) and places them around. Then he leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:
 
@@
 \subsubsection{Implicit disambiguation through visual perspective taking}
+\label{exp:disambVPT}
 
@@
 that the human refers to it and can eventually execute the command, \ie give
-it to the human. Alternatively, the robot could first verify with the human if
-that was the object being referred to or not before proceeding to execute the
+it to the human. Alternatively, the robot could first verify with the human whether
+that was the object being referred to before proceeding to execute the
 action. Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and
@@
 box, the toolbox or the cardboard box?'' Jerry could now simply answer the
-question. Instead, he decides to point at it while indicating: ``This box''
+question. Instead, he decides to point at it while indicating: ``That box''
 (Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
@@
 In the meantime, the \textsc{Dialogs} module is processing the human verbal input.
-When trying to resolve the reference ``this'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
+When trying to resolve the reference ``that'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
 \begin{center}
@@
 
-Finally, the \textsc{Dialogs} queries the ontology about the content of the box
+Finally, \textsc{Dialogs} queries the ontology about the content of the box
 and the question can be answered: ``Jido-E''. Note that the object's label is
 used instead of its ID. This way we enhance interaction using familiar names
-given by the users.
-
+given by the users:
 \begin{center}
@@
 
-At this point Jerry wants to know where the other tape is, and that is exactly
-what he asks Jido: ``And where is the other tape?''. In this occasion, the
+At this point Jerry wants to know where the other video tape is, and that is exactly
+what he asks Jido: ``And where is the other video tape?''. In this occasion, the
 \textsc{Dialogs} module is able to interpret that Jerry is not referring to the
-video which they were just talking about, but to the other one:
+video tape which they were just talking about, but to the other one:
 
@@
 
-Since there is only one possible ``other'' video (there are only two videos in
-the scene), it can directly answer Jerry: ``The other tape is on the table and
+Since there is only one possible ``other'' video tape (there are only two video tapes in
+the scene), it can directly answer Jerry: ``The other video tape is on the table and
 next to the toolbox.''
@@
 \section{Conclusion}
+\label{conclusion}
 
@@
 
-In this paper, we studied knowledge flows between three components: {\it(1)}
+In this paper, we have studied knowledge flows between three components: {\it(1)}
 {\sc ORO}, an ontology-based knowledge server that stores and maintains
-classified RDF statements produced by the other modules in agent-specific
+classified RDF statements produced by other modules in agent-specific
 models and allows information to be easily retrieved, either through queries or
@@
 \item{Knowledge is explicitly stored in one central and consistent repository
-of facts, accessible by all modules,}
+of facts, accessible by all modules.}
 \item{Knowledge is represented in a strict formalism (RDF statements) and
 with a clearly defined vocabulary (stated in the {\tt commonsense.oro.owl}
-ontology),}
+ontology).}
 \item{The first two points enable both a loosely-coupled
@@
 trigger events whose \textit{True} conditions can be inferred.}
-\item{Finally, this architecture allows for merging very different knowledge
+\item{Finally, this architecture allows for the combination of very different knowledge
 modalities in a single homogeneous environment, bringing mutual benefits to
@@
 to information alignment and fusion in complex robotic systems.
-
 The versatility of this paradigm could be illustrated by a simple imaginary
 scenario with a blind robot and a deaf robot. The blind robot does not see (no
-cameras or alike), but someone can verbally describe a scene to it, while the
-deaf robot has a good vision system but can not hear.  Without any
-changes to the software architecture described, supervisors of both robots would be able to perform equally well.
+cameras or alike), but someone can verbally describe a scene to it. On the other hand, the
+deaf robot has a good vision system, but cannot hear.  Without any
+changes to the software architecture described, supervision modules of both robots would be able to perform equally well.
 
@@
 validity} nor the context of a fact are represented in a satisfying way (we do
-store some kind of context -- the agent mental model for instance). Those
-informations are meta-informations on the knowledge, and the ORO framework,
-while it allows it through \emph{statement reification}, does not offer yet
+store some kind of context --the agent mental model for instance--). Those
+informations are meta-informations on the knowledge. While the ORO framework allows them through \emph{statement reification}, it does not offer yet
 a convenient way to store them. One obvious limitation that derives from
@@
 to a verb. 
-This could be improved by integrating a symbolic task
+This could be improved by linking a symbolic task
 planner to the \textsc{Dialogs} module to dynamically provide the list of
@@
 The three experiments that were presented in the paper all illustrate how the
-robot make use of its embodied nature to establish a meaningful communication
+robot makes use of its embodied nature to establish a meaningful communication
 with a human. Mainly, because the robot and the human share the same physical
@@
 However, we can build a model of the ``human point of view'' because the robot
-perceives the human, and is able to deduce, at least partially, what the human
+perceives the human, and is able to estimate, at least partially, what the human
 perceives or not. We infer that a human focuses on some object because it points at it, looks at it, and besides, the object is visible to him. This relies on the embodied
-nature of the interaction. And this allows us, in turn, to understand the
-meaning of sentences like ``Give me this''.
+nature of the interaction. In turn, this allows us to understand the
+meaning of sentences like ``Give me that''.
 
@@
 
-Let us conclude this article by a short glance back at Roy's list of challenges
-for human-robot dialogue: 
+Let us conclude this article briefly reviewing and linking Roy's list of challenges
+for human-robot dialogue with our current approach: 
 \begin{itemize}
@@
 	can be added, we have actually proposed a \emph{cross-modal
-	representation system},
+	representation system}.
 
@@
 	ability for the robot to \emph{associate words with perceptual and action
-	categories},
+	categories}.
 
 	\item The ORO knowledge base offers some support for the \emph{modeling of
-	context}, but a lot remains to be done in this respect,
+	context}, but a lot remains to be done in this respect.
 
@@
 	level. Generally speaking, it appears that complex robotic systems need
-	to operate with a dynamic granularity, depending on the task to achieve,
+	to operate with a dynamic granularity, depending on the task to achieve.
 
@@
 	essential tool to build predictive knowledge, and we are looking into this
-	direction,
+	direction.
 
@@
 	seems of particular importance for the robot to assess the state of the
-	interaction with the human,
+	interaction with the human.
 

### 2011-07-18 paper_knowledge.tex
Removed unused packages
Polished the introduction - moved some part elsewhere
- moved explanation on different models in oro to oro section
- moved discussion on layred architectures to the conclusion

@@
 
-\usepackage{graphicx}
 \usepackage[utf8]{inputenc}
-\usepackage{listings}
-\usepackage{fancyvrb}
 \usepackage[usenames]{color}
@@
 
-\usepackage{eso-pic} %% Background
-%\usepackage{floatflt}
-
 \usepackage{enumerate}
@@
 
+\usepackage{subfigure}
+
 \usepackage{pseudocode}
@@
 
-\usepackage{subfigure}
 %diagrams
 \usepackage{tikz}
-%\usetikzlibrary{shapes,decorations,trees}
 \usetikzlibrary{shapes,trees}
-
-% rotate text in tabular mode
-\usepackage{rotating}
 
@@
 interactions. We present an integrated approach, including a specialized symbolic
-knowledge representation approach based on Description Logics, and case studies
+knowledge representation system based on Description Logics, and case studies
 on several robotic platforms that demonstrate these cognitive capabilities.
@@
 (Fig.~\ref{fig|vpt}). ``Jido, give me that'', says Thomas while looking at a box that contains a
-video tape. The robot smoothly grasps the tape, and hands it to the him.
+video tape. The robot smoothly grasps the tape, and hands it to him.
+
+\begin{figure}%[!ht] 
+	\centering
+	\includegraphics[width=1.0\linewidth]{images/pt.jpg} 
+	\caption{Interacting with
+	the robot in an everyday setup: the human asks for help in vague terms, the
+	robot takes into account the human's spatial perspective to refine its
+	understanding of the question.} 
+	\label{fig|vpt} 
+\end{figure}
 
@@
 
-The first step is to understand the meaning of the sentence. To this end, we must acquire the sentence, convert it into a useful syntactic form (probably through
-speech recognition), and understand the semantics of the sentence, \ie What is
-referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
+The first step is to understand the meaning of the sentence. To this end, we
+must acquire the sentence, convert it into a useful syntactic form (probably
+through speech recognition), and understand the semantics of the sentence, \ie
+What is referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
 ``\textit{me}''? And ``\textit{that}''?
 
-Working in a situated context, we need to \emph{resolve} these
-semantics atoms, \ie ground them in the sensory-motor space of the robot. For
-instance, ``\textit{this}'' is a demonstrative pronoun that refers in this
-context to the object the human is focusing on (whatever \textit{focusing}
-means). In this example Thomas is looking at something, which is a possible
-cue. But it could as well be pointing at something or making reference to some
-previously mentioned concept.
+Working in a situated context, we need to \emph{resolve} these semantics atoms,
+\ie ground them in the sensory-motor space of the robot. For instance,
+``\textit{that}'' is a demonstrative pronoun that refers in this context to the
+object the human is focusing on.
 
 The next step is to extract and understand the \emph{intent} of the utterance
-as thought by the agent.
-In our example, Thomas obviously wants an action to be performed by the robot.
-The action parametrization is conveyed by the semantics attached to the words and the grammatical structure of the sentence.
-In our example, the type of action is
-given by the verb ``\textit{give}''. Assuming the robot has some procedural
-knowledge attached to this symbol, the action type can be considered as
-grounded for the robot. We can as well understand that the recipient of the
-action is the human, the performer is the robot itself, and the object acted
-upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
-that can be extracted from the sentence that allow for fully grounding the action.\footnote{This analysis has been inspired on the work of Austin et al.~\cite{Austin1962}, where this type of sentences correspond to \emph{speech acts}, comprising of \emph{locutionary act} (the meaning), \emph{illocutionary} (the intent) and possibly \emph{perlocutionary acts} (implicit speaker's expectation).}
+as thought by the agent. In our example, Thomas obviously wants an action to be
+performed by the robot. The action parametrization is conveyed by the semantics
+attached to the words and the grammatical structure of the sentence. In our
+example, the type of action is given by the verb ``\textit{give}''. Assuming
+the robot has some procedural knowledge attached to this symbol, the action
+type can be considered as grounded for the robot. We can as well understand
+that the recipient of the action is the human, the performer is the robot
+itself, and the object acted upon is the tape. The recipient, performer and object are three of the
+\emph{thematic roles}~\cite{Gruber1965} that qualify the \emph{give} action. They are necessary to fully ground the sentence\footnote{This analysis has been
+inspired on the work of Austin et al.~\cite{Austin1962}, where this type of
+sentences correspond to \emph{speech acts}, comprising of \emph{locutionary
+act} (the meaning), \emph{illocutionary} (the intent) and possibly
+\emph{perlocutionary acts} (implicit speaker's expectation).}.
 
@@
 
-\begin{figure}%[!ht] 
-	\centering
-	\includegraphics[width=1.0\linewidth]{images/pt.jpg} 
-	\caption{Interacting with
-	the robot in an everyday setup: the human asks for help in vague terms, the
-	robot takes into account the human's spatial perspective to refine its
-	understanding of the question.} 
-	\label{fig|vpt} 
-\end{figure}
-
 \subsection{Approach}
 
-In this work we propose an approach aiming at grounding the interaction with users, verbal and non-verbal, following the steps described above.
-To this end, we have developed three distinct, inter-related cognitive components, which we believe fulfill the minimum requisites for situated human interaction.
-Figure~\ref{fig|overall-process} illustrates the general context, while
-Figure~\ref{fig|main-components} summarizes the interactions between these three
-main components.
+In this work we propose an approach aiming at grounding the interaction with
+users. Figure~\ref{fig|overall-process} illustrates the general contex. We have
+developed three distinct, inter-related cognitive components, as illustrated in
+Figure~\ref{fig|main-components}.
 
@@
 \item \emph{Physical environment modeling} and \emph{spatial reasoning}
-(grouped under the term \emph{situation assessment}): these components are in
+(grouped under the term \emph{situation assessment}): this component is in
 charge of building and maintaining a coherent model of the physical world. This
@@
 \item \emph{Knowledge representation and management}: the robot is endowed with
-an active knowledge base that provides a logically sound symbolic model of its
-beliefs of the world, as well as models for each cognitive agent the robot
-interacts with. Each of these models is independent and logically consistent,
-enabling reasoning on different perspectives of the world that would otherwise
-be considered inconsistent (for instance, an object can be visible for the
-robot but not for the human. This object can have at the same time the property
-\concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}} using
-two different models).  Our platform relies on OWL-DL\footnote{\emph{Web
-Ontology Language - Description Logics}, a decidable subset of the first-order
-logics, \url{http://www.w3.org/TR/owl2-primer/}} ontologies and features
-continuous storage, querying and event triggering over the pool of facts known
-by the robot.
-
-Used in combination with the situation assessment framework, the robot is thus
-able to maintain different symbolic models of the world, one per agent. This
-proves an essential feature (\cite{Roy2005,Kruijff2010}) to enable
-perspective-aware grounding of natural language as we will see in next
-sections.
-
-\item \emph{Dialogue input processing}: a subsystem which includes natural
-language parsing capabilities, disambiguation routines and interactive concept
+an active knowledge base that provides a symbolic model of its beliefs of the
+world, as well as models for each cognitive agent the robot interacts with.
+Used in combination with the situation assessment framework, this proves an
+essential feature (\cite{Roy2005,Kruijff2010}) to enable perspective-aware
+grounding of natural language.
+
+Our platform relies on OWL-DL\footnote{\emph{Web Ontology Language -
+Description Logics}, a decidable subset of the first-order logics,
+\url{http://www.w3.org/TR/owl2-primer/}} ontologies and features continuous
+storage, querying and event triggering over the pool of facts known by the
+robot.
+
+\item \emph{Dialogue processing}: the third component includes natural
+language parsing capabilities, interactive disambiguation routines and concept
 anchoring. We focused our efforts on three classes of utterances commonly found
@@
 \emph{questions on declarative knowledge} (whose answers do not require
-explicit planning). This would roughly cover the \emph{representative}
+planning or prediction)\footnote{This would roughly cover the \emph{representative}
 (sometimes referred as \emph{assertive}) and \emph{directive} illocutionary
-acts in Searle's~\cite{Searle1976} classification.
+acts in Searle's~\cite{Searle1976} classification.}.
 
 \end{inparaenum}
-
-Based on these three components we have developed several grounding mechanisms
-for situated verbal interaction and integrated them into a broader control
-architecture for indoor service robots.
 
@@
 creating and maintaining the correspondence between symbols and sensor data
-that refer to the same physical object''}, Ziemke~\cite{Ziemke1999} who
-elaborates on the need of embodiment for symbol grounding, and
-Sloman~\cite{Sloman2007} who criticizes the \emph{``concept empricism"} that
-underlies symbol grounding.
+that refer to the same physical object''} and Ziemke~\cite{Ziemke1999} who
+elaborates on the need of embodiment for symbol grounding.
 
 Our contribution relates to two narrower fields: natural language in embodied
-interaction contexts and knowledge building and representation in robotic
+interaction contexts and knowledge acquisition and representation in robotic
 systems.
@@
 low-level percepts (including view-point independent SIFT based object
-recognition) to a high-level knowledge representation and reasoning system via
-a dedicated anchoring module. They base their knowledge model directly on the
-\textit{ResearchCyc} ontology (including the \textit{MicroTheories} concept),
-used in combination with the {\sc CycL} language. This enables second-order
-logic modeling and access to a large common-sense knowledge base.
+recognition) to a high-level knowledge representation and reasoning system.
+They base their knowledge model directly on the \textit{ResearchCyc} ontology
+(including the \textit{MicroTheories} concept), used in combination with the
+{\sc CycL} language. This enables second-order logic modeling and access to a
+large common-sense knowledge base.
 
@@
 dynamic, partially unknown human environments; and fully embodied (with arms,
-head, ability to move,...) autonomous robots that can perform manipulation.
-
-Four specific contributions are presented in this work: first, we introduce a
-novel and versatile knowledge base that models in a formal framework, based on
-first-order logics, not only the robot's own beliefs but also every other
-cognitive agent the robot interacts with.  This explicit modeling of other
-agents' belief states is used for the interaction and eases the implementation
-of various advanced cognitive behaviors like learning new concepts or
-interactive object discrimination.
+head, ability to move,...) autonomous robots that manipulate a large range of
+household objects.
+
+Three specific contributions are presented in this work: first, we introduce a
+versatile and light-weighted knowledge base that models in a formal framework,
+based on first-order logics, not only the robot's own beliefs but also every
+other cognitive agent the robot interacts with.  This explicit modeling of
+other agents' belief states is used for the interaction and eases the
+implementation of various advanced cognitive behaviors like
+False-Beliefs~\cite{Leslie2000} or interactive object discrimination.
 
@@
 on situated language grounding. We present several examples that include
-generic recognition and semantic validation of thematic roles (\ie the
-semantic roles played by the different nominal groups around a verb), which
-can be translated into input for a symbolic task planner, or disambiguation
+recognition and semantic validation of thematic roles or disambiguation
 based on attention foci.
 
-Lastly, our architecture moves away from standard layered approaches. The way
-our different components interact together is not easily represented as a
-layered architecture because interactions among components are mostly
-bidirectional.  This is especially visible for the dialogue input processing:
-this component does not simply act as an alternative perceptual input to the
-symbolic database, but also actively queries previously acquired knowledge to
-disambiguate and validate the newly created symbolic knowledge. This behaviour
-leads us to the idea of \emph{knowledge-oriented architecture}, which is
-detailed later on.
+Communcation between these components is build as streams of symbolic facts,
+where knowledge that the robot manipulate is made explicit.
+This leads us to the idea of a \emph{knowledge-oriented architecture} which is
+discussed at the end of the article.
 
@@
 components.
-Section~\ref{spark} presents {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge}), a
-module that merges perceptual information with a coherent geometric model and
+Section~\ref{spark} presents {\sc SPARK} (for \emph{SPAtial Reasoning \& Knowledge}), the
+component that merges perceptual information with a coherent geometric model and
 builds a symbolic interpretation of the world from the robot's point of view,
@@
 models of the world.
+
+Each of these models is independent and logically consistent,
+enabling reasoning on different perspectives of the world that would otherwise
+be considered as globally inconsistent (for instance, an object can be visible for the
+robot but not for the human. This object can have at the same time the property
+\concept{isVisible \textbf{true}} and \concept{isVisible \textbf{false}} in
+two different models).
 
@@
 
-In this paper, we have studied knowledge flows between three components: {\it(1)}
+In this paper, we have studied knowledge streams between three components: {\it(1)}
 {\sc ORO}, an ontology-based knowledge server that stores and maintains
@@
 of facts, accessible by all modules.}
-\item{Knowledge is represented in a strict formalism (RDF statements) and
+\item{Knowledge is represented in a strict formalism (OWL statements) and
 with a clearly defined vocabulary (stated in the {\tt commonsense.oro.owl}
@@
 the reasoner, we take full advantage of the inference abilities of ORO to
-trigger events whose \textit{True} conditions can be inferred.}
+trigger events whose \texttt{true} conditions can be inferred.}
 \item{Finally, this architecture allows for the combination of very different knowledge
@@
 \end{itemize}
+
+This architecture moves away from standard layered approaches: Interactions
+between components are mostly bidirectional and we do not have a concept of
+layers of abstraction (we do, however, have access to the robot lower levels
+to execute actions, but all cognition-related modules live at the same level).
+This is especially visible for the dialogue input processing: this component
+does not simply act as an alternative perceptual input to the symbolic
+database, but also actively queries previously acquired knowledge to
+disambiguate and validate the newly created symbolic knowledge.
 
@@
 enabling independent modeling of the beliefs of each agent, our architecture is
-especially well suited to set up cognitive and psychological experiments (such as the \emph{False-Belief} experiment~\cite{Leslie2000}), which we plan to further explore.
+especially well suited to set up cognitive and psychological experiments (such
+as the \emph{False-Belief} experiment~\cite{Leslie2000}), which we plan to
+further explore.
 

### 2011-07-18 paper_knowledge.tex
Removed screenshot of oro-view

@@
 
-\begin{figure}[!ht]
-\centering
-  \includegraphics[width=\columnwidth]{images/snapshot_oroview.png}
-  \caption{Live snapshot from the \emph{oro-view} viewer depicting the
-  knowledge model of the robot during an experiment. Dots represent
-  instances (ABox), while circles pyramids are classes (TBox)}
-  \label{fig|oro-view}
-\end{figure}
+%\begin{figure}[!ht]
+%\centering
+%  \includegraphics[width=\columnwidth]{images/snapshot_oroview.png}
+%  \caption{Live snapshot from the \emph{oro-view} viewer depicting the
+%  knowledge model of the robot during an experiment. Dots represent
+%  instances (ABox), while circles pyramids are classes (TBox)}
+%  \label{fig|oro-view}
+%\end{figure}
 
@@
 
-Figure~\ref{fig|oro-view} shows a snapshot of the knowledge model during an
-experiment. Both concepts from the common sense ontology and
-experiment-specific instances are visible.
-
 \subsection{Reasoning and dynamic knowledge structuring}
@@
 		\footnotesize \concept{looksAt(?agt, ?obj)} $\land$
-		\concept{pointsAt(?agt,?obj)} $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
+		\concept{pointsAt(?agt,?obj)} \par $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
 		\normalsize 

### 2011-07-19 paper_knowledge.tex
Give more freedom to Latex to place figures
Minor revisions by Raquel

@@
 Communcation between these components is build as streams of symbolic facts,
-where knowledge that the robot manipulate is made explicit.
-This leads us to the idea of a \emph{knowledge-oriented architecture} which is
+where knowledge manipulated by the robot is made explicit.
+This leads us to the idea of a \emph{knowledge-oriented architecture}, which is
 discussed at the end of the article.
@@
 
-\begin{figure}[!t] \centering
+\begin{figure}%[!t] 
+\centering
 \includegraphics[width=\columnwidth]{images/looks.jpg} \caption{Representation
@@
 
-\begin{figure}[!t] 
+\begin{figure}%[!t] 
 	\centering
@@
 
-\begin{figure}[!ht]
+\begin{figure}%[!ht]
 \centering
@@
 
-\begin{figure}[!ht]
+\begin{figure}%[!ht]
 \begin{center}
@@
 We first want to match the nominal group \emph{The yellow banana} to an already
-known concept (algorithm~\ref{algo|Resolution}), and second to translate the property \emph{is big} into a predicate (\emph{hasSize}) to state its semantics.
+known concept (algorithm~\ref{algo|Resolution}), and second to translate the property \emph{is big} into a predicate ({\tt hasSize}) to state its semantics.
 
@@
 further exploit the knowledge reasoner in this experiment, we ask the robot
-which object do not belong to a typical breakfast table, \ie we ask for the odd
+which object does not belong to a typical breakfast table, \ie we ask for the odd
 objects. In this case, it corresponds to objects that are neither of type
@@
 
-\begin{figure}[!t]
+\begin{figure}%[!t]
 \centering
@@
 
-\begin{figure}[!ht]
+\begin{figure}%[!ht]
   \centering
@@
 
-This architecture moves away from standard layered approaches: Interactions
+This architecture moves away from standard layered approaches. Interactions
 between components are mostly bidirectional and we do not have a concept of
-layers of abstraction (we do, however, have access to the robot lower levels
-to execute actions, but all cognition-related modules live at the same level).
-This is especially visible for the dialogue input processing: this component
-does not simply act as an alternative perceptual input to the symbolic
-database, but also actively queries previously acquired knowledge to
+layers of abstraction (we do, however, have access to the robot lower level
+modules to execute actions, but all cognition-related modules live at the same
+level). This is especially visible for the dialogue input processing. This
+component does not simply act as an alternative perceptual input to the
+symbolic database, but also actively queries previously acquired knowledge to
 disambiguate and validate the newly created symbolic knowledge.
@@
 especially well suited to set up cognitive and psychological experiments (such
-as the \emph{False-Belief} experiment~\cite{Leslie2000}), which we plan to
+as the \emph{False-Belief} experiment), which we plan to
 further explore.

### 2011-10-19 paper_knowledge.tex
Fixed capitalization of titles to match stupid american rules

@@
 
-\title{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}
+\title{Grounding the Interaction: Anchoring Situated Discourse in Everyday Human-Robot Interaction}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Grounding human interaction into robot knowledge}
+\section{Grounding Human Interaction Into Robot Knowledge}
 
@@
 
-\subsection{Related work}
+\subsection{Related Work}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ORO, a knowledge management platform} \label{cognitivekernel}
-
-\subsection{On knowledge representation} \label{knowledge_representation}
+\section{ORO, a Knowledge Management Platform} \label{cognitivekernel}
+
+\subsection{On Knowledge Representation} \label{knowledge_representation}
 
@@
 
-\subsection{Reasoning and dynamic knowledge structuring}
+\subsection{Reasoning and Dynamic Knowledge Structuring}
 
@@
 
-\subsection{ORO features}
+\subsection{ORO Features}
 \label{features}
@@
 
-\subsubsection{Base functionalities}
+\subsubsection{Base Functionalities}
 \label{base}
@@
 
-\subsubsection{Representation of alternative cognitive models}
+\subsubsection{Representation of Alternative Cognitive Models}
 \label{alterite}
@@
 
-\subsubsection{Memory profiles}
+\subsubsection{Memory Profiles}
 \label{memory}
@@
 
-\subsubsection{The events framework} 
+\subsubsection{The Events Framework} 
 \label{events}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Geometric reasoner for situation assessment}
+\section{Geometric Reasoner for Situation Assessment}
 \label{spark}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The natural language grounding process}
+\section{The Natural Language Grounding Process}
 \label{dialog}
@@
 
-\subsection{Informational content extraction}
+\subsection{Informational Content Extraction}
 
@@
 
-\subsection{Intentional content through verb resolution}
+\subsection{Intentional Content Through Verb Resolution}
 The sentence in the first example is built with the state verb \emph{be} at
@@
 
-\subsection{Informational content extraction requiring clarification}
+\subsection{Informational Content Extraction Requiring Clarification}
 \label{dialogs:disamb}
@@
 
-\subsection{\emph{Naming} experiment}
+\subsection{\emph{Naming} Experiment}
 \label{naming}
@@
 
-\subsection{\emph{Spy Game} experiment}
+\subsection{\emph{Spy Game} Experiment}
 \label{spygame}
@@
 
-\subsection{\emph{Moving to London} scenario}
+\subsection{\emph{Moving to London} Scenario}
 
@@
 
-\subsubsection{Initial situation assessment}
+\subsubsection{Initial Situation Assessment}
 
@@
 
-\subsubsection{Implicit disambiguation through visual perspective taking}
+\subsubsection{Implicit Disambiguation Through Visual Perspective Taking}
 \label{exp:disambVPT}
@@
 
-\subsubsection{Explicit disambiguation through verbal interaction and gestures}
+\subsubsection{Explicit Disambiguation Through Verbal Interaction and Gestures}
 \label{disamb-gestures}
@@
 
-\subsection{Towards an event-driven, knowledge-oriented architecture for personal robotics}
+\subsection{Towards an Event-Driven, Knowledge-Oriented Architecture for Personal Robotics}
 
@@
 
-\subsubsection{Modeling the real world}
+\subsubsection{Modeling the Real World}
 
@@
 
-\subsubsection{On thematic roles and action models}
+\subsubsection{On Thematic Roles and Action Models}
 
@@
 
-\subsubsection{Knowledge and embodiment}
+\subsubsection{Knowledge and Embodiment}
 

### 2011-10-19 paper_knowledge.tex
Fixed all typos & style mistakes noted by reviewers
The biger questions remain to be dealt with.

@@
 
-A messy table, covered with cardboard boxes, books, video tapes... Thomas is
-moving out and is packing everything with the help of Jido, his robot
-(Fig.~\ref{fig|vpt}). ``Jido, give me that'', says Thomas while looking at a box that contains a
-video tape. The robot smoothly grasps the tape, and hands it to him.
+{\em A messy table, covered with cardboard boxes, books, video tapes... Thomas
+is moving out and is packing everything with the help of Jido, his robot.
+``Jido, give me that'', says Thomas while looking at a box that contains a
+video tape. The robot smoothly grasps the tape, and hands it to him.}
 
@@
 
-While this kind of interaction should hopefully sound quite familiar in a
-foreseeable future, our robots are not yet quite up to the task, neither
-regarding natural language understanding nor plan-making and manipulation.
-To be combined together, those abilities require an unambiguous and shared
-representation of concepts (objects, agents, actions...) underlying the
-interaction: what are the prerequisites for such a sentence --``Jido,
-give me that''-- to be understood by the robot, correctly interpreted in the
-spatial context of the interaction, and eventually transformed into an action?
+While this kind of interaction (Fig.~\ref{fig|vpt}) should hopefully sound
+quite familiar in a foreseeable future, our robots are not yet quite up to the
+task, neither regarding natural language understanding nor plan-making and
+manipulation.  To be combined together, those abilities require an unambiguous
+and shared representation of concepts (objects, agents, actions...) underlying
+the interaction: what are the prerequisites for such a sentence --``Jido, give
+me that''-- to be understood by the robot, correctly interpreted in the spatial
+context of the interaction, and eventually transformed into an action?
 
@@
 
-The next step is to extract and understand the \emph{intent} of the utterance
+The next step is to extract and understand the \emph{intended meaning} of the utterance
 as thought by the agent. In our example, Thomas obviously wants an action to be
@@
 symbolic knowledge repositories, our work extends these previous contributions
-by tackling more realistic human-robot interactions: open speech; complex,
-dynamic, partially unknown human environments; and fully embodied (with arms,
-head, ability to move,...) autonomous robots that manipulate a large range of
-household objects.
+by tackling more realistic human-robot interactions: less restricted speech
+understanding; ability to deal with complex, partially unknown human
+environments; and fully embodied (with arms, head,...) autonomous robots that
+manipulate a large range of household objects.
 
@@
 These points highlight some original aspects of a larger cognitive architecture
-that has been deployed and tested on several robotic platforms
+that has been deployed and tested on several mobile robotic platforms
 (including both humanoid robots and service robots), demonstrating the
@@
 isIn kitchen}). We use the Jena\footnote{\url{http://jena.sourceforge.net/}}
-framework as the underlying library to load and build in-memory OWL model. We
+framework as the underlying library to load and build an in-memory OWL model. We
 use it in conjunction with the
@@
 entities would return \concept{socrates} as well. The language of the OWL
-family use the \emph{Open World Assumption} (if a fact can not be inferred as
+family make the \emph{Open World Assumption} (if a fact can not be inferred as
 true, it does not mean that it is inferred to be false), and the Pellet
@@
 
-Next sections describe how symbolic knowledge is actually produced and added to
+The next sections describe how symbolic knowledge is actually produced and added to
 ORO.
@@
 visible (otherwise the user would not be searching for it).
-In figure~\ref{fig::vis} the field of view of a person is illustrated with a
+In Figure~\ref{fig::vis} the field of view of a person is illustrated with a
 grey cone (broader one). While he is able to see the two small boxes on the
@@
 verbal groups are resolved and their associated \emph{thematic roles} are
-retrieved (verbal phrase resolution).
+retrieved (verb phrase resolution).
 
@@
 
-This second example  also shows the pronouns reference resolution: ``me'' is
-replaced by the id of the current speaker while ``you'' is replaced by
+This second example  also shows the pronoun reference resolution: ``me'' is
+replaced by the id of the current speaker, while ``you'' is replaced by
 \concept{myself} (\concept{myself} always represents the robot itself). When
@@
 
-While no examples involving questions have been detailed, \emph{W-} questions
-and \emph{yes/no} questions can be processed in a similar way by
-\textsc{Dialogs}. For instance, a question like ``What is on the table?''
-is grounded (to extract the relation \concept{isOn} and to find what \emph{table}
+While no examples involving questions have been detailed, factual \emph{wh-}
+questions and polar (\emph{yes/no}) questions can be processed in a similar way
+by \textsc{Dialogs}. For instance, a question like ``What is on the table?'' is
+grounded (to extract the relation \concept{isOn} and to find what \emph{table}
 refers to) and transformed into the following kind of query: \concept{find ?var
-[\stmt{?var isOn table1}]}.  Answers are converted back to a full sentence
-by the \emph{verbalization} module, and uttered to the human.
+[\stmt{?var isOn table1}]}.  Answers are converted back to a full sentence by
+the \emph{verbalization} module, and uttered to the human.
 
@@
 This section presents several experiments where the different features of our
-approach are represented. The experiments have been conducted on two different platforms: the
-\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
-holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
-System}, \url{http://www.ros.org}} middleware; and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
-Genom/pocolibs} stack. Both platforms share the use of ORO, the ontology server proposed in this work.
+approach are represented. The experiments have been conducted on two different
+platforms: the \emph{Rosie} manipulator from the Technical Universtiy of
+Munich, a dual-arm, holonomic service robot, running the
+ROS\footnote{\emph{Robotic Operating System}, \url{http://www.ros.org}}
+middleware; and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a
+similar, single-arm, service robot, running the LAAS's {\sc Genom/pocolibs}
+stack. Both platforms share the use of ORO, the ontology server described in
+this work.
 
@@
 through interaction with the user. This task has been implemented on the Rosie
-robot at Munich university.
+robot at TU Munich.
 
@@
 daily life situation. Tom and Jerry are moving to London, so they are packing
-things in boxes. The scenario takes places in the living-room, where Jido is observing while they move things here and there. To assess the
+things in boxes. The scenario takes places in the living room, where Jido is observing while they move things here and there. To assess the
 reasoning abilities of the robot they ask Jido for information (entered through
@@
 architecture, including reasoning involving motion planing (to compute
-reachability of object) and perspective taking, and {\it(3)} {\sc Dialogs}, a
+reachability of objects) and perspective taking, and {\it(3)} {\sc Dialogs}, a
 natural language processor that performs simple grammatical parsing of English
@@
 between components are mostly bidirectional and we do not have a concept of
-layers of abstraction (we do, however, have access to the robot lower level
-modules to execute actions, but all cognition-related modules live at the same
-level). This is especially visible for the dialogue input processing. This
+layers of abstraction (we do, however, have access to the lower level modules
+of the robot to execute actions, but all cognition-related modules reside at the
+same level). This is especially visible for the dialogue input processing. This
 component does not simply act as an alternative perceptual input to the
@@
 symbolic contents about the world to ORO where it becomes accessible to
-decisional layers. In parallel, ORO relies on reasoning in a \textit{top-bottom}
+decisional layers. In parallel, ORO relies on reasoning in a \textit{top-down}
 way to produce new facts that may trigger in return physical behaviours.
 
-We believe that this \emph{knowledge-oriented} approach has a strong potential
-not only to enable rich human-robot interaction, but also as a broader approach
-to information alignment and fusion in complex robotic systems.
-The versatility of this paradigm could be illustrated by a simple imaginary
-scenario with a blind robot and a deaf robot. The blind robot does not see (no
-cameras or alike), but someone can verbally describe a scene to it. On the other hand, the
-deaf robot has a good vision system, but cannot hear.  Without any
-changes to the software architecture described, supervision modules of both robots would be able to perform equally well.
+% We believe that this \emph{knowledge-oriented} approach has a strong potential
+% not only to enable rich human-robot interaction, but also as a broader approach
+% to information alignment and fusion in complex robotic systems.  The
+% versatility of this paradigm could be illustrated by a simple imaginary
+% scenario with a blind robot and a deaf robot. The blind robot does not see (no
+% cameras or alike), but someone can verbally describe a scene to it. On the
+% other hand, the deaf robot has a good vision system, but cannot hear.  Without
+% any changes to the software architecture described, supervision modules of both
+% robots would be able to perform equally well.
 
@@
 
-We hope as well that our contribution may contribute to fill the gap between
-robotics and psychology. Our architecture provides easier-to-reach entry points
-to implement some classical psychology tests to robots. We presented
-experiments focused on issues related to perspective taking. By explicitly
-enabling independent modeling of the beliefs of each agent, our architecture is
-especially well suited to set up cognitive and psychological experiments (such
-as the \emph{False-Belief} experiment), which we plan to
-further explore.
+% We hope as well that our contribution may contribute to fill the gap between
+% robotics and psychology. Our architecture provides easier-to-reach entry points
+% to implement some classical psychology tests to robots. We presented
+% experiments focused on issues related to perspective taking. By explicitly
+% enabling independent modeling of the beliefs of each agent, our architecture is
+% especially well suited to set up cognitive and psychological experiments (such
+% as the \emph{False-Belief} experiment), which we plan to
+% further explore.
 
@@
 first-order logic framework).  We would like to relate it to
-Sloman~\cite{Sloman2007} stance against the \emph{``Symbol Grounding meme''},
+Sloman's~\cite{Sloman2007} stance against the \emph{``Symbol Grounding meme''},
 where he argues that symbolic grounding is bound to the representation of
@@
 validity} nor the context of a fact are represented in a satisfying way (we do
-store some kind of context --the agent mental model for instance--). Those
-informations are meta-informations on the knowledge. While the ORO framework allows them through \emph{statement reification}, it does not offer yet
-a convenient way to store them. One obvious limitation that derives from
-the lack of efficient meta-knowledge is the absence of knowledge history.
-With ORO, the robot always lives in the present.
+store some kind of context --the agent's mental model for instance--). This
+information is meta-information on the knowledge. While the ORO framework
+allows them through \emph{statement reification}, it does not offer yet a
+convenient way to store them. One obvious limitation that derives from the lack
+of efficient meta-knowledge is the absence of knowledge history.  With ORO, the
+robot always lives in the present.
 
@@
 However, we can build a model of the ``human point of view'' because the robot
-perceives the human, and is able to estimate, at least partially, what the human
-perceives or not. We infer that a human focuses on some object because it points at it, looks at it, and besides, the object is visible to him. This relies on the embodied
-nature of the interaction. In turn, this allows us to understand the
-meaning of sentences like ``Give me that''.
+perceives the human, and is able to estimate, at least partially, what the
+human perceives or not. We infer that a human focuses on some object because
+he/she points at it, looks at it, and besides, the object is visible to him.
+This relies on the embodied nature of the interaction. In turn, this allows us
+to understand the meaning of sentences like ``Give me that''.
 

### 2011-10-21 paper_knowledge.tex
Added a footnote to explain behaviour with definite/undefinite articles

@@
 However both are visible and the cognitive ambiguity resolution cannot be
-applied. The only option is to ask Jerry which box he is referring to: ``Which
-box, the toolbox or the cardboard box?'' Jerry could now simply answer the
-question. Instead, he decides to point at it while indicating: ``That box''
-(Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
-cardBoardBox} as being pointed at and looked at by the human and updates the
-ontology with this new information using a rule available in the commonsense
-ontology:
+applied. The only option is to ask Jerry which box he is referring
+to\footnote{Note that Jerry is using the definite article \emph{the}: the robot
+has to determine which box is relevant. {\sc Dialogs} supports other kinds of
+quantification (existantial -- \emph{some} --, universal -- \emph{all} -- , explicit cardinality), as expressed by the combination of definite/indefinite and
+singular/plural forms of articles.}: ``Which box, the toolbox or the cardboard box?''
+Jerry could now simply answer the question. Instead, he decides to point at it
+while indicating: ``That box'' (Figure~\ref{fig|pointing}b). The robot's
+perception identifies the {\tt cardBoardBox} as being pointed at and looked
+at by the human and updates the ontology with this new information using a
+rule available in the commonsense ontology:
+
 \begin{center}

### 2011-10-21 paper_knowledge.tex
Fixed one more 'top-bottom' mistake

@@
 
-For us, \emph{grounding} is on the contrary a \emph{top-bottom} activity: the
+For us, \emph{grounding} is on the contrary a \emph{top-down} activity: the
 robot needs to automatically bind a representation (for instance, a word

### 2011-10-21 paper_knowledge.tex
Make clearer that we use tag to initialize knowledge

@@
 \subsection{Informational Content Extraction}
+\label{informational_content_extraction}
 
@@
 In this first simple case, the concept \concept{banana\_01} is unambiguously
-matched (since there is only one possible banana) and returned. Finally, we can now add
-the new information provided by the human, \ie the new statement
-\stmt{banana\_01 hasSize big}, to the human model in the ontology
-server.
+matched (since there is only one possible banana) and returned. Finally, we can
+now add the new information provided by the human, \ie the new statement
+\stmt{banana\_01 hasSize big}, to the human model in the ontology server.
 
@@
 % adjectives (big, small, etc.), the \concept{hasSize} predicate is being used.
+
+We discuss in section~\ref{modeling_real_world} how the initial knowledge (the
+existence of a yellow banana) can be acquired.
 
@@
 \subsubsection{Modeling the Real World}
+\label{modeling_real_world}
 
@@
 
+Note also that, as mentioned in section~\ref{spygame}, perception issues have
+been skipped in our experiments by using a tag-based object identification
+method. In section~\ref{informational_content_extraction}, we give an example
+where the human says ``the yellow banana is big''. It is assumed in the example
+that the robot already knows about a banana instance that is yellow. In our
+experiments, this kind of knowledge was either hard coded in scenario-specific
+ontologies (with facts like \stmt{BANANA1 type Banana} where \concept{BANANA1}
+is the id of the banana's tag) or taught to the robot with prescriptive
+sentences like ``Learn that this is a banana'' while pointing at the banana's
+tag.
+
+It would be interesting to broaden this approach with automatic classifiers
+(for colour, size, etc.). If the robot later discovers a yellowish and large
+object, an utterance like ``the yellow banana is big'' could be used to assert
+that this very object is in fact a banana. A similar approach has been 
+developed by~\cite{Vrecko2009}.
+
 Amongst the other shortcomings of our architecture, neither the \emph{domain of

### 2011-10-21 paper_knowledge.tex
Raquel contribution to try to cut down the number of pages

--- 2011-10-21 paper_knowledge.tex
@@
 In this work we propose an approach aiming at grounding the interaction with
-users. Figure~\ref{fig|overall-process} illustrates the general contex. We have
-developed three distinct, inter-related cognitive components, as illustrated in
-Figure~\ref{fig|main-components}.
+users (Figure~\ref{fig|overall-process} illustrates the general context). To this end, we have
+developed three distinct, inter-related cognitive components:
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!t]
 \centering
@@
 
-\begin{figure}[!ht]
-\centering
-  \includegraphics[width=0.9\linewidth]{images/grounding.png}
-  \caption{The three main components of our architecture and their connections. Submodules are presented in subsequent sections.}
-  \label{fig|main-components}
-\end{figure}
+%\begin{figure}[!ht]
+%\centering
+%  \includegraphics[width=0.9\linewidth]{images/grounding.png}
+%  \caption{The three main components of our architecture and their connections. Submodules are presented in subsequent sections.}
+%  \label{fig|main-components}
+%\end{figure}
 
@@
 visible (otherwise the user would not be searching for it).
-In Figure~\ref{fig::vis} the field of view of a person is illustrated with a
+In Figure~\ref{fig::sparkRepresentations}\emph{a} the field of view of a person is illustrated with a
 grey cone (broader one). While he is able to see the two small boxes on the
@@
 on, \ie where its focus of attention is directed. This model is based on a
-narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
+narrower field of view, the field of attention (FOA). 
+Figure~\ref{fig::sparkRepresentations}\emph{a}
 shows the field of attention of a person with a green cone (narrower one). In
@@
 
-\begin{figure}%[!t] 
-\centering
-\includegraphics[width=\columnwidth]{images/looks.jpg} \caption{Representation
-of the field of view (FOV) and the field of attention (FOA) of the human.}
-\label{fig::vis} \end{figure}
+%\begin{figure}%[!t] 
+%\centering
+%\includegraphics[width=0.8\columnwidth]{images/looks.jpg} \caption{Representation
+%of the field of view (FOV) and the field of attention (FOA) of the human.}
+%\label{fig::vis} 
+%\end{figure}
 
@@
 point where the user is able to get the object afterward. 
-Figure~\ref{fig::reach} shows different reachability postures for each object
+Figure~\ref{fig::sparkRepresentations}\emph{b} shows different reachability postures for each object
 on the table. In the example, the bottle and the box are both reachable for the
@@
 
-\begin{figure}%[!t] 
-	\centering
-	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
-	\caption{Different reaching postures for the human: the bottle and the 
-	box are reachable, while the teddy bear is not.} 
-	\label{fig::reach} 
-\end{figure}
+%\begin{figure}%[!t] 
+%	\centering
+%	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
+%	\caption{Different reaching postures for the human: the bottle and the 
+%	box are reachable, while the teddy bear is not.} 
+%	\label{fig::reach} 
+%\end{figure}
 
 \end{itemize}
+
+\begin{figure*}[!t]
+	\subfigure[]
+	{\includegraphics[width=0.35\linewidth]{images/looks.jpg} \label{(a)}} \quad
+	\subfigure[]
+	{\includegraphics[width=0.3\linewidth]{images/reach.jpg} \label{(b)}}\qquad	
+	\subfigure[]
+	{\includegraphics[width=0.25\linewidth]{images/spatial_relation.png} \label{(c)}}
+
+	\caption{(a) Field of view (FOV) and the field of attention (FOA) of the human.
+(b) Different reaching postures for the human.
+(c) Spatial relations between two objects: A) \concept{isOn} relation, B) \concept{isIn} relation, and C) \concept{isNextTo} relation.}
+	\label{fig::sparkRepresentations}
+\end{figure*}
 
@@
 next to, in}, etc. In this work we compute three main relations based on the
-bounding box and center of mass of the objects (Figure~\ref{fig::sprelation}):
+bounding box and center of mass of the objects 
+(Figure~\ref{fig::sparkRepresentations}\emph{c}):
 
@@
 
-\begin{figure} 
-	\centering
-	\includegraphics[width=0.8\columnwidth]{images/spatial_relation.png}
-	\caption{Spatial relations between two objects: A) \concept{isOn} relation, 
-	B) \concept{isIn} relation, and C) \concept{isNextTo} relation.} 
-	\label{fig::sprelation} 
-\end{figure}
+%\begin{figure} 
+%	\centering
+%	\includegraphics[width=0.6\columnwidth]{images/spatial_relation.png}
+%	\caption{Spatial relations between two objects: A) \concept{isOn} relation, 
+%	B) \concept{isIn} relation, and C) \concept{isNextTo} relation.} 
+%	\label{fig::sprelation} 
+%\end{figure}
 
@@
 
-To this end we have developed a dedicated module called {\sc Dialogs} that processes human
-input in natural language, grounds the concepts in the robot's knowledge and
-eventually translates the discourse in a set of queries or declarative OWL/RDF
-statements.
-
-\begin{figure}%[!ht]
+\begin{figure}[!t]
 \centering
@@
 
-Figure~\ref{fig|dialog} shows the {\sc Dialogs} module architecture. The user's
+To this end we have developed a dedicated module called {\sc Dialogs} that processes human
+input in natural language, grounds the concepts in the robot's knowledge and
+eventually translates the discourse in a set of queries or declarative OWL/RDF
+statements. Figure~\ref{fig|dialog} shows the {\sc Dialogs} module architecture. The user's
 input is first pre-processed. For instance, \emph{I'm} constructs are expanded
@@
 based on three examples. In these examples we assume that some initial facts 
-are present in the knowledge base, both in the robot's own model and in the 
-human's model.  Since the robot tries to ground a
-human utterance, all queries are sent to the human model in order to interpret
-it from the human perspective.
+are present in the knowledge base (section~\ref{modeling_real_world} discusses how 
+the initial knowledge can be acquired), both in the robot's own model and in the human's model.
+Since the robot tries to ground a human utterance, all queries are sent to the human model in order to interpret it from the human perspective.
 
 \subsection{Informational Content Extraction}
+\label{informational_content_extraction}
+
+%\begin{figure}
+%    \centering
+%	\begin{tabular}{p{7cm}}
+%	\emph{Initial knowledge model of} \texttt{human\_01}\\
+%	\hline
+%    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
+%    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
+%	
+%	\vspace{0.5em}
+%	\emph{Human input}\\
+%	\hline
+%	\hspace{0.3cm}``The yellow banana is big!'' \\
+%
+%	\vspace{0.5em}
+%	\emph{Generated partial statements}\\
+%	\hline
+%	\hspace{0.3cm}\stmt{?obj type Banana} \\
+%    	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
+%    	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
+%
+%	\vspace{0.5em}
+%	\emph{Newly created statements}\\
+%	\hline
+%	\hspace{0.3cm}\stmt{banana\_01 hasSize big} \\
+%	\end{tabular}
+%\caption{First example of natural language grounding: the nominal group ``the
+%yellow banana'' is matched with the individual \concept{banana\_01}.
+%``$\Rightarrow$'' represents the output of the ontology server.}
+%\label{dialog|ex1}
+%\end{figure}
 
@@
     \centering
-	\begin{tabular}{p{7cm}}
-	\emph{Initial knowledge model of} \texttt{human\_01}\\
+%	\begin{tabular}{p{0.5\columnwidth} | p{0.5\columnwidth}}}
+	\begin{tabular}{l|l}
+	\emph{Initial knowledge} \texttt{human\_01} &
+	\emph{Human input}\\
+
 	\hline
-    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
-    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
-
-	\vspace{0.5em}
-	\emph{Human input}\\
+
+    	\stmt{banana\_01 type Banana} &
+	``The yellow banana is big!'' \\
+
+    	\stmt{banana\_01 hasColor yellow} & \\
+	\vspace{0.5em}\\
 	\hline
-	\hspace{0.3cm}``The yellow banana is big!'' \\
-
-	\vspace{0.5em}
-	\emph{Generated partial statements}\\
-	\hline
-	\hspace{0.3cm}\stmt{?obj type Banana} \\
-    	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
-    	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
-
-	\vspace{0.5em}
+
+	\emph{Generated partial statements} &
 	\emph{Newly created statements}\\
 	\hline
-	\hspace{0.3cm}\stmt{banana\_01 hasSize big} \\
+
+	\stmt{?obj type Banana} &
+	\hspace{0.2cm}\stmt{banana\_01 hasSize big} \\
+
+    	\stmt{?obj hasColor yellow} & \\
+    	\hspace{0.2cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
+
+	\hline
 	\end{tabular}
-\caption{First example of natural language grounding: the nominal group ``the
-yellow banana'' is matched with the individual \concept{banana\_01}.
+\caption{First example: content extraction.
 ``$\Rightarrow$'' represents the output of the ontology server.}
@@
 
+%\begin{figure}
+%    \centering
+%	\begin{tabular}{p{7cm}}
+%	\emph{Initial knowledge model of} \texttt{human\_01}\\
+%	\hline
+%    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
+%    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
+%	\end{tabular} \\
+%
+%	\vspace{0.5em}
+%
+%	\begin{tabular}{p{7cm}}
+%	\emph{Human input}\\
+%	\hline
+%    	\hspace{0.3cm}``Give me the banana.'' \\
+%	\end{tabular} \\
+%
+%	\vspace{0.5em}
+%
+%	\begin{tabular}{p{7cm}}
+%	\emph{Generated partial statements}\\
+%	\hline
+%    	\hspace{0.3cm}\stmt{?obj type Banana} \\
+%	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
+%
+%	\end{tabular} \\
+%
+%	\vspace{0.5em}
+%
+%	\begin{tabular}{p{7cm}}
+%	\emph{Newly created statements}\\
+%	\hline
+%    	\hspace{0.3cm}\stmt{human\_01 desires situation\_a3f74} \\
+%    	\hspace{0.3cm}\stmt{situation\_a3f74 type Give} \\
+%    	\hspace{0.3cm}\stmt{situation\_a3f74 performedBy myself} \\
+%    	\hspace{0.3cm}\stmt{situation\_a3f74 actsOnObject banana\_01} \\
+%    	\hspace{0.3cm}\stmt{situation\_a3f74 receivedBy human\_01} \\
+%	\end{tabular}
+%
+%\caption{Second example: processing an order.}
+%\label{dialog|ex2}
+%\end{figure}
+
 \begin{figure}
     \centering
-	\begin{tabular}{p{7cm}}
-	\emph{Initial knowledge model of} \texttt{human\_01}\\
+	\begin{tabular}{l|l}
+	\emph{Initial knowledge} \texttt{human\_01} &
+	\emph{Human input}\\
+
 	\hline
-    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
-    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
-	\end{tabular} \\
-
-	\vspace{0.5em}
-
-	\begin{tabular}{p{7cm}}
-	\emph{Human input}\\
+
+    	\stmt{banana\_01 type Banana} &
+	``Give me the banana.'' \\
+
+    	\stmt{banana\_01 hasColor yellow} & \\
+	\vspace{0.5em}\\
 	\hline
-    	\hspace{0.3cm}``Give me the banana.'' \\
-	\end{tabular} \\
-
-	\vspace{0.5em}
-
-	\begin{tabular}{p{7cm}}
-	\emph{Generated partial statements}\\
-	\hline
-    	\hspace{0.3cm}\stmt{?obj type Banana} \\
-	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
-
-	\end{tabular} \\
-
-	\vspace{0.5em}
-
-	\begin{tabular}{p{7cm}}
+
+	\emph{Generated partial statements} &
 	\emph{Newly created statements}\\
 	\hline
-    	\hspace{0.3cm}\stmt{human\_01 desires situation\_a3f74} \\
-    	\hspace{0.3cm}\stmt{situation\_a3f74 type Give} \\
-    	\hspace{0.3cm}\stmt{situation\_a3f74 performedBy myself} \\
-    	\hspace{0.3cm}\stmt{situation\_a3f74 actsOnObject banana\_01} \\
-    	\hspace{0.3cm}\stmt{situation\_a3f74 receivedBy human\_01} \\
+    	\stmt{?obj type Banana} & 
+	\stmt{human\_01 desires sit\_a3} \\
+
+	\hspace{0.2cm}$\Rightarrow$ \concept{?obj = banana\_01}
+    	& \stmt{sit\_a3 performedBy myself} \\
+    	& \stmt{sit\_a3 actsOnObject banana\_01} \\
+    	& \stmt{sit\_a3 receivedBy human\_01} \\
 	\end{tabular}
@@
 environment, while the robot queries the human until either discovering the
-object or giving up~\cite{Ros2010a}. A categorization example is presented in
-Figure~\ref{fig|objectsSpyGame}. The game starts with the human user giving a
+object or giving up~\cite{Ros2010a}. The game starts with the human user giving a
 first hint (communication is done through a keyboard and screen), allowing the
@@
 
-\begin{figure}%[!ht]
-\centering
-\begin{scriptsize}
-\begin{tikzpicture} %[level/.style={sibling distance=30mm/#1},scale=0.8]
-	%[edge from parent fork down,
-	%every node/.style={fill=black!30,rounded corners},
-	%[parent anchor=east,child anchor=west,grow=east,
-	%edge from parent/.style={thick,draw}]
-	\node {Artifact}
-	child {node {Tableware}
-		child {node {Bottle}}
-		child {node {Cup}}
-		}
-	child {node {GameObject}}
-	child {node {Furniture}
-			child {node {Table}}
-			child {node {Chair}}
-			child {node {Shelf}}};
-\end{tikzpicture}
-\end{scriptsize}
-\caption{Example of object categorization used in the scenario.}
-\label{fig|objectsSpyGame}			
-\end{figure}
+%\begin{figure}%[!ht]
+%\centering
+%\begin{scriptsize}
+%\begin{tikzpicture} %[level/.style={sibling distance=30mm/#1},scale=0.8]
+%	%[edge from parent fork down,
+%	%every node/.style={fill=black!30,rounded corners},
+%	%[parent anchor=east,child anchor=west,grow=east,
+%	%edge from parent/.style={thick,draw}]
+%	\node {Artifact}
+%	child {node {Tableware}
+%		child {node {Bottle}}
+%		child {node {Cup}}
+%		}
+%	child {node {GameObject}}
+%	child {node {Furniture}
+%			child {node {Table}}
+%			child {node {Chair}}
+%			child {node {Shelf}}};
+%\end{tikzpicture}
+%\end{scriptsize}
+%\caption{Example of object categorization used in the scenario.}
+%\label{fig|objectsSpyGame}			
+%\end{figure}
 
@@
 
-\begin{figure}%[!t]
+\begin{figure*}
 \centering
 	\subfigure[]
-  {\includegraphics[width=0.95\linewidth]{images/situationAssessment1.png}\label{a}}
+	{\includegraphics[width=0.3\linewidth]{images/situationAssessment1.png}\label{(a)}}
 	\subfigure[]
-  {\includegraphics[width=0.95\linewidth]{images/situationAssessment2.png}\label{b}}
-	\caption{Situation assessment during initialization: a) before placing the
-	objects on the table, and b) after placing them. In each image the snapshots
-	correspond to: real environment (top-left sub-image); processed image to
-	identify the tagged objects (bottom-left sub-image); and the 3D environment
-	(right sub-image).}
-	\label{fig|sitAssessRoman}
-\end{figure}
+	{\includegraphics[width=0.3\linewidth]{images/situationAssessment2.png}\label{(b)}}
+	\subfigure[]
+  	{\includegraphics[width=0.3\linewidth]{images/inTheBox2.jpg}\label{c}}
+	\caption{Situation assessment during initialization (before (a) and after (b) placing the objects on the table).
+In each image the snapshots correspond to: real environment (top-left sub-image); processed image to identify the tagged objects (bottom-left sub-image); and the 3D environment (right sub-image).(c) Jerry asks Jido for the content of the box while pointing at it.}
+	\label{fig|movingToLondon}
+\end{figure*}
+
+%\begin{figure}%[!t]
+%\centering
+%  \subfigure[]
+%  {\includegraphics[width=0.95\linewidth]{images/situationAssessment1.png}\label{a}}
+%  \subfigure[]
+%  {\includegraphics[width=0.95\linewidth]{images/situationAssessment2.png}\label{b}}
+%	\caption{Situation assessment during initialization: a) before placing the
+%	objects on the table, and b) after placing them. In each image the snapshots
+%	correspond to: real environment (top-left sub-image); processed image to
+%	identify the tagged objects (bottom-left sub-image); and the 3D environment
+%	(right sub-image).}
+%	\label{fig|sitAssessRoman}
+%\end{figure}
 
@@
 The initial information in ORO corresponds to:
-
 \begin{center}
-    \begin{tabular}{l}
-      \stmt{table type Table}\\
-      \stmt{cardBoardBox type Box}\\
-      \stmt{toolbox type Box}\\
-      \stmt{videoTape1 type VideoTape}\\
-      \stmt{videoTape1 label "Lord of the Robots"}\\
-      \stmt{videoTape2 type VideoTape}\\
-      \stmt{videoTape2 label "Jido-E"}\\
-      \stmt{Tom type Human}\\
-      \stmt{Jerry type Human}
+    \begin{tabular}{ll}
+      \stmt{table type Table} & \stmt{Tom type Human}\\
+      \stmt{cardBoardBox type Box} & \stmt{Jerry type Human}\\
+      \stmt{toolbox type Box} & \stmt{videoTape1 type VideoTape}\\
+      \stmt{videoTape2 type VideoTape} & \stmt{videoTape2 label "Jido-E"}   
       \end{tabular}
+      \stmt{videoTape1 label "Lord of the Robots"}
 \end{center}
@@
 \end{center}
-
-Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a) and places them around. Then he leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:
-
+Tom enters carrying several objects (Figure~\ref{fig|movingToLondon}a) and places them around. Then he leaves. (Figure~\ref{fig|movingToLondon}b). The following facts are computed and sent to ORO:
 \begin{center}
-    \begin{tabular}{l}
-      \stmt{toolbox isOn table}\\
-      \stmt{videoTape1 isOn table}\\
-      \stmt{videoTape1 isNextTo toolbox}\\
-      \stmt{toolbox isNextTo videoTape1}\\
-      \stmt{videoTape2 isIn cardBoardBox}
+    \begin{tabular}{ll}
+      \stmt{toolbox isOn table} & \stmt{toolbox isNextTo videoTape1}\\
+      \stmt{videoTape1 isOn table} & \stmt{videoTape2 isIn cardBoardBox}
       \end{tabular}
+      \stmt{videoTape1 isNextTo toolbox}
 \end{center}
@@
 identify the object the human is referring to:
-
 \begin{center} 
@@
 \end{center}
-
 There are two video tapes in the scene: one on
@@
 base returns both: 
-
 \begin{center}
@@
 \end{center}
-
 However, only one is visible for Tom (the one on the
@@
 
-\begin{figure}%[!ht]
-  \centering
-  \subfigure[]
-  {\includegraphics[width=0.9\linewidth]{images/inTheBox1.jpg}\label{a}}
-  \subfigure[]
-  {\includegraphics[width=0.9\linewidth]{images/inTheBox2.jpg}\label{b}}
-\caption{(a) Jerry asks Jido for the content of the box and (b) points at it.}
-  \label{fig|pointing}
-\end{figure}
-
-Figure~\ref{fig|pointing}a depicts the third situation where Jerry enters the
+%\begin{figure}%[!ht]
+%  \centering
+%  \subfigure[]
+%  {\includegraphics[width=0.9\linewidth]{images/inTheBox1.jpg}\label{a}}
+%  \subfigure[]
+%  {\includegraphics[width=0.9\linewidth]{images/inTheBox2.jpg}\label{b}}
+%\caption{(a) Jerry asks Jido for the content of the box and (b) points at it.}
+%  \label{fig|pointing}
+%\end{figure}
+
+In this last situation Jerry enters the
 living room without knowing where Tom had placed the video tapes. So he first
@@
 situation, there are two available boxes: 
-
 \begin{center}
@@
 \end{center}
-
 However both are visible and the cognitive ambiguity resolution cannot be
@@
 Jerry could now simply answer the question. Instead, he decides to point at it
-while indicating: ``That box'' (Figure~\ref{fig|pointing}b). The robot's
+while indicating: ``That box'' (Figure~\ref{fig|movingToLondon}c). The robot's
 perception identifies the {\tt cardBoardBox} as being pointed at and looked
@@
 rule available in the commonsense ontology:
-
 \begin{center}
-\begin{tabular}{l}
-\stmt{Jerry pointsAt carboardBox}\\
+\begin{tabular}{r}
+\stmt{Jerry pointsAt carboardBox}, 
 \stmt{Jerry looksAt carboardBox}\\
@@
 video tape which they were just talking about, but to the other one:
-
 \begin{center}
@@
 next to the toolbox.''
-
 \begin{center}
 \begin{tabular}{l}
-\stmt{videoTape1 isOn table}\\
+\stmt{videoTape1 isOn table}, 
 \stmt{videoTape1 isNextTo toolbox}
@@
 
+Note also that perception issues have been solved in our experiments by using a tag-based object identification
+method. In section~\ref{informational_content_extraction} we give an example
+where the human says ``the yellow banana is big''. It is assumed in the example
+that the robot already knows about a banana instance that is yellow. In our
+experiments, this kind of knowledge was either hard coded in scenario-specific
+ontologies (\eg \stmt{banana\_01 type Banana} where \concept{banana\_01}
+is the id of the banana's tag) or taught to the robot with prescriptive
+sentences like ``Learn that this is a banana'' while pointing at the banana's
+tag. It would be interesting to extend this approach with automatic classifiers
+(for colour, size, etc.). If the robot later discovers a yellowish and large
+object, an utterance like ``the yellow banana is big'' could be used to assert
+that this object is a banana.  A similar approach focused on the combination of visual perception and communication modalities to achieve visual learning has been 
+developed by~\cite{Vrecko2009}.
+
 While the examples we develop are all based on symbols that have a physical
@@
 validity} nor the context of a fact are represented in a satisfying way (we do
-store some kind of context --the agent's mental model for instance--). This
+store some kind of context --the agent's mental model for instance). This
 information is meta-information on the knowledge. While the ORO framework
